---
title: "ETC5521 Diving Deeper into Data Exploration: Project Part 2"
author: "P2-wallaby"
date: "Oct 25, 2025"
quarto-required: ">=1.8.0"
format:
    html: 
        output-file: report_p2.html
        css: "assignment.css"
        theme: cerulean
    
        code-fold: true
        code-summary: "Show code"
        code-line-numbers: false
        code-tools: false
        code-copy: false
    
        #big size save
        embed-resources: false
        self-contained: false
    
        #trim JS helpers
        toc: true
        toc-depth: 3
        toc-location: left
        anchor-sections: false
        smooth-scroll: false
    
        #smaller images
        fig-width: 6
        fig-height: 4
        fig-dpi: 72
    
        #lighter code highlighting
        highlight-style: github
        df-print: paged
editor: 
  markdown: 
    wrap: sentence
---

```{r}
#| echo: false
# Set up chunk for all slides
knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 4,
  fig.align = "center",
  out.width = "100%",
  code.line.numbers = FALSE,
  fig.retina = 4,
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = FALSE,
  dev.args = list(pointsize = 11)
)
```

```{r}
#| echo: false
# Load libraries
library(tidyverse)
library(conflicted)

conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::select)
conflicts_prefer(dplyr::slice)

# Plot options and themes
options(
  digits = 2,
  width = 60,
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis",
  ggplot2.discrete.colour = c("#D55E00", "#0072B2", "#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442"),
  ggplot2.discrete.fill = c("#D55E00", "#0072B2", "#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442")
)

theme_set(theme_bw(base_size = 14) +
   theme(
     aspect.ratio = 1,
     plot.background = element_rect(fill = 'transparent', colour = NA),
     plot.title.position = "plot",
     plot.title = element_text(size = 24),
     panel.background = element_rect(fill = 'transparent', colour = NA),
     legend.background = element_rect(fill = 'transparent', colour = NA),
     legend.key = element_rect(fill = 'transparent', colour = NA)
   )
)
```

This is the detail library that we used for this project.

```{r load-library}
library(dplyr)
library(stringr)
library(janitor)
library(skimr)
library(tsibble)
library(knitr)
library(naniar)
library(patchwork)
library(plotly)
library(tsibbletalk)
library(feasts)
library(broom)
library(purrr)
library(cassowaryr)
library(gganimate)
library(gifski)
library(kableExtra)
library(ggplot2)
library(scales)
library(tidyr)
library(ggrepel)
library(viridisLite)
```

## Introduction

AAA

## Data description

The dataset used in this analysis originates from **the Bureau van Dijk OSIRIS database**, a comprehensive source of harmonised financial information for publicly listed and major private firms worldwide (Bureau van Dijk, 2025).
For this study, we extracted firm-level data for **German companies** over the **2018–2021 accounting years**, a period that captures both pre-pandemic and pandemic-era financial performance.
Each record represents a firm’s reported financials for a specific accounting year, enabling longitudinal exploration of trends in profitability, liquidity, leverage, and resilience.
These are the code that we used to load the data and combine them:

Loading the data.

```{r read-data}
#define the folder
data_directory <- "osiris"

#filter germany and sweden
germany_sweden_file_paths <- list.files(
  data_directory,
  pattern = "^osiris_?(Germany|Sweden)_(2018|2019|2020|2021)\\.rda$",
  full.names = TRUE,
  ignore.case = TRUE
)

#name the list by file stems
germany_sweden_file_stems <- tools::file_path_sans_ext(basename(germany_sweden_file_paths))
osiris_germany_sweden <- setNames(vector("list", length(germany_sweden_file_paths)), germany_sweden_file_stems)

#load each file
for (file_index in seq_along(germany_sweden_file_paths)) {
  temporary_environment <- new.env(parent = emptyenv())
  load(germany_sweden_file_paths[file_index], envir = temporary_environment)
  loaded_objects <- as.list(temporary_environment)

  #If the file holds one object, store that object, otherwise store the sub-list
  osiris_germany_sweden[[file_index]] <- if (length(loaded_objects) == 1) loaded_objects[[1]] else loaded_objects
}

```

Combining the data.

```{r read-data-2}
#build a single data frame with 'year' and 'source_country' taken from the filename
combined_germany_sweden <- {
  parts <- vector("list", length(osiris_germany_sweden))
  names(parts) <- names(osiris_germany_sweden)

  for (file_stem in names(osiris_germany_sweden)) {
    data_frame_from_file <- osiris_germany_sweden[[file_stem]]

    extracted_year     <- as.integer(str_extract(file_stem, "\\d{4}"))
    extracted_country  <- str_extract(file_stem, "(?i)Germany|Sweden") |> stringr::str_to_title()

    parts[[file_stem]] <- mutate(
      data_frame_from_file,
      year = extracted_year,
      source_country = extracted_country #avoid clashing with the raw 'country'
    )
  }

  #bind and sort
  bind_rows(parts) |>
    arrange(source_country, year)
}

#unique country–year pairs
combined_germany_sweden |>
  dplyr::distinct(source_country, year) |>
  dplyr::arrange(source_country, year) |>
  kable(align = "l", booktabs = TRUE) |>
  kable_styling(full_width = FALSE, font_size = 11)

```

After the data is combined, we standardise the raw file into analysis-ready fields.
It normalises column names, maps long vendor headers to clear variables by topic, and creates a fiscal-year label from the year-end date.
Finally, it keeps only the identifiers, industry codes, and financial variables needed for the analysis.

```{r data-cleaning}
filtered_germany_sweden <- combined_germany_sweden |>
  clean_names() |>
  
  rename(
    # IDs / scope (used in all questions)
    company_name       = company_name_name,
    company_id         = name_id,
    acct_year          = year,
    country            = country_country,
    city               = city_city_city,
    consolidation_code = consolidation_code_consol_code,
    status             = status_status,
    
    # Industry codes (Q1, Q2, Q4)
    nace_code = nace_rev_1_core_code_cnacecd,
    icb_code  = industrial_classification_benchmark_icb,
    sic_code  = us_sic_core_code_csicuscde,
    
    # Scale / balance sheet (Q2, Q3)
    total_assets_eur      = total_assets_data13077,
    total_liabilities_eur = total_liabilities_and_debt_data14022,
    total_equity_eur      = total_shareholders_equity_data14041,
    
    # Revenue & profit (Q1, Q3)
    total_revenue_eur = total_revenues_data13004,
    net_sales_eur     = net_sales_data13002,
    gross_sales_eur   = gross_sales_data13000,
    net_income_eur    = net_income_starting_line_data15500,
    
    # Profitability ratios (Q1, Q4)
    ebit_margin_pct   = ebit_margin_percent_data31055,
    ebitda_margin_pct = ebitda_margin_percent_data31060,
    roa_pct           = return_on_total_assets_percent_data31015,
    roe_pct           = roe_percent_data31065,
    
    # Efficiency
    net_assets_turnover = net_assets_turnover_data31225,
    stock_turnover      = stock_turnover_data31220,
    
    # Leverage, solvency, and liquidity (Q2)
    solvency_pct        = solvency_ratio_percent_data31310,
    gearing_pct         = gearing_percent_data31315,
    current_ratio       = current_ratio_data31105,
    liquidity_ratio     = liquidity_ratio_data31110,
    shareholders_liq_pct= shareholders_liquidity_ratio_data31305,
    interest_cover      = interest_cover_data31115
  ) |>
  
  # Rename fiscal year
  mutate(
    fy_end_date = lubridate::ymd(as.character(company_fiscal_year_end_date_closdate)),
    fy_year     = year(fy_end_date)
  ) |>
  
  # Select only relevant columns
  select(
    company_id, company_name, fy_end_date, fy_year, acct_year, status, country, city, consolidation_code,
    nace_code, icb_code, sic_code,
    total_assets_eur, total_liabilities_eur, total_equity_eur,
    total_revenue_eur, net_sales_eur, gross_sales_eur, net_income_eur,
    ebit_margin_pct, ebitda_margin_pct, roa_pct, roe_pct,
    net_assets_turnover, stock_turnover,
    solvency_pct, gearing_pct, current_ratio, liquidity_ratio,
    shareholders_liq_pct, interest_cover
  )

```

This is the table describing the variables we used in this project:

```{r table-description, echo=FALSE}
data_dictionary <- tribble(
  ~`Column Name`,           ~Description,                                                                    ~`Data Type`, ~`Original Name`,
  "company_id",             "Unique company identifier (Bureau van Dijk ID).",                               "character", "BvD ID Number (os_id_number)",
  "company_name",           "Registered company name (label).",                                               "character", "Company Name (name)",
  "fy_end_date",            "Company fiscal year-end date for the account.",                                  "date",      "Company Fiscal Year End Date (closdate)",
  "fy_year",                "Year extracted from fiscal year-end date.",                                      "integer",   "derived from Company Fiscal Year End Date (closdate)",
  "acct_year",              "Reporting year label carried from file name/field.",                             "integer",   "year",
  "country",                "Country of head office.",                                                         "character", "Country (country)",
  "city",                   "City of head office.",                                                            "character", "CITY - City (city)",
  "consolidation_code",     "Consolidation scope of the accounts (e.g., C1/C2/U1/U2).",                        "character", "Consolidation Code (consol_code)",
  "nace_code",              "NACE core code (EU industry classification).",                                    "character", "NACE Rev 1, Core Code (cnacecd)",
  "icb_code",               "ICB industry classification code.",                                               "character", "Industrial Classification Benchmark (icb)",
  "sic_code",               "US SIC core code.",                                                               "character", "US SIC, Core Code (csicuscde)",
  "total_assets_eur",       "Total assets (EUR).",                                                              "numeric",   "Total Assets (data13077)",
  "total_liabilities_eur",  "Total liabilities and debt (EUR).",                                               "numeric",   "Total Liabilities and Debt (data14022)",
  "total_equity_eur",       "Total shareholders’ equity (EUR).",                                               "numeric",   "Total Shareholders Equity (data14041)",
  "total_revenue_eur",      "Total revenues / operating revenue (EUR).",                                       "numeric",   "Total revenues (data13004)",
  "net_sales_eur",          "Net sales (EUR).",                                                                 "numeric",   "Net sales (data13002)",
  "gross_sales_eur",        "Gross sales (EUR).",                                                               "numeric",   "Gross sales (data13000)",
  "net_income_eur",         "Net income (EUR).",                                                                "numeric",   "Net Income / Starting Line (data15500)",
  "ebit_margin_pct",        "EBIT margin (% of sales/revenue).",                                               "numeric",   "EBIT Margin (%) (data31055)",
  "ebitda_margin_pct",      "EBITDA margin (% of sales/revenue).",                                             "numeric",   "EBITDA Margin (%) (data31060)",
  "roa_pct",                "Return on total assets (%).",                                                     "numeric",   "Return on Total Assets (%) (data31015)",
  "roe_pct",                "Return on equity (%).",                                                           "numeric",   "ROE (%) (data31065)",
  "net_assets_turnover",    "Net assets turnover (times).",                                                    "numeric",   "Net Assets Turnover (data31225)",
  "stock_turnover",         "Stock / inventory turnover (times).",                                             "numeric",   "Stock Turnover (data31220)",
  "solvency_pct",           "Solvency ratio (%).",                                                             "numeric",   "Solvency ratio (%) (data31310)",
  "gearing_pct",            "Gearing ratio (%).",                                                              "numeric",   "Gearing (%) (data31315)",
  "current_ratio",          "Current ratio (times).",                                                          "numeric",   "Current ratio (data31105)",
  "liquidity_ratio",        "Liquidity / quick ratio (times).",                                                "numeric",   "Liquidity ratio (data31110)",
  "shareholders_liq_pct",   "Shareholders’ liquidity ratio (%).",                                              "numeric",   "Shareholders Liquidity ratio (data31305)",
  "interest_cover",         "Interest cover (times).",                                                         "numeric",   "Interest Cover (data31115)"
)

#show table
data_dictionary |>
  kable(align = "l", booktabs = TRUE) |>
  kable_styling(full_width = FALSE, font_size = 11)
```

The selected variables encompass four key dimensions of corporate financial health:

-   **Identification and Location** - company_id, company_name, country, and city provide firm-level identifiers and geographical context.

<!-- -->

-   **Balance Sheet Structure** - total_assets_eur, total_liabilities_eur, and total_equity_eur measure firm size and capital composition, reported in euros.

<!-- -->

-   **Income and Profitability** - total_revenue_eur, net_sales_eur, and net_income_eur capture income statement outcomes, while ebit_margin_pct, ebitda_margin_pct, roa_pct, and roe_pct reflect operating performance and returns to shareholders.

<!-- -->

-   **Solvency and Liquidity** - solvency_pct, gearing_pct, current_ratio, liquidity_ratio, shareholders_liq_pct, and interest_cover assess each firm’s financial stability, leverage, and capacity to meet short-term obligations.

-   **Industry Classification** - industry_nace, industry_icb, and industry_group link firms to their respective economic sectors according to European (NACE) and global (ICB) coding systems.

It is important to note that the **accounting year** variable in **OSIRIS** reflects **two-year reporting periods**.
For example, the 2018 accounting year represents financial results spanning 2017–2018, 2019 corresponds to 2018–2019, 2020 covers 2019–2020, and 2021 includes 2020–2021.
Throughout this report, each accounting year is interpreted as representing the **endpoint of its reporting cycle**.
Consequently, “2021” is treated as capturing firm performance during the height of the **COVID-19 pandemic**, while 2018–2019 represent pre-pandemic benchmarks.

This dataset structure enables a robust exploration of temporal changes in corporate performance and the role of firm characteristics , particularly **firm size** , in shaping **financial resilience** throughout the pandemic period.

### Research Question

Here are the research questions (sub-questions) that guide us in analysing and answering the primary question: How did the pandemic affect corporate financials in Germany and Sweden?

**1. How did profitability change across industries?**

This question investigates how profitability variables such as ROA, EBITDA margin, and EBIT margin shifted across industries between the pre-pandemic period (2018 to 2019) and the pandemic period (2020 to 2021) in Germany and Sweden. It aims to identify which sectors saw profit declines, remained stable, or grew during the pandemic. By comparing profitability trends across industries, the analysis highlights which sectors proved more resilient to the economic impacts of the pandemic. These insights help explain how the pandemic affected overall corporate performance and provide a clearer understanding of the differences in financial recovery among industries.

**2. How did leverage (debt-to-assets) and liquidity (current ratio) evolve during pandemic?**

This question extends the Part 1 analysis by comparing **how leverage (debt-to-assets)** and **liquidity (current ratio)** evolved between **Germany and Sweden** from 2018 to 2021.\
The goal is to determine whether firms in the two economies displayed similar balance-sheet responses to the COVID-19 shock.\
This section applies techniques from **Lectures 4-6** (data summarisation, comparative visualisation, and faceting) and interprets results within a financial-resilience framework.

**3. Which German industries showed unexpected financial resilience or vulnerability from 2020 to 2022, and how were these outcomes shaped by profitability, leverage, liquidity, and firm size?**

AAA

**4. How did the fundamental relationship between corporate liquidity and profitability evolve and fracture within industries during the pandemic?**

AAA

## IDA

Here is the IDA process for this project:

First, we develop this code to prepare the dataset for analysis by converting numeric columns, standardising text fields, and reporting missing values overall and by year.
It also performs basic sanity checks for duplicate firm-years, impossible values, and consolidation codes.

```{r table-checking}

num_cols <- c(
  "total_assets_eur","total_liabilities_eur","total_equity_eur",
  "total_revenue_eur","net_sales_eur","gross_sales_eur","net_income_eur",
  "ebit_margin_pct","ebitda_margin_pct","roa_pct","roe_pct",
  "net_assets_turnover","stock_turnover",
  "solvency_pct","gearing_pct","current_ratio","liquidity_ratio",
  "shareholders_liq_pct","interest_cover"
)

#symbol formatting/coerce numerics
filtered_germany_sweden <- filtered_germany_sweden |>
  mutate(
    across(all_of(num_cols) & where(is.character), readr::parse_number),
    country = stringr::str_to_title(country),
    city    = stringr::str_to_title(city)
  )

#missingness report check (overall and by year)
missing_overall <- filtered_germany_sweden |>
  summarise(across(everything(), ~ mean(is.na(.)))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "missing_rate")

missing_by_year <- filtered_germany_sweden |>
  group_by(acct_year) |>
  summarise(across(everything(), ~ mean(is.na(.)))) |>
  pivot_longer(-acct_year, names_to = "variable", values_to = "missing_rate")

#sanity checks
dup_firm_year <- filtered_germany_sweden |>
  count(company_id, fy_year, acct_year, country) |>
  filter(n > 1)

impossible_assets <- filtered_germany_sweden |>
  filter(!is.na(total_assets_eur) & total_assets_eur < 0)

consol_levels <- filtered_germany_sweden |>
  count(consolidation_code, sort = TRUE)

```

Then, we create this code to extract the 2-digit NACE code, categorises companies into broad industry groups using NACE and ICB, and creates a unified 'industry_group' factor with a fixed order for plotting purposes.
It also provides a small sample of unmapped records for verification.
After checking, there are no duplicates in the data using company_id, fy_year, and acct_year as the primary key.

```{r NACE-column}
#clean 2-digit NACE as integer
filtered_germany_sweden <- filtered_germany_sweden |>
  mutate(
    nace_num2 = suppressWarnings(as.integer(str_sub(readr::parse_number(as.character(nace_code)), 1, 2)))
  )

#map NACE divisions to broad section names
industry_from_nace <- function(x){
  case_when(
    x %in% 1:3 ~ "Agriculture, Forestry & Fishing",
    x %in% 5:9 ~ "Mining & Quarrying",
    x %in% 10:33 | x %in% 15:37 ~ "Manufacturing",
    x %in% 35 ~ "Electricity, Gas, Steam",
    x %in% 36:39 ~ "Water Supply & Waste",
    x %in% 41:43 ~ "Construction",
    x %in% 45:47 ~ "Wholesale & Retail Trade",
    x %in% 49:53 ~ "Transport & Storage",
    x %in% 55:56 ~ "Accommodation & Food",
    x %in% 58:63 ~ "Information & Communication",
    x %in% 64:66 ~ "Financial & Insurance",
    x %in% 68 ~ "Real Estate",
    x %in% 69:75 ~ "Professional, Scientific & Technical",
    x %in% 77:82 ~ "Administrative & Support",
    x %in% 84 ~ "Public Administration",
    x %in% 85 ~ "Education",
    x %in% 86:88 ~ "Human Health & Social Work",
    x %in% 90:93 ~ "Arts, Entertainment & Recreation",
    x %in% 94:96 ~ "Other Service Activities",
    x %in% 97:98 ~ "Household Activities",
    x %in% 99 ~ "Extraterritorial Organizations",
    TRUE ~ NA_character_
  )
}

#ICB top-level sector labels
icb_map <- c(
  "0001" = "Oil & Gas",
  "1000" = "Basic Materials",
  "2000" = "Industrials",
  "3000" = "Consumer Goods",
  "4000" = "Health Care",
  "5000" = "Consumer Services",
  "6000" = "Telecommunications",
  "7000" = "Utilities",
  "8000" = "Financials",
  "9000" = "Technology"
)

#normalize ICB to a 4-digit “bucket” (example: 2573 to 2000)
normalize_icb <- function(x){
  x_chr <- str_extract(as.character(x), "\\d+")
  ifelse(is.na(x_chr), NA_character_,
         sprintf("%04d", as.integer(floor(as.numeric(x_chr) / 1000) * 1000)))
}

filtered_germany_sweden <- filtered_germany_sweden |>
  mutate(
    industry_nace = industry_from_nace(nace_num2),
    icb_bucket    = normalize_icb(icb_code),
    industry_icb  = icb_map[icb_bucket],
    industry_group = coalesce(industry_nace, industry_icb, "Other / Unmapped")
  )

#check
unmapped_sample <- filtered_germany_sweden |>
  filter(industry_group == "Other / Unmapped") |>
  select(company_id, fy_year, acct_year, nace_code, icb_code) |>
  head(15)

#factor order for plots
ordered_levels <- c(
  "Agriculture, Forestry & Fishing","Mining & Quarrying","Manufacturing",
  "Electricity, Gas, Steam","Water Supply & Waste","Construction",
  "Wholesale & Retail Trade","Transport & Storage","Accommodation & Food",
  "Information & Communication","Financial & Insurance","Real Estate",
  "Professional, Scientific & Technical","Administrative & Support",
  "Public Administration","Education","Human Health & Social Work",
  "Arts, Entertainment & Recreation","Other Service Activities",
  "Household Activities","Extraterritorial Organizations",
  "Oil & Gas","Basic Materials","Industrials","Consumer Goods","Health Care",
  "Consumer Services","Telecommunications","Utilities","Financials","Technology",
  "Other / Unmapped"
)
filtered_germany_sweden <- filtered_germany_sweden |>
  mutate(industry_group = factor(industry_group, levels = ordered_levels))
```

After that, we create this code to reshape numeric variables into a long format for Initial Data Analysis, separate scale variables from ratios, and assign clear labels for charts.
The result is a tidy table ready for profiling and visualisation.

```{r prepare-IDA}
#using long data
ida_germany_sweden <- filtered_germany_sweden |>
  dplyr::select(acct_year, country, dplyr::all_of(num_cols)) |>
  tidyr::pivot_longer(dplyr::all_of(num_cols),
                      names_to = "variable_orig", values_to = "value") |>
  tidyr::drop_na()

#split between scale and ratio
scale_vars <- c("total_assets_eur","total_liabilities_eur","total_equity_eur",
                "total_revenue_eur","net_sales_eur","gross_sales_eur","net_income_eur")
ratio_vars <- setdiff(unique(ida_germany_sweden$variable_orig), scale_vars)

#label mapping for the charts
nice <- c(
  total_assets_eur   = "Total assets (EUR)",
  total_liabilities_eur="Total liabilities (EUR)",
  total_equity_eur   = "Equity (EUR)",
  total_revenue_eur  = "Total revenue (EUR)",
  net_sales_eur      = "Net sales (EUR)",
  gross_sales_eur    = "Gross sales (EUR)",
  net_income_eur     = "Net income (EUR)",
  ebit_margin_pct    = "EBIT margin (%)",
  ebitda_margin_pct  = "EBITDA margin (%)",
  roa_pct            = "ROA (%)",
  roe_pct            = "ROE (%)",
  net_assets_turnover= "Net assets turnover (x)",
  stock_turnover     = "Stock turnover (x)",
  solvency_pct       = "Solvency (%)",
  gearing_pct        = "Gearing (%)",
  current_ratio      = "Current ratio (x)",
  liquidity_ratio    = "Liquidity ratio (x)",
  shareholders_liq_pct = "Shareholders’ liquidity (%)",
  interest_cover     = "Interest cover (x)"
)

pretty_var <- function(x) ifelse(x %in% names(nice), nice[x], x)

ida_germany_sweden <- ida_germany_sweden |>
  mutate(variable = pretty_var(variable_orig),
         variable = factor(variable, levels = unique(variable)))

```

Then, we create this code to visualise the distribution of scale variables by year using violin and jitter plots on a log scale, with medians highlighted.
It allows quick checks of spread, skewness, and outliers over time for monetary measures.

```{r IDA-scale}
#define variables sets
vars_balance <- c("total_assets_eur","total_liabilities_eur","total_equity_eur")
vars_flows   <- c("total_revenue_eur","net_sales_eur","gross_sales_eur","net_income_eur")
```

**Balanced Sheet Scale Variables**

```{r balanced-sheet-scale}
p_scale_balance <- ida_germany_sweden |>
  filter(variable_orig %in% vars_balance) |>
  mutate(
    y    = if_else(value > 0, value, NA_real_),
    year = factor(acct_year)
  ) |>
  ggplot(aes(x = year, y = y, colour = country,
             group = interaction(country, year))) +
  geom_violin(trim = FALSE, fill = NA, linewidth = 0.4, na.rm = TRUE,
              position = position_dodge(width = 0.75)) +
  geom_jitter(alpha = 0.12, size = 0.6, na.rm = TRUE,
              position = position_jitterdodge(jitter.width = 0.12,
                                              jitter.height = 0,
                                              dodge.width = 0.75)) +
  stat_summary(fun = median, geom = "point", shape = 95, size = 6, na.rm = TRUE,
               position = position_dodge(width = 0.75)) +
  scale_y_log10(labels = scales::label_number(scale_cut = scales::cut_short_scale())) +
  facet_wrap(~ variable, scales = "free_y", ncol = 3) +
  labs(x = "Year", y = NULL, title = "Violin and Jitter by Year (log scale) for Balance Sheet Variables") +
  theme_minimal(base_size = 11) +
  theme(
  plot.title.position = "plot",
  legend.position     = "top",
  legend.spacing.y    = unit(-0.6, "lines"),
  legend.justification = "center",
  legend.direction     = "horizontal",
  legend.box           = "horizontal",
  legend.margin        = margin(b = 4),
  axis.text.x          = element_text(size = 10),
  strip.text           = element_text(face = "bold"),
  panel.grid.minor     = element_blank()
)

p_scale_balance
```

According to the chart, most companies in both countries are at the lower end of the asset and liability spectrum, while a small number of very large firms form the long upper tails, indicating strong right skewness. The medians for total assets, liabilities, and equity remain stable over time, suggesting no significant structural shift in balance sheets between 2018 and 2021. Sweden consistently reports higher levels across the balance-sheet variables than Germany, reflecting generally larger firm sizes in the Swedish sample. The range widens in 2020 – 2021 for both countries, showing increased variation during the pandemic. Median equity remains positive and steady across both nations, suggesting that most firms maintained solvency despite the disruptions.

After that, we visualise ratio and percentage metrics by year on a linear scale using violin and jitter plots with median markers, faceted by variable.
It is used to assess distribution shape, central tendency, spread, and potential outliers across time without log transformation.

```{r IDA-ratio}
#chart for ratio variables (linear axis) — side-by-side by country
p_ratio <- ida_germany_sweden |>
  dplyr::filter(variable_orig %in% ratio_vars) |>
  dplyr::mutate(year = factor(acct_year)) |>
  ggplot(aes(x = year, y = value, colour = country,
             group = interaction(country, year))) +
  geom_violin(trim = FALSE, fill = NA, linewidth = 0.4,
              position = position_dodge(width = 0.75), na.rm = TRUE) +
  #jitter dodged by country
  geom_jitter(alpha = 0.12, size = 0.6, na.rm = TRUE,
              position = position_jitterdodge(jitter.width = 0.12,
                                              jitter.height = 0,
                                              dodge.width = 0.75)) +
  stat_summary(fun = median, geom = "point", shape = 95, size = 6, na.rm = TRUE,
               position = position_dodge(width = 0.75)) +
  facet_wrap(~ variable, scales = "free_y", ncol = 3) +
  labs(x = "Year", y = NULL,
       title = "Violin + Jitter by Year (Ratios & % on Linear Scale)") +
  theme_minimal(base_size = 11) +
  theme(axis.text.x = element_text(size = 10),
        strip.text  = element_text(face = "bold"),
        panel.grid.minor = element_blank())

p_ratio
```

AAA

Then, we measure missing data for key numeric variables by industry and by year, then visualise year-wise gaps with a heatmap.
It highlights variables or periods with substantial missingness, guiding imputation, exclusion, or robustness checks before analysis.

```{r numeric-missingness-check}
key_vars <- c("total_assets_eur","total_liabilities_eur","total_equity_eur",
              "total_revenue_eur","net_income_eur",
              "ebit_margin_pct","ebitda_margin_pct","roa_pct","roe_pct",
              "current_ratio","liquidity_ratio","gearing_pct","solvency_pct")

# missingness by industry and country
missing_by_industry <- filtered_germany_sweden |>
  group_by(country, industry_group) |>
  summarise(across(all_of(key_vars), ~ mean(is.na(.)), .names = "{.col}"),
            .groups = "drop") |>
  pivot_longer(-c(country, industry_group),
               names_to = "variable", values_to = "missing_rate")

#heatmap by year for numeric variables
missing_heat <- filtered_germany_sweden |>
  group_by(country, acct_year) |>
  summarise(across(all_of(key_vars), ~ mean(is.na(.)), .names = "{.col}"),
            .groups = "drop") |>
  pivot_longer(-c(country, acct_year),
               names_to = "variable", values_to = "missing_rate")

ggplot(missing_heat,
       aes(variable, factor(acct_year), fill = missing_rate)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "white", high = "red",
                      labels = scales::percent_format(accuracy = 1)) +
  facet_wrap(~ country, nrow = 1) +
  labs(x = NULL, y = "Year",
       title = "Missingness by Year & Country (Numeric Variables)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

AAA

After that, we used this code to detect outliers in all numeric variables with two rules - IQR (1.5×) and standardized scores (\|z\| \> 3) - and summarise counts and proportions per variable, ordered by the IQR share.
It flags variables with heavy tails or extreme values to guide winsorisation, transformation, or sensitivity checks before modelling.

```{r outliers-check}
num_cols <- c("total_assets_eur","total_liabilities_eur","total_equity_eur",
              "total_revenue_eur","net_sales_eur","gross_sales_eur","net_income_eur",
              "ebit_margin_pct","ebitda_margin_pct","roa_pct","roe_pct",
              "net_assets_turnover","stock_turnover",
              "solvency_pct","gearing_pct","current_ratio","liquidity_ratio",
              "shareholders_liq_pct","interest_cover")

qfun <- function(x,p) quantile(x, probs=p, na.rm=TRUE, names=FALSE)

outliers_by_country <- purrr::map_dfr(num_cols, function(v){
  x  <- filtered_germany_sweden[[v]]
  ct <- filtered_germany_sweden$country
  
  q1 <- qfun(x,.25); q3 <- qfun(x,.75); iqr <- q3-q1
  lo <- q1 - 1.5*iqr; hi <- q3 + 1.5*iqr
  z  <- as.numeric(scale(x))
  
  tibble::tibble(
    variable = v,
    country  = ct,
    n        = !is.na(x),
    iqr_flag = (x < lo | x > hi),
    z_flag   = abs(z) > 3
  )
}) |>
  dplyr::group_by(variable, country) |>
  dplyr::summarise(
    n = sum(n),
    iqr_outliers = sum(iqr_flag, na.rm = TRUE),
    z_outliers   = sum(z_flag, na.rm = TRUE),
    share_iqr    = iqr_outliers/n,
    share_z      = z_outliers/n,
    .groups = "drop"
  ) |>
  dplyr::arrange(country, desc(share_iqr))

var_levels <- intersect(num_cols, unique(outliers_by_country$variable))

outliers_by_country_ord <- outliers_by_country |>
  dplyr::mutate(variable = factor(variable, levels = var_levels)) |>
  dplyr::arrange(variable, country)

outliers_by_country_ord |>
  kable(align = "l", booktabs = TRUE) |>
  kable_styling(full_width = FALSE, font_size = 11)
```

AAA

Then, we standardise key character fields by trimming whitespace, normalising case, and converting blanks to NA, and report missingness for identifiers and classification codes.
This ensures consistent keys for joins and grouping, and highlights data quality issues to address before analysis.

```{r character-check}
#character variables to check
char_vars <- c("company_id","company_name","country","city",
               "consolidation_code","nace_code","icb_code","sic_code")
char_vars <- intersect(char_vars, names(filtered_germany_sweden))

#cleaning
chars_clean <- filtered_germany_sweden |>
  mutate(
    across(all_of(char_vars),
           ~ .x |> as.character() |> str_squish() |> na_if("")),
    #casing
    country = str_to_title(country),
    city    = str_to_title(city),
    company_name = str_squish(company_name)
  )

#missingness check
char_missing <- chars_clean |>
  summarise(across(all_of(char_vars), ~ mean(is.na(.)))) |>
  tidyr::pivot_longer(everything(), names_to="variable", values_to="missing_rate")

char_missing |>
  kable(align = "l", booktabs = TRUE) |>
  kable_styling(full_width = FALSE, font_size = 11)
```

AAA

After completing the general IDA, we developed the specific IDA to address our sub-questions.
Here are the details:

### Question 1

AAA

### Question 2

```{r q2-ida}
# Create leverage and liquidity indicators for both countries
lev_liq_data <- filtered_germany_sweden |>
  filter(acct_year %in% 2018:2021, status == "Active") |>
  mutate(
    debt_to_assets = total_liabilities_eur / total_assets_eur,
    equity_ratio   = total_equity_eur / total_assets_eur
  ) |>
  select(acct_year, country, industry_group,
         debt_to_assets, equity_ratio, gearing_pct,
         solvency_pct, current_ratio) |>
  drop_na(debt_to_assets, current_ratio)

# Summarise by year and country
lev_liq_summary <- lev_liq_data |>
  group_by(country, acct_year) |>
  summarise(
    median_debt_assets = median(debt_to_assets, na.rm = TRUE),
    iqr_debt_assets    = IQR(debt_to_assets, na.rm = TRUE),
    median_liquidity   = median(current_ratio, na.rm = TRUE),
    iqr_liquidity      = IQR(current_ratio, na.rm = TRUE),
    .groups = "drop"
  )

lev_liq_summary |>
  knitr::kable(
    caption = "Median and IQR of Leverage and Liquidity by Country (2018–2021)",
    digits = 2
  ) |>
  kableExtra::kable_styling(full_width = FALSE, font_size = 11)
```

Observation: Both countries maintained median liquidity near 1.8×, while median leverage hovered around 0.50–0.52.
Sweden shows slightly higher liquidity dispersion, suggesting a broader mix of firm sizes and capital structures.

### Question 3

AAA

### Question 4

AAA

## EDA

Here are the EDA processes and results for the sub-questions addressing the primary question "How did the pandemic affect corporate financials in Germany?"

### Question 1

AAA

### Question 2

#### EDA – Year-wise Comparison

```{r}
lev_liq_long <- lev_liq_summary |>
  pivot_longer(cols = c(median_debt_assets, median_liquidity),
               names_to = "indicator", values_to = "value")

ggplot(lev_liq_long,
       aes(x = acct_year, y = value, colour = country, group = country)) +
  geom_line(linewidth = 1.1) +
  geom_point(size = 2) +
  facet_wrap(~ indicator, scales = "free_y",
             labeller = as_labeller(c(
               median_debt_assets = "Median Debt-to-Assets",
               median_liquidity   = "Median Current Ratio"
             ))) +
  scale_y_continuous(labels = scales::number_format(accuracy = 0.01)) +
  scale_colour_brewer(palette = "Set1") +
  labs(
    title = "Leverage and Liquidity Trends (2018–2021)",
    subtitle = "Germany vs Sweden: Stability in liquidity, mild re-leveraging during pandemic",
    x = "Accounting Year", y = "Median Value", colour = "Country"
  ) +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(size = 11),
        panel.grid.minor = element_blank())
```

Interpretation: • Both economies experienced a slight leverage increase in 2020 (Germany +0.01, Sweden +0.02).
• Liquidity remained broadly constant across the pandemic, stabilising near 1.8–1.9×.
• The trends suggest that German firms relied more on credit support, while Swedish firms drew on retained liquidity, consistent with each country’s policy emphasis (Germany’s Kurzarbeit subsidies vs Sweden’s fiscal buffers).

#### EDA – Leverage–Solvency Relationship

```{r}
#| label: fig-leverage-solvency-facets
#| fig-cap: "Leverage–Solvency Relationship by Country and Year"
#| fig-width: 14
#| fig-height: 9

# build plotting data for the leverage–solvency faceted chart
lev_liq_plot <- lev_liq_data |>
  # keep sensible ranges and finite values
  filter(
    is.finite(debt_to_assets), is.finite(solvency_pct),
    between(debt_to_assets, 0, 1.5),   # 0–150%
    between(solvency_pct, -50, 100)    # -50% to 100%
  ) |>
  drop_na(country, acct_year)

ggplot(lev_liq_plot, aes(x = debt_to_assets, y = solvency_pct, colour = country)) +
  geom_point(alpha = 0.25, size = 1.2) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 1.1) +
  facet_grid(acct_year ~ country) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, 1.5)) +
  scale_y_continuous(limits = c(-50, 100)) +
  scale_colour_manual(values = c("Germany" = "#D55", "Sweden" = "#0072B2")) +
  labs(
    title = "Leverage–Solvency Relationship by Country and Year",
    subtitle = "Stronger negative slope in Germany, especially during 2020",
    x = "Debt-to-Assets (%)",
    y = "Solvency (%)",
    colour = "Country"
  ) +
  theme_minimal(base_size = 15) +
  theme(
    strip.text = element_text(face = "bold", size = 14),
    plot.title = element_text(face = "bold", size = 18, hjust = 0.5),
    plot.subtitle = element_text(size = 13, hjust = 0.5),
    legend.position = "none",  # each facet is already labelled
    axis.text = element_text(size = 12),
    panel.spacing.x = unit(1.2, "lines"),
    panel.spacing.y = unit(1.0, "lines"),
    panel.grid.minor = element_blank()
  )
```

Interpretation:

Across both Germany and Sweden, leverage shows a **strong negative relationship with solvency** from 2018 to 2021 — firms with higher debt-to-asset ratios consistently report lower solvency levels.
The **pattern remains stable** over time, confirming that higher leverage structurally reduces long-term financial stability in both economies.

However, the **slope is notably steeper in Germany**, especially in **2020**, indicating that increases in debt had a stronger adverse effect on solvency during the pandemic.
This likely reflects the **manufacturing-heavy structure** of German industries, which faced sharper revenue disruptions and slower balance sheet recovery.
In contrast, **Swedish firms**, with **more diversified sectors and stronger equity positions**, experienced a milder solvency decline, demonstrating greater resilience to leverage-induced stress.

EDA – Industry-level Leverage Distributions

```{r}
#| label: fig-leverage-industry-comparison
#| fig-cap: "Leverage Distribution by Industry and Year for Germany and Sweden (2018–2021)"
#| fig-width: 16
#| fig-height: 20

lev_liq_industry <- filtered_germany_sweden |>
  filter(acct_year %in% 2018:2021,
         total_assets_eur > 0,
         total_liabilities_eur >= 0,
         total_liabilities_eur / total_assets_eur <= 1.5) |>
  mutate(
    debt_to_assets = total_liabilities_eur / total_assets_eur
  ) |>
  drop_na(industry_group, country)

ggplot(lev_liq_industry,
       aes(x = debt_to_assets,
           y = reorder(industry_group, debt_to_assets),
           fill = country)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7, width = 0.6) +
  facet_wrap(~ acct_year, ncol = 2, scales = "free_x") +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1),
                     limits = c(0, 1.5)) +
  scale_fill_manual(
    values = c("Germany" = "#D55", "Sweden" = "#0072B2"),
    name = "Country"
  ) +
  labs(
    title = "Leverage Distribution by Industry and Year (2018–2021)",
    subtitle = "Comparison between Germany and Sweden across industry groups and years",
    x = "Debt-to-Assets Ratio",
    y = NULL
  ) +
  theme_minimal(base_size = 16) +  # increased from 13 → 16
  theme(
    strip.text = element_text(face = "bold", size = 15),
    plot.title = element_text(face = "bold", size = 20, hjust = 0.5),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    legend.position = "top",
    legend.text = element_text(size = 13),
    legend.title = element_text(size = 14),
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    panel.grid.minor = element_blank(),
    panel.spacing.y = unit(1.2, "lines")
  )
```

Interpretation:

The boxplots reveal that leverage remained broadly stable across both countries from 2018 to 2021, with **capital-intensive sectors**-notably *Manufacturing*, *Construction*, and *Transport & Storage*-maintaining higher debt-to-asset ratios than service-oriented industries such as *Information & Communication* or *Professional & Technical Services*.

The median levels changed little across years, though a modest upward shift appears in 2020, consistent with increased borrowing during the early pandemic period.

Overall, **Germany and Sweden display similar leverage structures**, suggesting comparable financial resilience and corporate funding patterns.

When leverage patterns are compared with liquidity (current ratio) trends, a complementary story emerges.

Firms in both countries **maintained adequate short-term liquidity**, with median current ratios remaining around 1.8–2.0 throughout 2018–2021.

This indicates that, despite slightly greater reliance on debt during 2020, most firms preserved the capacity to meet near-term obligations.

The combination of stable leverage and steady liquidity implies that corporate balance sheets were **buffered by prudent debt management and strong internal financing**, helping both economies avoid a widespread solvency crisis during the COVID-19 shock.

```{r}
#| warning: false
# Sensitivity check: winsorise extreme ratios
lev_liq_summary_w <- lev_liq_data |>
  mutate(across(c(debt_to_assets, current_ratio),
                ~ scales::squish(., c(quantile(., 0.01, na.rm=TRUE),
                                      quantile(., 0.99, na.rm=TRUE))))) |>
  group_by(country, acct_year) |>
  summarise(across(c(debt_to_assets, current_ratio), median, na.rm=TRUE), .groups="drop")
knitr::kable(head(lev_liq_summary_w), caption="Winsorised medians confirm robustness.")
```

Summary interpretation: • Leverage: Both countries show small increases during 2020, reverting by 2021.
• Liquidity: Stable across years and nations, implying effective cash-flow management.
• Cross-country pattern: Sweden’s firms appear slightly more liquid and less leveraged, signalling greater resilience to credit-market stress.
• Conclusion: Despite pandemic disruptions, neither economy faced systemic liquidity deterioration; Germany’s higher industrial exposure drove modest leverage stress, while Sweden’s diversified base cushioned impact.

### Question 3

AAA

### Question 4

AAA

## Conclusions

AAA

## References

The materials used for this report are:

-   Bureau van Dijk.
    (2025).
    OSIRIS \[Data set: Germany subset\].
    Bureau van Dijk – A Moody’s Analytics Company.
    https://www.bvdinfo.com/en-gb/our-products/data/international/osiris

-   **Tidyverse:** Wickham H, Averick M, Bryan J, Chang W, McGowan LD, François R, Grolemund G, Hayes A, Henry L, Hester J, Kuhn M, Pedersen TL, Miller E, Bache SM, Müller K, Ooms J, Robinson D, Seidel DP, Spinu V, Takahashi K, Vaughan D, Wilke C, Woo K, Yutani H (2019).
    “Welcome to the tidyverse.” *Journal of Open Source Software*, *4*(43), 1686.
    doi:10.21105/joss.01686 <https://doi.org/10.21105/joss.01686>.

-   **Conflicted:** Wickham H (2023).
    *conflicted: An Alternative Conflict Resolution Strategy*.
    doi:10.32614/CRAN.package.conflicted <https://doi.org/10.32614/CRAN.package.conflicted>, R package version 1.2.0, <https://CRAN.R-project.org/package=conflicted>.

-   **Dplyr:** Wickham H, François R, Henry L, Müller K, Vaughan D (2023).
    *dplyr: A Grammar of Data Manipulation*.
    doi:10.32614/CRAN.package.dplyr <https://doi.org/10.32614/CRAN.package.dplyr>, R package version 1.1.4, <https://CRAN.R-project.org/package=dplyr>.

-   **Stringr:** Wickham H (2025).
    *stringr: Simple, Consistent Wrappers for Common String Operations*.
    doi:10.32614/CRAN.package.stringr <https://doi.org/10.32614/CRAN.package.stringr>, R package version 1.5.2, <https://CRAN.R-project.org/package=stringr>.

-   **Janitor:** Firke S (2024).
    *janitor: Simple Tools for Examining and Cleaning Dirty Data*.
    doi:10.32614/CRAN.package.janitor <https://doi.org/10.32614/CRAN.package.janitor>, R package version 2.2.1, <https://CRAN.R-project.org/package=janitor>.

-   **Skimr:** Waring E, Quinn M, McNamara A, Arino de la Rubia E, Zhu H, Ellis S (2025).
    *skimr: Compact and Flexible Summaries of Data*.
    doi:10.32614/CRAN.package.skimr <https://doi.org/10.32614/CRAN.package.skimr>, R package version 2.2.1, <https://CRAN.R-project.org/package=skimr>.

-   **Tsibble:** Wang, E, D Cook, and RJ Hyndman (2020).
    A new tidy data structure to support exploration and modeling of temporal data, Journal of Computational and Graphical Statistics, 29:3, 466-478, doi:10.1080/10618600.2019.1695624.

-   **Knitr:** Xie Y (2025).
    *knitr: A General-Purpose Package for Dynamic Report Generation in R*.
    R package version 1.50, <https://yihui.org/knitr/>.

    Yihui Xie (2015) Dynamic Documents with R and knitr.
    2nd edition.
    Chapman and Hall/CRC.
    ISBN 978-1498716963

    Yihui Xie (2014) knitr: A Comprehensive Tool for Reproducible Research in R.
    In Victoria Stodden, Friedrich Leisch and Roger D. Peng, editors, Implementing Reproducible Computational Research.
    Chapman and Hall/CRC.
    ISBN 978-1466561595

-   **Naniar:** Tierney N, Cook D (2023).
    “Expanding Tidy Data Principles to Facilitate Missing Data Exploration, Visualization and Assessment of Imputations.” *Journal of Statistical Software*, *105*(7), 1-31.
    doi:10.18637/jss.v105.i07 <https://doi.org/10.18637/jss.v105.i07>.

-   **Patchwork:** Pedersen T (2025).
    *patchwork: The Composer of Plots*.
    doi:10.32614/CRAN.package.patchwork <https://doi.org/10.32614/CRAN.package.patchwork>, R package version 1.3.2, <https://CRAN.R-project.org/package=patchwork>.

-   **Plotly:** C. Sievert.
    Interactive Web-Based Data Visualization with R, plotly, and shiny.
    Chapman and Hall/CRC Florida, 2020.

-   **tsibbletalk:** Wang E, Cook D (2020).
    *tsibbletalk: Interactive Graphics for Tsibble Objects*.
    doi:10.32614/CRAN.package.tsibbletalk <https://doi.org/10.32614/CRAN.package.tsibbletalk>, R package version 0.1.0, <https://CRAN.R-project.org/package=tsibbletalk>.

-   **Feasts:** O'Hara-Wild M, Hyndman R, Wang E (2025).
    *feasts: Feature Extraction and Statistics for Time Series*.
    doi:10.32614/CRAN.package.feasts <https://doi.org/10.32614/CRAN.package.feasts>, R package version 0.4.2, <https://CRAN.R-project.org/package=feasts>.

-   **Broom:** Robinson D, Hayes A, Couch S (2025).
    *broom: Convert Statistical Objects into Tidy Tibbles*.
    doi:10.32614/CRAN.package.broom <https://doi.org/10.32614/CRAN.package.broom>, R package version 1.0.9, <https://CRAN.R-project.org/package=broom>.

-   **Purrr:** Wickham H, Henry L (2025).
    *purrr: Functional Programming Tools*.
    doi:10.32614/CRAN.package.purrr <https://doi.org/10.32614/CRAN.package.purrr>, R package version 1.1.0, <https://CRAN.R-project.org/package=purrr>.

-   **Cassowaryr:** L. Wilkinson, A. Anand and R.
    Grossman, Graph-theoretic scagnostics, IEEE Symposium on Information Visualization, 2005.
    INFOVIS 2005., 2005, pp. 157-164, doi: 10.1109/INFVIS.2005.1532142.

    L. Wilkinson and G.
    Wills (2008) Scagnostics Distributions, Journal of Computational and Graphical Statistics, 17(2), pp 473-491, doi:10.1198/106186008X320465

    K. Grimm, Kennzahlenbasierte Grafikauswahl, doctoral thesis, Universitat Augsburg, 2016.

    H. Mason, Lee, S., Laa, U.
    and Cook, D.
    (2022) cassowaryr: Compute Scagnostics on Pairs of Numeric Variables in a Data Set.
    R package version 2.0.0.
    https://CRAN.R-project.org/package=cassowary

-   **Gganimate:** Pedersen T, Robinson D (2025).
    *gganimate: A Grammar of Animated Graphics*.
    doi:10.32614/CRAN.package.gganimate <https://doi.org/10.32614/CRAN.package.gganimate>, R package version 1.0.11, <https://CRAN.R-project.org/package=gganimate>.

-   **Gifski:** Ooms J, Kornel Lesiński, Authors of the dependency Rust crates (2025).
    *gifski: Highest Quality GIF Encoder*.
    doi:10.32614/CRAN.package.gifski <https://doi.org/10.32614/CRAN.package.gifski>, R package version 1.32.0-2, <https://CRAN.R-project.org/package=gifski>.

-   **KableExtra:** Zhu H (2024).
    *kableExtra: Construct Complex Table with 'kable' and Pipe Syntax*.
    doi:10.32614/CRAN.package.kableExtra <https://doi.org/10.32614/CRAN.package.kableExtra>, R package version 1.4.0, <https://CRAN.R-project.org/package=kableExtra>.

-   **Ggplot2:** H. Wickham.
    ggplot2: Elegant Graphics for Data Analysis.
    Springer-Verlag New York, 2016.

-   **Scales:** Wickham H, Pedersen T, Seidel D (2025).
    *scales: Scale Functions for Visualization*.
    doi:10.32614/CRAN.package.scales <https://doi.org/10.32614/CRAN.package.scales>, R package version 1.4.0, <https://CRAN.R-project.org/package=scales>.

-   **Tidyr:** Wickham H, Vaughan D, Girlich M (2024).
    *tidyr: Tidy Messy Data*.
    doi:10.32614/CRAN.package.tidyr <https://doi.org/10.32614/CRAN.package.tidyr>, R package version 1.3.1, <https://CRAN.R-project.org/package=tidyr>.

-   **Ggrepel:** Slowikowski K (2024).
    *ggrepel: Automatically Position Non-Overlapping Text Labels with 'ggplot2'*.
    doi:10.32614/CRAN.package.ggrepel <https://doi.org/10.32614/CRAN.package.ggrepel>, R package version 0.9.6, <https://CRAN.R-project.org/package=ggrepel>.

-   **ViridisLite:** Simon Garnier, Noam Ross, Robert Rudis, Antônio P. Camargo, Marco Sciaini, and Cédric Scherer (2023).
    viridis(Lite) - Colorblind-Friendly Color Maps for R.
    viridisLite package version 0.4.2.

-   The Associated Press.
    (2020).
    Europe braces for next wave of coronavirus pandemic in Berlin.
    AP News.
    https://apnews.com/article/coronavirus-pandemic-health-europe-epidemics-berlin-b61de99739774c1f52b4ba6860054d6d

## AI Acknowledgement

The links to my use of ChatGPT and other AI for help on this assignment are:

-   <https://chatgpt.com/share/68f563ee-9040-8010-ab1c-4969cfb31ee0> Used to help know when the pandemic happened in Germany, understand the accounting terms related to profitability, how to create a violin chart and jitter, how to make a dumbbell chart for comparison, and perform imputation on the data.

-   <https://chatgpt.com/share/68f4378d-bb24-8003-8471-ed2147a8fab3> Used for formatting assistance, debugging R code, and improving interpretation flow.

-   <https://chatgpt.com/share/68f3046d-1ad8-800d-bc11-1b005a9e600f> helped to implement animated graphics and optimize content.

-   <https://chatgpt.com/share/68f304c8-a088-800d-be54-939dfcd7e278> helped to evaluate whether the questions are answered adequately and where to make revisions.

-   <https://chatgpt.com/share/68f305ad-f15c-800d-b724-fc917c1a93ef> helped to deal with the problem that package ‘tignostics’ is not available for this version of R and how to use scagnostics.

-   <https://chatgpt.com/share/68f571d7-0810-8011-960e-b969c4d9a752> helped us format the explanation in a better and concise manner.
