---
title: "ETC5521 Diving Deeper into Data Exploration: Project Part 1"
author: "P1-wallaby"
date: "Oct 6, 2025"
quarto-required: ">=1.8.0"
format:
    html: 
        output-file: report.html
        css: "assignment.css"
        embed-resources: true
        code-fold: true
        code-tools: true
        code-summary: "Show code"
        code-line-numbers: true
        toc: true
        toc-depth: 3
        toc-location: left
        theme: cerulean
editor: 
  markdown: 
    wrap: sentence
---
:::: {.panel-tabset}

```{r}
#| echo: false
# Set up chunk for all slides
knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 4,
  fig.align = "center",
  out.width = "100%",
  code.line.numbers = FALSE,
  fig.retina = 4,
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = FALSE,
  dev.args = list(pointsize = 11)
)
```

```{r}
#| echo: false
# Load libraries
library(tidyverse)
library(conflicted)

conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::select)
conflicts_prefer(dplyr::slice)

# Plot options and themes
options(
  digits = 2,
  width = 60,
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis",
  ggplot2.discrete.colour = c("#D55E00", "#0072B2", "#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442"),
  ggplot2.discrete.fill = c("#D55E00", "#0072B2", "#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442")
)

theme_set(theme_bw(base_size = 14) +
   theme(
     aspect.ratio = 1,
     plot.background = element_rect(fill = 'transparent', colour = NA),
     plot.title.position = "plot",
     plot.title = element_text(size = 24),
     panel.background = element_rect(fill = 'transparent', colour = NA),
     legend.background = element_rect(fill = 'transparent', colour = NA),
     legend.key = element_rect(fill = 'transparent', colour = NA)
   )
)
```

This is the detail library that we used for this project.

```{r load-library}
library(dplyr)
library(stringr)
library(janitor)
library(skimr)
library(tsibble)
library(knitr)
library(naniar)
library(patchwork)
library(plotly)
library(tsibbletalk)
library(feasts)
library(broom)
library(purrr)
library(cassowaryr)
library(gganimate)
library(gifski)
library(kableExtra)
library(ggplot2)
library(scales)
library(tidyr)
library(ggrepel)
library(viridisLite)
```

## Introduction

The COVID-19 pandemic caused significant disruptions across global markets, testing the financial resilience of firms in every sector.
In Germany, the first confirmed case was recorded on 27 January 2020.
Nationwide contact restrictions began on 22 March 2020.
Most nationwide measures were lifted between 20 March and 2 April 2022, and the final federal provisions expired on 7 April 2023.

This project examines how German companies responded to these shocks during 2020 – 2021, using financial data from the Bureau van Dijk OSIRIS Database.
A reliable global source of harmonised corporate information.
The dataset available for this study covers 2018 - 2021.
The period from 2018–2019 provides the pre-pandemic baseline, while 2020 – 2021 captures the pandemic period.
Data for 2022 are not available in this extract.

Our analysis focuses on how industries differed in their financial outcomes during this period, investigating changes in profitability, leverage, liquidity, and firm size.
By identifying which sectors showed unexpected resilience or vulnerability, we aim to understand how structural and strategic differences shaped recovery from the pandemic’s economic impacts.
All analytical procedures follow the ETC5521 workflow, ensuring reproducibility, transparency, and alignment with the course framework for exploratory data analysis.

## Data description

The dataset used in this analysis originates from **the Bureau van Dijk OSIRIS database**, a comprehensive source of harmonised financial information for publicly listed and major private firms worldwide.
For this study, we extracted firm-level data for **German companies** over the **2018–2021 accounting years**, a period that captures both pre-pandemic and pandemic-era financial performance.
Each record represents a firm’s reported financials for a specific accounting year, enabling longitudinal exploration of trends in profitability, liquidity, leverage, and resilience.
These are the code that we used to load the data and combine them:

Loading the data.

```{r read-data}
#define the folder
data_directory <- "osiris"

#filter only germany
germany_file_paths <- list.files(
  data_directory,
  pattern = "^osiris_?Germany_(2018|2019|2020|2021)\\.rda$",
  full.names = TRUE,
  ignore.case = TRUE
)

#name the list by file stems
germany_file_stems <- tools::file_path_sans_ext(basename(germany_file_paths))
osiris_germany <- setNames(vector("list", length(germany_file_paths)), germany_file_stems)

#load each file
for (file_index in seq_along(germany_file_paths)) {
  temporary_environment <- new.env(parent = emptyenv())
  load(germany_file_paths[file_index], envir = temporary_environment)
  loaded_objects <- as.list(temporary_environment)

  #If the file holds one object, store that object; otherwise store the sub-list
  osiris_germany[[file_index]] <- if (length(loaded_objects) == 1) loaded_objects[[1]] else loaded_objects
}
```

Combining the data.

```{r read-data-2}
#build a single data frame with a 'year' column taken from the filename
combined_germany <- {
  germany_parts <- vector("list", length(osiris_germany))
  names(germany_parts) <- names(osiris_germany)

  for (file_stem in names(osiris_germany)) {
    data_frame_from_file <- osiris_germany[[file_stem]]
    extracted_year <- as.integer(str_extract(file_stem, "\\d{4}"))

    germany_parts[[file_stem]] <- mutate(data_frame_from_file, year = extracted_year)
  }

  #bind and sort by year (optional)
  bind_rows(germany_parts) |>
    arrange(year)
}
```

After the data is combined, we standardise the raw file into analysis-ready fields.
It normalises column names, maps long vendor headers to clear variables by topic, and creates a fiscal-year label from the year-end date.
Finally, it keeps only the identifiers, industry codes, and financial variables needed for the analysis.

```{r data-cleaning}
filtered_germany <- combined_germany |>
  clean_names() |>
  
  rename(
    # IDs / scope (used in all questions)
    company_name       = company_name_name,
    company_id         = name_id,
    acct_year          = year,
    country            = country_country,
    city               = city_city_city,
    consolidation_code = consolidation_code_consol_code,
    status             = status_status,
    
    # Industry codes (Q1, Q2, Q4)
    nace_code = nace_rev_1_core_code_cnacecd,
    icb_code  = industrial_classification_benchmark_icb,
    sic_code  = us_sic_core_code_csicuscde,
    
    # Scale / balance sheet (Q2, Q3)
    total_assets_eur      = total_assets_data13077,
    total_liabilities_eur = total_liabilities_and_debt_data14022,
    total_equity_eur      = total_shareholders_equity_data14041,
    
    # Revenue & profit (Q1, Q3)
    total_revenue_eur = total_revenues_data13004,
    net_sales_eur     = net_sales_data13002,
    gross_sales_eur   = gross_sales_data13000,
    net_income_eur    = net_income_starting_line_data15500,
    
    # Profitability ratios (Q1, Q4)
    ebit_margin_pct   = ebit_margin_percent_data31055,
    ebitda_margin_pct = ebitda_margin_percent_data31060,
    roa_pct           = return_on_total_assets_percent_data31015,
    roe_pct           = roe_percent_data31065,
    
    # Efficiency
    net_assets_turnover = net_assets_turnover_data31225,
    stock_turnover      = stock_turnover_data31220,
    
    # Leverage, solvency, and liquidity (Q2)
    solvency_pct        = solvency_ratio_percent_data31310,
    gearing_pct         = gearing_percent_data31315,
    current_ratio       = current_ratio_data31105,
    liquidity_ratio     = liquidity_ratio_data31110,
    shareholders_liq_pct= shareholders_liquidity_ratio_data31305,
    interest_cover      = interest_cover_data31115
  ) |>
  
  # Rename fiscal year
  mutate(
    fy_end_date = lubridate::ymd(as.character(company_fiscal_year_end_date_closdate)),
    fy_year     = year(fy_end_date)
  ) |>
  
  # Select only relevant columns
  select(
    company_id, company_name, fy_end_date, fy_year, acct_year, status, country, city, consolidation_code,
    nace_code, icb_code, sic_code,
    total_assets_eur, total_liabilities_eur, total_equity_eur,
    total_revenue_eur, net_sales_eur, gross_sales_eur, net_income_eur,
    ebit_margin_pct, ebitda_margin_pct, roa_pct, roe_pct,
    net_assets_turnover, stock_turnover,
    solvency_pct, gearing_pct, current_ratio, liquidity_ratio,
    shareholders_liq_pct, interest_cover
  )

```

This is the table describing the variables we used in this project:

```{r table-description, echo=FALSE}
data_dictionary <- tribble(
  ~`Column Name`,           ~Description,                                                                    ~`Data Type`, ~`Original Name`,
  "company_id",             "Unique company identifier (Bureau van Dijk ID).",                               "character", "BvD ID Number (os_id_number)",
  "company_name",           "Registered company name (label).",                                               "character", "Company Name (name)",
  "fy_end_date",            "Company fiscal year-end date for the account.",                                  "date",      "Company Fiscal Year End Date (closdate)",
  "fy_year",                "Year extracted from fiscal year-end date.",                                      "integer",   "derived from Company Fiscal Year End Date (closdate)",
  "acct_year",              "Reporting year label carried from file name/field.",                             "integer",   "year",
  "country",                "Country of head office.",                                                         "character", "Country (country)",
  "city",                   "City of head office.",                                                            "character", "CITY - City (city)",
  "consolidation_code",     "Consolidation scope of the accounts (e.g., C1/C2/U1/U2).",                        "character", "Consolidation Code (consol_code)",
  "nace_code",              "NACE core code (EU industry classification).",                                    "character", "NACE Rev 1, Core Code (cnacecd)",
  "icb_code",               "ICB industry classification code.",                                               "character", "Industrial Classification Benchmark (icb)",
  "sic_code",               "US SIC core code.",                                                               "character", "US SIC, Core Code (csicuscde)",
  "total_assets_eur",       "Total assets (EUR).",                                                              "numeric",   "Total Assets (data13077)",
  "total_liabilities_eur",  "Total liabilities and debt (EUR).",                                               "numeric",   "Total Liabilities and Debt (data14022)",
  "total_equity_eur",       "Total shareholders’ equity (EUR).",                                               "numeric",   "Total Shareholders Equity (data14041)",
  "total_revenue_eur",      "Total revenues / operating revenue (EUR).",                                       "numeric",   "Total revenues (data13004)",
  "net_sales_eur",          "Net sales (EUR).",                                                                 "numeric",   "Net sales (data13002)",
  "gross_sales_eur",        "Gross sales (EUR).",                                                               "numeric",   "Gross sales (data13000)",
  "net_income_eur",         "Net income (EUR).",                                                                "numeric",   "Net Income / Starting Line (data15500)",
  "ebit_margin_pct",        "EBIT margin (% of sales/revenue).",                                               "numeric",   "EBIT Margin (%) (data31055)",
  "ebitda_margin_pct",      "EBITDA margin (% of sales/revenue).",                                             "numeric",   "EBITDA Margin (%) (data31060)",
  "roa_pct",                "Return on total assets (%).",                                                     "numeric",   "Return on Total Assets (%) (data31015)",
  "roe_pct",                "Return on equity (%).",                                                           "numeric",   "ROE (%) (data31065)",
  "net_assets_turnover",    "Net assets turnover (times).",                                                    "numeric",   "Net Assets Turnover (data31225)",
  "stock_turnover",         "Stock / inventory turnover (times).",                                             "numeric",   "Stock Turnover (data31220)",
  "solvency_pct",           "Solvency ratio (%).",                                                             "numeric",   "Solvency ratio (%) (data31310)",
  "gearing_pct",            "Gearing ratio (%).",                                                              "numeric",   "Gearing (%) (data31315)",
  "current_ratio",          "Current ratio (times).",                                                          "numeric",   "Current ratio (data31105)",
  "liquidity_ratio",        "Liquidity / quick ratio (times).",                                                "numeric",   "Liquidity ratio (data31110)",
  "shareholders_liq_pct",   "Shareholders’ liquidity ratio (%).",                                              "numeric",   "Shareholders Liquidity ratio (data31305)",
  "interest_cover",         "Interest cover (times).",                                                         "numeric",   "Interest Cover (data31115)"
)

#show table
data_dictionary |>
  kable(align = "l", booktabs = TRUE) |>
  kable_styling(full_width = FALSE, font_size = 11)
```

The selected variables encompass four key dimensions of corporate financial health:

-   **Identification and Location** - company_id, company_name, country, and city provide firm-level identifiers and geographical context.

<!-- -->

-   **Balance Sheet Structure** - total_assets_eur, total_liabilities_eur, and total_equity_eur measure firm size and capital composition, reported in euros.

<!-- -->

-   **Income and Profitability** - total_revenue_eur, net_sales_eur, and net_income_eur capture income statement outcomes, while ebit_margin_pct, ebitda_margin_pct, roa_pct, and roe_pct reflect operating performance and returns to shareholders.

<!-- -->

-   **Solvency and Liquidity** - solvency_pct, gearing_pct, current_ratio, liquidity_ratio, shareholders_liq_pct, and interest_cover assess each firm’s financial stability, leverage, and capacity to meet short-term obligations.

-   **Industry Classification** - industry_nace, industry_icb, and industry_group link firms to their respective economic sectors according to European (NACE) and global (ICB) coding systems.

It is important to note that the **accounting year** variable in **OSIRIS** reflects **two-year reporting periods**.
For example, the 2018 accounting year represents financial results spanning 2017–2018, 2019 corresponds to 2018–2019, 2020 covers 2019–2020, and 2021 includes 2020–2021.
Throughout this report, each accounting year is interpreted as representing the **endpoint of its reporting cycle**.
Consequently, “2021” is treated as capturing firm performance during the height of the **COVID-19 pandemic**, while 2018–2019 represent pre-pandemic benchmarks.

This dataset structure enables a robust exploration of temporal changes in corporate performance and the role of firm characteristics , particularly **firm size** , in shaping **financial resilience** throughout the pandemic period.

### Research Question

Here are the research questions (sub-questions) that guide us in analysing and answering the primary question: How did the pandemic affect corporate financials in Germany?

**1. How did profitability change across industries?**

This question assesses how profitability variables, such as ROA, EBITDA, or margins shifted across industries over time, highlighting which sectors improved, remained steady, or declined during the pre-pandemic period (2018 – 2019) and the pandemic (2020 – 2021) in Germany.
These cross-industry differences reveal relative exposure, resilience, and recovery pathways during the pandemic compared to the pre-pandemic period.
Understanding the direction and scale of profitability changes directly addresses the main question of how the pandemic affected corporate finances.

**2. How did leverage (debt-to-assets) and liquidity (current ratio) evolve during pandemic?**

This question examines how German firms managed their financial structure and short-term liquidity during the COVID-19 period.
By tracking leverage and liquidity trends, we can assess whether firms relied more on debt financing or preserved liquidity to withstand operational shocks.
Understanding these shifts helps explain how the pandemic influenced firms’ balance sheet resilience and capacity to absorb financial stress.

**3. Which German industries showed unexpected financial resilience or vulnerability from 2020 to 2022, and how were these outcomes shaped by profitability, leverage, liquidity, and firm size?**

During 2020–2022, industries such as healthcare, information technology, and utilities demonstrated notable financial resilience, supported by strong liquidity and stable profitability.
In contrast, sectors like hospitality, transportation, and manufacturing showed greater vulnerability due to higher leverage and lower profit margins.
Overall, firm size and liquidity emerged as key buffers against financial stress, while high debt levels amplified vulnerability across more capital-intensive industries.

**4. How did the fundamental relationship between corporate liquidity and profitability evolve and fracture within industries during the pandemic?**

The COVID-19 pandemic (2020–2022) placed severe pressure on global corporate financial structures.
Firms faced the challenge of maintaining short-term liquidity without compromising profitability.
Financial theory suggests that adequate liquidity supports stable profits by ensuring operational continuity.
However, this relationship may have fractured across industries amid uneven shocks and policy responses.
This question examines how liquidity (Current Ratio) and profitability (ROA) evolved among German firms during the pandemic.
It helps reveal whether the pandemic reshaped the traditional balance between liquidity and profitability in corporate financial resilience.

## IDA

Here is the IDA process for this project:

First, we develop this code to prepare the dataset for analysis by converting numeric columns, standardising text fields, and reporting missing values overall and by year.
It also performs basic sanity checks for duplicate firm-years, impossible values, and consolidation codes.

```{r table-checking}

num_cols <- c(
  "total_assets_eur","total_liabilities_eur","total_equity_eur",
  "total_revenue_eur","net_sales_eur","gross_sales_eur","net_income_eur",
  "ebit_margin_pct","ebitda_margin_pct","roa_pct","roe_pct",
  "net_assets_turnover","stock_turnover",
  "solvency_pct","gearing_pct","current_ratio","liquidity_ratio",
  "shareholders_liq_pct","interest_cover"
)

#symbol formatting/coerce numerics
filtered_germany <- filtered_germany |>
  mutate(
    across(all_of(num_cols) & where(is.character), readr::parse_number),
    country = stringr::str_to_title(country),
    city    = stringr::str_to_title(city)
  )

#missingness report check (overall and by year)
missing_overall <- filtered_germany |>
  summarise(across(everything(), ~ mean(is.na(.)))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "missing_rate")

missing_by_year <- filtered_germany |>
  group_by(acct_year) |>
  summarise(across(everything(), ~ mean(is.na(.)))) |>
  pivot_longer(-acct_year, names_to = "variable", values_to = "missing_rate")

#sanity checks (optional but recommended)
dup_firm_year <- filtered_germany |>
  count(company_id, fy_year, acct_year) |>
  filter(n > 1)

impossible_assets <- filtered_germany |>
  filter(!is.na(total_assets_eur) & total_assets_eur < 0)

consol_levels <- filtered_germany |>
  count(consolidation_code, sort = TRUE)

```

Then, we create this code to extract the 2-digit NACE code, categorises companies into broad industry groups using NACE and ICB, and creates a unified 'industry_group' factor with a fixed order for plotting purposes.
It also provides a small sample of unmapped records for verification.
After checking, there are no duplicates in the data using company_id, fy_year, and acct_year as the primary key.

```{r NACE-column}
#clean 2-digit NACE as integer
filtered_germany <- filtered_germany |>
  mutate(
    nace_num2 = suppressWarnings(as.integer(str_sub(readr::parse_number(as.character(nace_code)), 1, 2)))
  )

#map NACE divisions to broad section names
industry_from_nace <- function(x){
  case_when(
    x %in% 1:3 ~ "Agriculture, Forestry & Fishing",
    x %in% 5:9 ~ "Mining & Quarrying",
    x %in% 10:33 | x %in% 15:37 ~ "Manufacturing",
    x %in% 35 ~ "Electricity, Gas, Steam",
    x %in% 36:39 ~ "Water Supply & Waste",
    x %in% 41:43 ~ "Construction",
    x %in% 45:47 ~ "Wholesale & Retail Trade",
    x %in% 49:53 ~ "Transport & Storage",
    x %in% 55:56 ~ "Accommodation & Food",
    x %in% 58:63 ~ "Information & Communication",
    x %in% 64:66 ~ "Financial & Insurance",
    x %in% 68 ~ "Real Estate",
    x %in% 69:75 ~ "Professional, Scientific & Technical",
    x %in% 77:82 ~ "Administrative & Support",
    x %in% 84 ~ "Public Administration",
    x %in% 85 ~ "Education",
    x %in% 86:88 ~ "Human Health & Social Work",
    x %in% 90:93 ~ "Arts, Entertainment & Recreation",
    x %in% 94:96 ~ "Other Service Activities",
    x %in% 97:98 ~ "Household Activities",
    x %in% 99 ~ "Extraterritorial Organizations",
    TRUE ~ NA_character_
  )
}

#ICB top-level sector labels
icb_map <- c(
  "0001" = "Oil & Gas",
  "1000" = "Basic Materials",
  "2000" = "Industrials",
  "3000" = "Consumer Goods",
  "4000" = "Health Care",
  "5000" = "Consumer Services",
  "6000" = "Telecommunications",
  "7000" = "Utilities",
  "8000" = "Financials",
  "9000" = "Technology"
)

#normalize ICB to a 4-digit “bucket” (example: 2573 to 2000)
normalize_icb <- function(x){
  x_chr <- str_extract(as.character(x), "\\d+")
  ifelse(is.na(x_chr), NA_character_,
         sprintf("%04d", as.integer(floor(as.numeric(x_chr) / 1000) * 1000)))
}

filtered_germany <- filtered_germany |>
  mutate(
    industry_nace = industry_from_nace(nace_num2),
    icb_bucket    = normalize_icb(icb_code),
    industry_icb  = icb_map[icb_bucket],
    industry_group = coalesce(industry_nace, industry_icb, "Other / Unmapped")
  )

#check
unmapped_sample <- filtered_germany |>
  filter(industry_group == "Other / Unmapped") |>
  select(company_id, fy_year, acct_year, nace_code, icb_code) |>
  head(15)

#factor order for plots
ordered_levels <- c(
  "Agriculture, Forestry & Fishing","Mining & Quarrying","Manufacturing",
  "Electricity, Gas, Steam","Water Supply & Waste","Construction",
  "Wholesale & Retail Trade","Transport & Storage","Accommodation & Food",
  "Information & Communication","Financial & Insurance","Real Estate",
  "Professional, Scientific & Technical","Administrative & Support",
  "Public Administration","Education","Human Health & Social Work",
  "Arts, Entertainment & Recreation","Other Service Activities",
  "Household Activities","Extraterritorial Organizations",
  "Oil & Gas","Basic Materials","Industrials","Consumer Goods","Health Care",
  "Consumer Services","Telecommunications","Utilities","Financials","Technology",
  "Other / Unmapped"
)
filtered_germany <- filtered_germany |>
  mutate(industry_group = factor(industry_group, levels = ordered_levels))
```

After that, we use this code to reshape numeric variables into a long format for Initial Data Analysis, separate scale variables from ratios, and assign clear labels for charts.
The result is a tidy table ready for profiling and visualisation.

```{r prepare-IDA}
#using long data
ida_germany <- filtered_germany |>
  select(acct_year, all_of(num_cols)) |>
  pivot_longer(-acct_year, names_to = "variable_orig", values_to = "value") |>
  drop_na()

#split between scale and ratio
scale_vars <- c("total_assets_eur","total_liabilities_eur","total_equity_eur",
                "total_revenue_eur","net_sales_eur","gross_sales_eur","net_income_eur")
ratio_vars <- setdiff(unique(ida_germany$variable_orig), scale_vars)

#label mapping for the charts
nice <- c(
  total_assets_eur   = "Total assets (EUR)",
  total_liabilities_eur="Total liabilities (EUR)",
  total_equity_eur   = "Equity (EUR)",
  total_revenue_eur  = "Total revenue (EUR)",
  net_sales_eur      = "Net sales (EUR)",
  gross_sales_eur    = "Gross sales (EUR)",
  net_income_eur     = "Net income (EUR)",
  ebit_margin_pct    = "EBIT margin (%)",
  ebitda_margin_pct  = "EBITDA margin (%)",
  roa_pct            = "ROA (%)",
  roe_pct            = "ROE (%)",
  net_assets_turnover= "Net assets turnover (x)",
  stock_turnover     = "Stock turnover (x)",
  solvency_pct       = "Solvency (%)",
  gearing_pct        = "Gearing (%)",
  current_ratio      = "Current ratio (x)",
  liquidity_ratio    = "Liquidity ratio (x)",
  shareholders_liq_pct = "Shareholders’ liquidity (%)",
  interest_cover     = "Interest cover (x)"
)

pretty_var <- function(x) ifelse(x %in% names(nice), nice[x], x)

ida_germany <- ida_germany |>
  mutate(variable = pretty_var(variable_orig),
         variable = factor(variable, levels = unique(variable)))

```

Then, we create this code to visualise the distribution of scale variables by year using violin and jitter plots on a log scale, with medians highlighted.
It allows quick checks of spread, skewness, and outliers over time for monetary measures.

```{r IDA-scale}
#chart for scale variables (log axis)
p_scale <- ida_germany |>
  filter(variable_orig %in% scale_vars) |>
  mutate(y = if_else(value > 0, value, NA_real_)) |>
  ggplot(aes(x = factor(acct_year), y = y)) +
  geom_violin(trim = FALSE, fill = NA, linewidth = 0.4, na.rm = TRUE) +
  geom_jitter(width = 0.12, height = 0, alpha = 0.12, size = 0.6, na.rm = TRUE) +
  stat_summary(fun = median, geom = "point", shape = 95, size = 6, na.rm = TRUE) +
  scale_y_log10(labels = scales::label_number(scale_cut = scales::cut_short_scale())) +
  facet_wrap(~ variable, scales = "free_y", ncol = 3) +
  labs(x = "Year", y = NULL, title = "Violin + Jitter by Year (log scale for Monetary Variables)") +
  theme_minimal(base_size = 11) +
  theme(axis.text.x = element_text(size = 10),
        strip.text  = element_text(face = "bold"),
        panel.grid.minor = element_blank())

p_scale
```

According to the chart, the monetary amounts (EUR) variables are heavily right-skewed: most firms are small to mid-sized, with a long tail from a few very large firms (hence the log scale in figures).
From 2018 to 2021, medians are relatively stable, with a slight increase in liabilities relative to assets and sales.
Distributions widen in 2020 and 2021, showing more variability in assets, revenues, and especially net income.
Net income shows the most significant relative spread, indicating higher volatility in profitability during the pandemic.
Therefore, in EDA, the median and IQR are used rather than the mean.

After that, we visualise ratio and percentage metrics by year on a linear scale using violin and jitter plots with median markers, faceted by variable.
It is used to assess distribution shape, central tendency, spread, and potential outliers across time without log transformation.

```{r IDA-ratio}
#chart for ratio variables (linear axis)
p_ratio <- ida_germany |>
  filter(variable_orig %in% ratio_vars) |>
  ggplot(aes(x = factor(acct_year), y = value)) +
  geom_violin(trim = FALSE, fill = NA, linewidth = 0.4) +
  geom_jitter(width = 0.12, height = 0, alpha = 0.12, size = 0.6) +
  stat_summary(fun = median, geom = "point", shape = 95, size = 6) +
  facet_wrap(~ variable, scales = "free_y", ncol = 3) +
  labs(x = "Year", y = NULL, title = "Violin + Jitter by Year (Ratios & % on Linear Scale)") +
  theme_minimal(base_size = 11) +
  theme(axis.text.x = element_text(size = 10),
        strip.text  = element_text(face = "bold"),
        panel.grid.minor = element_blank())

p_ratio
```

According to the chart, the profitability ratios (EBIT/EBITDA margins, ROA) cluster near small positives, with wider variation in 2020–2021.
ROE has the heaviest tails and many negatives, indicating pressure on equity returns during the pandemic.
Solvency is relatively stable with a slight dip around 2020; gearing is more dispersed and somewhat higher in 2020–2021 for some firms.
Liquidity (current/quick ratios) is generally low, with some moderate levels, but it shows long right tails, reflecting a few firms with strong buffers.
Efficiency (net assets and stock turnover) is broadly stable.
Therefore, in EDA, we report median and IQR rather than means, and we cap extreme values in figures only to improve readability while keeping original values for calculations.

Then, we measure missing data for key numeric variables by industry and by year, then visualise year-wise gaps with a heatmap.
It highlights variables or periods with substantial missingness, guiding imputation, exclusion, or robustness checks before analysis.

```{r numeric-missingness-check}
key_vars <- c("total_assets_eur","total_liabilities_eur","total_equity_eur",
              "total_revenue_eur","net_income_eur",
              "ebit_margin_pct","ebitda_margin_pct","roa_pct","roe_pct",
              "current_ratio","liquidity_ratio","gearing_pct","solvency_pct")

#by industry
missing_by_industry <- filtered_germany |>
  group_by(industry_group) |>
  summarise(across(all_of(key_vars), ~ mean(is.na(.))), .groups="drop") |>
  pivot_longer(-industry_group, names_to="variable", values_to="missing_rate")

#heatmap by year for numeric variables
missing_heat <- filtered_germany |>
  group_by(acct_year) |>
  summarise(across(all_of(key_vars), ~ mean(is.na(.))), .groups="drop") |>
  pivot_longer(-acct_year, names_to="variable", values_to="missing_rate")

ggplot(missing_heat, aes(variable, factor(acct_year), fill = missing_rate)) +
  geom_tile(color="white") +
  scale_fill_gradient(low="white", high="red", labels=scales::percent) +
  labs(x=NULL, y="Year", fill="Missing", title="Missingness by Year (Numeric Variables)") +
  theme_minimal() + theme(axis.text.x=element_text(angle=45, hjust=1))
```

According to the heatmap, missingness is concentrated in the gearing_pct and liquidity_ratio fields with around 20 – 25%, moderate for ebit_margin_pct/ebitda_margin_pct with around 10 – 15%, and low for net_income_eur/ROA/ROE.
The balance-sheet totals and revenue are essentially complete.
The pattern is stable across 2018 – 2021, with no pandemic-specific spike.
This means ratio analyses will have smaller sample sizes than level variables.

After that, we used this code to detect outliers in all numeric variables with two rules - IQR (1.5×) and standardized scores (\|z\| \> 3) - and summarise counts and proportions per variable, ordered by the IQR share.
It flags variables with heavy tails or extreme values to guide winsorisation, transformation, or sensitivity checks before modelling.

```{r outliers-check}
num_cols <- c("total_assets_eur","total_liabilities_eur","total_equity_eur",
              "total_revenue_eur","net_sales_eur","gross_sales_eur","net_income_eur",
              "ebit_margin_pct","ebitda_margin_pct","roa_pct","roe_pct",
              "net_assets_turnover","stock_turnover",
              "solvency_pct","gearing_pct","current_ratio","liquidity_ratio",
              "shareholders_liq_pct","interest_cover")

qfun <- function(x,p) quantile(x, probs=p, na.rm=TRUE, names=FALSE)

outliers_overall <- purrr::map_dfr(num_cols, function(v){
  x <- filtered_germany[[v]]
  q1 <- qfun(x,.25); q3 <- qfun(x,.75); iqr <- q3-q1
  lo <- q1 - 1.5*iqr; hi <- q3 + 1.5*iqr
  z  <- as.numeric(scale(x))
  tibble::tibble(
    variable=v,
    n=sum(!is.na(x)),
    iqr_outliers=sum(x<lo | x>hi, na.rm=TRUE),
    z_outliers=sum(abs(z)>3, na.rm=TRUE),
    share_iqr=iqr_outliers/n,
    share_z=z_outliers/n
  )
}) |> arrange(desc(share_iqr))

outliers_overall |>
  kable(align = "l", booktabs = TRUE) |>
  kable_styling(full_width = FALSE, font_size = 11)
```

According to the table, the share of extreme values per variable is assessed using two rules: the IQR fence (share_iqr) and a z-score cutoff (share_z).
Outliers are most common in net_income_eur (around 25%), EBIT margin (about 19%), and interest_cover (about 18%), while they are infrequent in solvency_pct (around 1%) and less common for turnover and gearing.
The z-score method identifies far fewer points because many variables are right-skewed, which makes mean/SD-based detection conservative.
The IQR rule provides a more informative screening here.

Then, we standardise key character fields by trimming whitespace, normalising case, and converting blanks to NA, and report missingness for identifiers and classification codes.
This ensures consistent keys for joins and grouping, and highlights data quality issues to address before analysis.

```{r character-check}
#character variables to check
char_vars <- c("company_id","company_name","country","city",
               "consolidation_code","nace_code","icb_code","sic_code")
char_vars <- intersect(char_vars, names(filtered_germany))

#cleaning
chars_clean <- filtered_germany |>
  mutate(
    across(all_of(char_vars),
           ~ .x |> as.character() |> str_squish() |> na_if("")),
    #casing
    country = str_to_title(country),
    city    = str_to_title(city),
    company_name = str_squish(company_name)
  )

#missingness check
char_missing <- chars_clean |>
  summarise(across(all_of(char_vars), ~ mean(is.na(.)))) |>
  tidyr::pivot_longer(everything(), names_to="variable", values_to="missing_rate")

char_missing |>
  kable(align = "l", booktabs = TRUE) |>
  kable_styling(full_width = FALSE, font_size = 11)
```

According to the table, identifiers and locations are complete: company_id, company_name, country, city, and consolidation_code show no missing data.
Industry codes are mostly available: NACE has about 2.0% missing, ICB around 6.2%, and SIC approximately 0.1%.
This indicates that while a small number of firms cannot be classified by NACE alone, most can be identified using the ICB/SIC fallbacks, which are categorised as industrial_group.

After completing the general IDA, we developed the specific IDA to address our sub-questions.
Here are the details:

### Question 1

We first add a period label (Pre-pandemic 2018–2019 vs. Pandemic 2020–2021).
For visuals, net_income_eur is transformed using asinh() so we can plot both gains and losses on a comparable scale.
Where ROA/ROE are missing, we recalculate simple fallbacks from the accounting totals and record the sources of the values (reported vs. recomputed).

```{r q1_prepare_profitability_inputs}
stopifnot("net_income_eur" %in% names(filtered_germany))

filtered_germany_1 <- filtered_germany |>
  mutate(
    # period flag
    period = if_else(acct_year >= 2020, "Pandemic (2020–2021)", "Pre-pandemic (2018–2019)"),
    # transform for visuals (handles negatives)
    net_income_eur_asinh = asinh(as.numeric(net_income_eur)),
    # recompute ROA / ROE (fallbacks)
    roa_calc = 100 * (net_income_eur / total_assets_eur),
    roe_calc = if_else(total_equity_eur > 0,
                       100 * (net_income_eur / total_equity_eur),
                       NA_real_),
    roa_use = coalesce(roa_pct, roa_calc),
    roe_use = coalesce(roe_pct, roe_calc),
    roa_src = case_when(
      !is.na(roa_pct) ~ "reported",
      !is.na(roa_calc) ~ "recomputed",
      TRUE ~ NA_character_
    ),
    roe_src = case_when(
      !is.na(roe_pct) ~ "reported",
      !is.na(roe_calc) ~ "recomputed",
      TRUE ~ NA_character_
    )
  )
```

Then, we perform a conservative within-firm, within-period “one-gap” fill, patching only a single missing year when adjacent values match.
It then restores reported values where available.

```{r q1_firm_one_gap_fill}
#Firm “one-gap” fill within the same period only

fill_one_gap <- function(x){
  f <- dplyr::lag(x); b <- dplyr::lead(x)
  ifelse(is.na(x) & !is.na(f) & !is.na(b) & (f == b), f, x)
}

filtered_germany_1 <- filtered_germany_1 |>
  arrange(company_id, acct_year) |>
  group_by(company_id, period) |>  #period boundary blocks carryover
  mutate(
    ebit_step1   = fill_one_gap(ebit_margin_pct),
    ebitda_step1 = fill_one_gap(ebitda_margin_pct),
    roa_step1    = fill_one_gap(roa_use),
    roe_step1    = fill_one_gap(roe_use)
  ) |>
  ungroup() |>
  mutate(
    ebit_step1   = coalesce(ebit_margin_pct,   ebit_step1),
    ebitda_step1 = coalesce(ebitda_margin_pct, ebitda_step1),
    roa_step1    = coalesce(roa_use,           roa_step1),
    roe_step1    = coalesce(roe_use,           roe_step1)
  )
```

After that, we impute remaining gaps following a strict hierarchy: first industry and year, then industry and period, and finally period, without crossing period boundaries.
This process produces the final Q1 variables (EBIT, EBITDA, ROA, ROE) and records the source of each value.

```{r q1_impute_hierarchy_finalize}
#median fills in strict order (no cross-period borrowing)
#industry × year medians
med_ixy <- filtered_germany_1 |>
  group_by(industry_group, acct_year) |>
  summarise(
    med_ebit   = median(ebit_step1,   na.rm=TRUE),
    med_ebitda = median(ebitda_step1, na.rm=TRUE),
    med_roa    = median(roa_step1,    na.rm=TRUE),
    med_roe    = median(roe_step1,    na.rm=TRUE),
    .groups="drop"
  )

#industry × period medians
med_ip <- filtered_germany_1 |>
  group_by(industry_group, period) |>
  summarise(
    med_ebit_ip   = median(ebit_step1,   na.rm=TRUE),
    med_ebitda_ip = median(ebitda_step1, na.rm=TRUE),
    med_roa_ip    = median(roa_step1,    na.rm=TRUE),
    med_roe_ip    = median(roe_step1,    na.rm=TRUE),
    .groups="drop"
  )

#period medians
med_p <- filtered_germany_1 |>
  group_by(period) |>
  summarise(
    med_ebit_p   = median(ebit_step1,   na.rm=TRUE),
    med_ebitda_p = median(ebitda_step1, na.rm=TRUE),
    med_roa_p    = median(roa_step1,    na.rm=TRUE),
    med_roe_p    = median(roe_step1,    na.rm=TRUE),
    .groups="drop"
  )

#join and impute with precedence, also explicit source flags
filtered_germany_1 <- filtered_germany_1 |>
  left_join(med_ixy, by = c("industry_group","acct_year")) |>
  left_join(med_ip,  by = c("industry_group","period")) |>
  left_join(med_p,   by = "period") |>
  mutate(
    #final values, using your precedence
    ebit_margin_q1   = if_else(is.na(ebitda_step1), ebit_step1, ebit_step1),
    ebit_margin_q1   = if_else(is.na(ebit_step1),   coalesce(med_ebit,   med_ebit_ip,   med_ebit_p),   ebit_step1),
    ebitda_margin_q1 = if_else(is.na(ebitda_step1), coalesce(med_ebitda, med_ebitda_ip, med_ebitda_p), ebitda_step1),
    roa_q1           = if_else(is.na(roa_step1),    coalesce(med_roa,    med_roa_ip,    med_roa_p),    roa_step1),
    roe_q1           = if_else(is.na(roe_step1),    coalesce(med_roe,    med_roe_ip,    med_roe_p),    roe_step1),

    #provenance for EBIT/EBITDA margins
    ebit_src = dplyr::case_when(
      !is.na(ebit_margin_pct) ~ "reported",
      is.na(ebit_margin_pct) & !is.na(ebit_step1) ~ "firm-1gap(period)",
      is.na(ebit_step1) & !is.na(med_ebit) ~ "ind×year",
      is.na(ebit_step1) & is.na(med_ebit) & !is.na(med_ebit_ip) ~ "ind×period",
      TRUE ~ "period"
    ),
    ebitda_src = dplyr::case_when(
      !is.na(ebitda_margin_pct) ~ "reported",
      is.na(ebitda_margin_pct) & !is.na(ebitda_step1) ~ "firm-1gap(period)",
      is.na(ebitda_step1) & !is.na(med_ebitda) ~ "ind×year",
      is.na(ebitda_step1) & is.na(med_ebitda) & !is.na(med_ebitda_ip) ~ "ind×period",
      TRUE ~ "period"
    ),

    #ROA/ROE provenance
    roa_src = dplyr::case_when(
      !is.na(roa_src) ~ roa_src,
      is.na(roa_src) & !is.na(roa_step1) ~ "firm-1gap(period)",
      is.na(roa_step1) & !is.na(med_roa) ~ "ind×year",
      is.na(roa_step1) & is.na(med_roa) & !is.na(med_roa_ip) ~ "ind×period",
      TRUE ~ "period"
    ),
    roe_src = dplyr::case_when(
      !is.na(roe_src) ~ roe_src,
      is.na(roe_src) & !is.na(roe_step1) ~ "firm-1gap(period)",
      is.na(roe_step1) & !is.na(med_roe) ~ "ind×year",
      is.na(roe_step1) & is.na(med_roe) & !is.na(med_roe_ip) ~ "ind×period",
      TRUE ~ "period"
    )
  ) |>
  select(-ends_with("_step1"))
```

Then, we summarise imputation provenance by period, counting observations that are reported, firm-filled, or imputed from industry/year/period medians.
It offers a transparent audit of data construction for Question 1.

```{r 1_imputation_summary}
impute_summary_q1 <- filtered_germany_1 |>
  summarise(
    .by = period,
    n_rows = n(),
    ebit_reported   = sum(ebit_src   == "reported",            na.rm=TRUE),
    ebit_firmgap    = sum(ebit_src   == "firm-1gap(period)",   na.rm=TRUE),
    ebit_ind_year   = sum(ebit_src   == "ind×year",            na.rm=TRUE),
    ebit_ind_period = sum(ebit_src   == "ind×period",          na.rm=TRUE),
    ebit_period     = sum(ebit_src   == "period",              na.rm=TRUE),

    ebitda_reported   = sum(ebitda_src   == "reported",          na.rm=TRUE),
    ebitda_firmgap    = sum(ebitda_src   == "firm-1gap(period)", na.rm=TRUE),
    ebitda_ind_year   = sum(ebitda_src   == "ind×year",          na.rm=TRUE),
    ebitda_ind_period = sum(ebitda_src   == "ind×period",        na.rm=TRUE),
    ebitda_period     = sum(ebitda_src   == "period",            na.rm=TRUE),

    roa_reported    = sum(roa_src == "reported",            na.rm=TRUE),
    roa_recomputed  = sum(roa_src == "recomputed",          na.rm=TRUE),
    roa_firmgap     = sum(roa_src == "firm-1gap(period)",   na.rm=TRUE),
    roa_ind_year    = sum(roa_src == "ind×year",            na.rm=TRUE),
    roa_ind_period  = sum(roa_src == "ind×period",          na.rm=TRUE),
    roa_period      = sum(roa_src == "period",              na.rm=TRUE),

    roe_reported    = sum(roe_src == "reported",            na.rm=TRUE),
    roe_recomputed  = sum(roe_src == "recomputed",          na.rm=TRUE),
    roe_firmgap     = sum(roe_src == "firm-1gap(period)",   na.rm=TRUE),
    roe_ind_year    = sum(roe_src == "ind×year",            na.rm=TRUE),
    roe_ind_period  = sum(roe_src == "ind×period",          na.rm=TRUE),
    roe_period      = sum(roe_src == "period",              na.rm=TRUE)
  )

impute_summary_q1 |>
  kable(align = "l", booktabs = TRUE) |>
  kable_styling(full_width = FALSE, font_size = 11)
```

According to the table, EBIT/EBITDA margins: Coverage is high, with about 84% of rows reported directly (3,399/4,071 pre, and 3,337/3,975 pandemic), and the remaining approximately 16% filled using (industry and year) medians.
No firm-gap or period-level fills were needed.
Furthermore, ROA: around 98 – 99% is available (or recomputed from net income/assets).
Only about 1% was recomputed (52 pre and 43 pandemic), and less than 0.3% used (industry and year) fallback (11 pre; 2 pandemic).
Moreover, ROE: approximately 96% are reported, around 0.5% are recomputed (16 – 18 rows), and about 3.6 – 3.8% use (industry and year) medians.
Overall, imputations are minimal, transparent, and unlikely to influence results.

### Question 2

The analysis focuses on leverage and liquidity indicators for active German firms between 2018 and 2021.
Key derived variables include the debt-to-assets ratio and the equity ratio, along with reported measures of gearing, solvency, and the current ratio.
Missing and extreme values were excluded to maintain consistency across year-to-year comparisons.

```{r}
# Create leverage and liquidity indicators 
lev_liq_data <- filtered_germany |>
  filter(acct_year %in% 2018:2021, status == "Active") |>
  mutate(
    debt_to_assets = total_liabilities_eur / total_assets_eur,
    equity_ratio   = total_equity_eur / total_assets_eur
  ) |>
  select(acct_year, industry_group, company_id, company_name,
         debt_to_assets, equity_ratio, gearing_pct, solvency_pct, current_ratio) |>
  drop_na(debt_to_assets, current_ratio)

# Summary statistics by year
lev_liq_summary <- lev_liq_data |>
  group_by(acct_year) |>
  summarise(
    median_debt_assets = median(debt_to_assets, na.rm = TRUE),
    iqr_debt_assets    = IQR(debt_to_assets, na.rm = TRUE),
    median_liquidity   = median(current_ratio, na.rm = TRUE),
    iqr_liquidity      = IQR(current_ratio, na.rm = TRUE),
    median_gearing     = median(gearing_pct, na.rm = TRUE),
    median_solvency    = median(solvency_pct, na.rm = TRUE),
    .groups = "drop"
  )

lev_liq_summary |>
  knitr::kable(digits = 2, caption = "Leverage and Liquidity Summary (2018–2021)") |>
  kableExtra::kable_styling(full_width = FALSE, font_size = 11)
# Based on lectures: 3 & 4
```

From 2018 to 2021, the median debt-to-assets ratio remained steady at around 0.50–0.51, while the median current ratio stayed near 1.8×, indicating stable leverage and liquidity levels through the pandemic period.
A slight rise in gearing (from 62% to 66%) alongside a modest decline in solvency (from 50% to 49%) suggests firms gradually relied more on debt financing but maintained short-term financial health.

### Question 3

1.  Filter relevant years and active firms for pandemic analysis (Lecture - 2 & 3)

```{r Filtering relevant data}
filtered_germany_q3 <- filtered_germany  |> 
  filter(acct_year >= 2020, acct_year <= 2022, status == "Active")
```

This code filters the dataset **filtered_germany** to keep only the records relevant for the **pandemic period (2020–2022)** and only **active firms.**

2.  Summary checks on key financial variables for completeness and basic stats (Lecture - 2,3 & 5)

```{r Summary checks}
options(scipen = 999)

financial_vars <- c("ebit_margin_pct", "gearing_pct", "current_ratio", "total_assets_eur")

summary_stats <- filtered_germany_q3 |>
  summarise(across(
    all_of(financial_vars),
    list(
      missing = ~sum(is.na(.)),
      mean = ~mean(., na.rm = TRUE),
      median = ~median(., na.rm = TRUE),
      sd = ~sd(., na.rm = TRUE),
      min = ~min(., na.rm = TRUE),
      max = ~max(., na.rm = TRUE)
    ),
    .names = "{.col}_{.fn}"
  ))

summary_stats_tidy <- summary_stats |>
  pivot_longer(
    everything(),
    names_to = c("Variable", "Stat"),
    names_pattern = "(.+)_(.+)",
    values_to = "Value"
  ) |>
  pivot_wider(names_from = Stat, values_from = Value)

# Force decimal precision
summary_stats_tidy <- summary_stats_tidy |>
  mutate(across(where(is.numeric), ~round(., 5)))

kable(summary_stats_tidy,
      caption = "Summary Statistics for Key Financial Variables (2020–2022)",
      col.names = c("Variable", "Missing", "Mean", "Median", "SD", "Min", "Max"),
      align = "lrrrrrr",
      escape = FALSE
) |>
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = FALSE, font_size = 15) |>
  column_spec(1, width = "5cm") |>
  scroll_box(width = "100%")

```

This code calculates key summary statistics - including missing values, mean, median, standard deviation, minimum, and maximum - for major financial variables such as profitability, leverage, liquidity, and firm size.
It provides a quick check on data completeness and helps understand the overall distribution of these variables for active German firms between 2020 and 2022.

3.  Outlier detection using IQR for key variables (function) (Lecture 5 & 6)

```{r Outlier detection}
outlier_summary <- function(x) {
  q1 <- quantile(x, 0.25, na.rm=TRUE)
  q3 <- quantile(x, 0.75, na.rm=TRUE)
  iqr <- q3 - q1
  lower <- q1 - 1.5 * iqr
  upper <- q3 + 1.5 * iqr
  sum(x < lower | x > upper, na.rm=TRUE)
}
outlier_counts <- filtered_germany_q3  |> 
  summarise(across(all_of(financial_vars), outlier_summary))
kable(outlier_counts, caption = "Count of Outliers for Key Financial Variables (IQR Method)")  |> 
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

```

This code defines a custom function to detect outliers in each key financial variable using the Interquartile Range (IQR) method.
It calculates the lower and upper bounds (1.5 × IQR below Q1 and above Q3) and counts how many observations fall outside this range.
The resulting table shows how many potential outliers exist for each financial metric, helping identify variables that may need further cleaning or transformation.

4.  Check distribution of missing data across years and variables (Lecture - 4 & 6)

```{r Distribution of missing data}
missing_long <- filtered_germany_q3  |> 
  select(acct_year, all_of(financial_vars))  |> 
  pivot_longer(cols = -acct_year, names_to = "variable", values_to = "value")  |> 
  mutate(missing = is.na(value))  |> 
  group_by(acct_year, variable)  |> 
  summarise(prop_missing = mean(missing), .groups = "drop")

ggplot(missing_long, aes(x = variable, y = prop_missing, fill = factor(acct_year))) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Proportion of Missing Values by Variable and Year",
       x = "Variable", y = "Proportion Missing", fill = "Year") +
  theme_minimal()
```

This code examines how much data is missing for key financial variables over time and presents it visually.
It starts by reshaping the dataset into a long format so that each variable and its value are in separate rows.
Then, it marks whether each observation is missing and calculates the proportion of missing values for each variable within each year.
Finally, it uses a grouped bar chart to show these proportions by variable and year.
This helps identify which financial indicators have higher missingness and whether data completeness has improved or worsened across the 2020–2022 period, offering insights into data reliability for further analysis.

5.  Prepare industry code mapping table for reader reference (sample snippet) (Lecture - 2 & 7)

```{r Mapping tables}
nace_codes <- filtered_germany_q3 |> 
  distinct(nace_num2, industry_group)  |> 
  arrange(nace_num2)

kable(nace_codes, caption = "NACE Level 2 Code to Industry Mapping") |> 
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

This code generates a reference table to help readers interpret industry-level results later in the report.
It begins by selecting unique combinations of `nace_num2` (the NACE Level 2 industry code) and `industry_group` from the filtered dataset, ensuring that each code is listed only once.
The table is then arranged in ascending order of `nace_num2` for readability.
Finally, the `kable()` and `kable_styling()` functions are used to display the results in a clean, professional format.
This table serves as a useful lookup guide, allowing readers to easily understand which industry each NACE code represents when reviewing the subsequent analyses and visualizations.

### Question 4

1.  Data Cleaning on Consolidation Codes

The raw data from Orbis contained financial statements under different consolidation codes (C1, C2, C\*, U1, U2).
To ensure consistency and comparability, we prioritized consolidated accounts which represent the entire economic entity.
For each firm-year observation, we retained the highest-priority available consolidated statement following the order: C1 \> C2 \> C\*.
Unconsolidated statements (U1, U2) were excluded from the analysis.

```{r Q4 IDA 1. Data Cleaning on Consolidation Codes}
germany_q4 <- filtered_germany |>
  mutate(
    consolidation_priority = case_when(
      consolidation_code == "C1" ~ 1,  # highest priority
      consolidation_code == "C2" ~ 2,
      consolidation_code == "C*" ~ 3,
      consolidation_code == "U1" ~ 99, # Very low priority, about to be filtered
      consolidation_code == "U2" ~ 99,
      TRUE ~ 99  # Handle any unexpected code
    )
  ) |>
  
  # Group by company and year, and keep only the highest priority records
  group_by(company_id, fy_year) |>
  arrange(consolidation_priority, .by_group = TRUE) |>
  slice(1) |> # Take the first row of each group, which has the highest priority
  ungroup() |>
  # Filter out the remaining non-consolidated records
  filter(consolidation_priority %in% c(1, 2, 3))
```

2.  Select the columns to be analyzed

```{r Q4 IDA 2. Select the columns to be analyzed}
germany_q4 <- germany_q4 |>
  select(company_id, fy_year, roa_pct, current_ratio, industry_group)
```


The analysis in this study is based on a dataset of company financial information. It is important to note that in this question we use the fiscal year, not the calendar year. A company's fiscal year is its own 12-month reporting period, which may not end in December. Using the fiscal year is more accurate for this research because it matches the company's true business cycle. This means that when we refer to "2020" in the data, we are talking about the fiscal year that was most affected by the pandemic, even if it does not exactly match the calendar year 2020.

3.  Convert data into a tsibble object using the tsibble package (Lecture 9, slide 5 & Reading week 9)

Each company was defined as a key and the financial year as the index.
Basic screening was performed to check for missing values, outliers, and duplicates.

```{r Q4 IDA 3. Convert data into a tsibble object using the tsibble package (Lecture 9, slide 5 & Reading week 9)}
germany_ts <- germany_q4 |>
  as_tsibble(key = company_id, index = fy_year) |>
  arrange(company_id, fy_year)
```

4.  Missing value - Check for missing value (Tutorial 8)

The initial examination of missing values revealed a relatively low proportion of incomplete data in the dataset.
As shown in @fig-missing_clu, the missing value analysis showed that only 0.7% of the total data was missing, while 99.3% of observations were complete.
ROA contained between 1% and 1.5% missing values across the dataset, while the current ratio had a slightly higher proportion of missing data, ranging from 2% to 2.5% (@fig-missing_plots).

```{r Q4 IDA 4. Missing value -Check for missing value}
#| label: fig-missing_plots
#| fig-cap: "The plots of Missing Value Distribution"

# Check for missing ROA and CR
miss_var <- 
  gg_miss_var(germany_ts, show_pct = TRUE) + 
  labs(title = "Distribution of Missing Values", 
       subtitle = "For all variables")

# Check for missing values by industry group
miss_industry <- germany_ts |> 
  group_by(industry_group) |> 
  gg_miss_var(show_pct = TRUE) + 
  labs(title = "Distribution of Missing Values", 
       subtitle = "By industry group")


# Missing case pattern
miss_case <- gg_miss_case(germany_ts, show_pct = TRUE) + 
  labs(title = "Distribution of Missing Case Pattern")
  
miss_var / miss_industry / miss_case
```

```{r Q4 EDA 4. Missing value -Check for missing value 2}
#| label: fig-missing_clu
#| fig-cap: "The plot of clustering to visualize missing patterns"

m3 <- vis_miss(germany_ts, cluster = TRUE, sort_miss = TRUE) +
  theme(axis.text.x = element_text(size = 8, angle = 45, hjust = 1),
        axis.text.y = element_text(size = 8))

m3
```

**Address missingness issues**

To address missing data, observations with missing values in either ROA or current ratio were removed from the dataset.
Additionally, to ensure temporal consistency across the study period, only industry groups that contained data for all three years (2019-2021) were retained.
The 'Other / Unmapped' category was removed from the study as it comprises companies without clear industry classification.

```{r Q4 IDA 4. Address missingness issues}
germany_ts <- germany_ts |>
  as_tibble() |>   # Convert to a normal tibble
  # Remove observations with missing ROA or CR
  filter(!is.na(roa_pct) & !is.na(current_ratio)) |>
   # Remove industries that don't have data in 2019-2021
  group_by(industry_group) |>
  filter(all(2019:2021 %in% fy_year)) |>
  filter(industry_group != "Other / Unmapped") |>
  ungroup() |>
  # Convert back to tsibble format
  as_tsibble(key = company_id, index = fy_year) |>
  arrange(company_id, fy_year)
```

5.  Summary

The trends in ROA and Current Ratio from 2017 to 2021 reveal a distinct pattern, potentially linked to the pandemic.
ROA showed significant volatility: it was stable pre-pandemic, fell to a negative value in 2020, but then recovered strongly to a peak of 5% in 2021.
In contrast, the Current Ratio remained relatively stable throughout the period, indicating consistent short-term liquidity despite the economic disruption.
This suggests that while profitability was heavily impacted, the company's ability to meet its short-term obligations was maintained.

```{r Q4 IDA 5. Summary}
#| label: tbl-summary
#| tbl-cap: "Trends in Mean Return on Assets (ROA) and Current Ratio from Pre-Pandemic to Post-Pandemic Years (2017-2021)"

kable(germany_ts |>
  index_by(fy_year) |>
  summarise(mean_roa = mean(roa_pct, na.rm = TRUE),
            mean_cr = mean(current_ratio, na.rm = TRUE)))
```

## EDA

Here are the EDA processes and results for the sub-questions addressing the primary question "How did the pandemic affect corporate financials in Germany?"

### Question 1

**1. How did profitability change across industries?**

This code recodes years into Pre (2018 – 2019) versus Pandemic (2020 – 2021), reshapes five profitability measures (EBIT, EBITDA, ROA, ROE, net income) to long format, and plots violin plots with jitter, median, and IQR overlays.
It allows us to compare central tendency and spread before versus during the pandemic across all firms; net income uses the asinh function, so negatives remain visible.

```{r q1_profitability_distributions_overall}
#recode period to short labels
filtered_germany_1 <- filtered_germany_1 |>
  mutate(period = if_else(acct_year >= 2020, "Pandemic (2020–21)", "Pre (2018–19)"),
         period = factor(period, levels = c("Pre (2018–19)", "Pandemic (2020–21)")))

#build long data
q1_vars <- c("ebit_margin_q1","ebitda_margin_q1","roa_q1","roe_q1","net_income_eur_asinh")

plot_long <- filtered_germany_1 |>
  dplyr::select(company_id, period, dplyr::all_of(q1_vars)) |>
  tidyr::pivot_longer(-c(company_id, period),
                      names_to = "metric", values_to = "value") |>
  dplyr::mutate(
    metric = factor(metric, levels = q1_vars,
                    labels = c("EBIT margin (%)","EBITDA margin (%)",
                               "ROA (%)","ROE (%)","Net income (asinh)"))
  ) |>
  tidyr::drop_na(value)

iqr_bar <- function(y) data.frame(
  y    = median(y, na.rm=TRUE),
  ymin = quantile(y, 0.25, na.rm=TRUE),
  ymax = quantile(y, 0.75, na.rm=TRUE)
)

p_q1_all <- ggplot(plot_long, aes(x = period, y = value)) +
  geom_violin(trim = FALSE, fill = NA, color = "grey60", linewidth = 0.45, show.legend = FALSE) +
  geom_jitter(aes(color = "Data"), width = 0.08, height = 0, alpha = 0.10, size = 0.55, show.legend = TRUE) +
  stat_summary(aes(color = "IQR"), fun.data = iqr_bar, geom = "errorbar",
               width = 0.12, linewidth = 0.6, show.legend = TRUE) +
  stat_summary(aes(color = "Median"), fun = median, geom = "point",
               shape = 95, size = 8, show.legend = TRUE) +
  facet_wrap(~ metric, scales = "free_y", ncol = 2) +
  scale_x_discrete(labels = c(
    `Pre (2018–19)`      = "Pre\n(2018 – 2019)",
    `Pandemic (2020–21)` = "Pandemic\n(2020 – 2021)"
  )) +
  scale_color_manual(
    name = NULL,
    values = c("Data" = "grey35", "IQR" = "firebrick", "Median" = "black")
  ) +
  guides(color = guide_legend(override.aes = list(
    shape    = c(16, NA, 95),   # point for Data, line for IQR, thick tick for Median
    linetype = c(0, 1, 0),
    size     = c(2, 1, 6),
    alpha    = 1
  ))) +
  coord_cartesian(clip = "off") +
  labs(x = NULL, y = NULL,
       title = "Pre vs Pandemic: Profitability Distributions by Metric\n(Median + IQR)") +
  theme_minimal(base_size = 11) +
  theme(
    strip.text  = element_text(face = "bold"),
    axis.text.x = element_text(size = 10, margin = margin(t = 8), lineheight = 0.95),
    plot.margin = margin(10, 15, 24, 10),
    legend.position = "top"
  )

p_q1_all
```

The chart compares profitability before (2018–19) and during the pandemic (2020 – 2021) for all firms.
Across EBIT and EBITDA margins, the medians remain largely unchanged, but the spreads widen in 2020 – 2021, indicating increased dispersion in operating performance.
ROA and especially ROE show many more extreme negatives in 2020 – 2021, indicating that a greater share of firms reported losses or equity write-downs, even though the median stayed near zero.
Net income (asinh) also becomes more dispersed in 2020 –21, with a heavier lower tail, confirming higher volatility.
A key insight is that central tendencies barely moved while tail risk increased.

This code retains the mapped industries, calculates industry medians for EBITDA margin during the Pre and Pandemic periods, and creates a dumbbell chart showing the two medians and the change (Pandemic – Pre), coloured by increase or decrease.
Using medians ensures the comparison is robust against outliers and highlights which sectors improved or worsened the most.

```{r q1_industry_ebitda_median_delta}
#keep only mapped industry
germ_mapped <- filtered_germany_1 |>
  filter(!is.na(ebitda_margin_q1),
         !is.na(industry_group),
         industry_group != "Other / Unmapped") |>
  droplevels()

#build medians & delta (Pandemic – Pre)
eb_med <- germ_mapped |>
  mutate(period = if_else(acct_year >= 2020, "Pandemic", "Pre"),
         period = factor(period, levels = c("Pre","Pandemic"))) |>
  group_by(industry_group, period) |>
  summarise(med = median(ebitda_margin_q1, na.rm = TRUE), .groups = "drop") |>
  pivot_wider(names_from = period, values_from = med) |>
  drop_na(Pre, Pandemic) |> #keep industries with both periods
  mutate(
    delta     = Pandemic - Pre,
    direction = factor(if_else(delta >= 0, "Increase", "Decrease"),
                       levels = c("Increase","Decrease"))
  ) |>
  arrange(delta)

#create and colored dumbbell
ggplot(eb_med, aes(y = reorder(industry_group, delta))) +
  geom_segment(aes(x = Pre, xend = Pandemic, yend = industry_group, color = direction),
               linewidth = 1) +
  geom_point(aes(x = Pre),      color = "#4E79A7", size = 2) +  #pre (blue)
  geom_point(aes(x = Pandemic), color = "#E15759", size = 2) +  #pandemic (red)
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey70") +
  scale_color_manual(values = c("Increase" = "#2CA02C", "Decrease" = "#D62728")) +
  guides(color = guide_legend(title = NULL, override.aes = list(linewidth = 3))) +
  labs(title = "Change in Median EBITDA Margin\n(Pandemic – Pre)",
       x = "Median EBITDA Margin (%)", y = NULL) +
  theme_minimal(base_size = 11)

```

This dumbbell chart compares the median EBITDA margin (%) in each industry before and during the pandemic.
It shows the direction and magnitude of profitability changes across sectors on a single axis.
Overall, most industries moved only slightly, but a few stand out: Accommodation & Food shows the biggest improvement, bouncing from slightly negative to positive margins, and Technology also saw an increase.
The largest drop is in Health Care, which fell sharply from a high pre-pandemic median to a much lower level.
Cyclical sectors like Industrials, Construction, Arts & Entertainment, and Financials experienced small to moderate decreases, while Utilities and Education remain near flat.
Using medians (rather than means) makes the comparison more robust to outliers and highlights typical firms within each industry.

### Question 2

**2. How did leverage (debt-to-assets) and liquidity (current ratio) evolve during pandemic?**

This question examines whether German firms altered their capital structure or short-term liquidity positions in response to the COVID-19 shock.The analysis proceeds in three parts:

1.  Time trends of leverage and liquidity

```{r}
#| label: fig-leverage-liquidity-two-panels
lev_liq_long <- lev_liq_summary |>
  select(acct_year, median_debt_assets, median_liquidity) |>
  pivot_longer(-acct_year, names_to = "indicator", values_to = "value")

# Debt-to-Assets panel — y-axis ticks at 0.50 and 0.51 only
p_debt <- lev_liq_long |>
  filter(indicator == "median_debt_assets") |>
  ggplot(aes(acct_year, value)) +
  geom_line(colour = "#E69F00", linewidth = 1.2) +
  geom_point(colour = "#E69F00", size = 3) +
  scale_y_continuous(
    limits = c(0.495, 0.515),
    breaks = c(0.50, 0.51),
    labels = scales::number_format(accuracy = 0.01)
  ) +
  labs(x = NULL, y = "Median Value", title = "Debt-to-Assets") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"),
        panel.grid.minor = element_blank())

# Current Ratio panel — its own sensible scale
p_curr <- lev_liq_long |>
  filter(indicator == "median_liquidity") |>
  ggplot(aes(acct_year, value)) +
  geom_line(colour = "#0072B2", linewidth = 1.2) +
  geom_point(colour = "#0072B2", size = 3) +
  scale_y_continuous(labels = scales::number_format(accuracy = 0.01)) +
  labs(x = "Accounting Year", y = "Median Value", title = "Current Ratio") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"),
        panel.grid.minor = element_blank())

# place side-by-side with a common title
p_debt + p_curr + plot_annotation(title = "Trends in Leverage and Liquidity, 2018–2021") &
  theme(plot.title = element_text(face = "bold", size = 15))

# Based on Lectures: 3, 4, 5 & 6
```

The plots show that liquidity (current ratio) remained **broadly stable** from 2018 to 2021, fluctuating only slightly around **1.8** .
Meanwhile, leverage (debt-to-assets) **rose modestly** from 0.50 to 0.51, peaking in 2020 before easing slightly in 2021.
This pattern suggests that while firms **marginally increased** their reliance on debt financing during the pandemic, they continued to maintain sufficient short-term liquidity.
Overall, German firms demonstrated a balanced financial position, reflecting resilience and prudence in managing both debt and liquidity pressures through the COVID-19 period.

2.  Leverage vs Solvency (are highly levered firms less solvent?)

```{r}
#| label: fig-leverage-solvency
lev_liq_plot <- lev_liq_data |>
  filter(is.finite(debt_to_assets),
         debt_to_assets >= 0, debt_to_assets <= 1.5,   # 0–150%
         solvency_pct >= -50, solvency_pct <= 100)

ggplot(lev_liq_plot, aes(x = debt_to_assets, y = solvency_pct)) +
  geom_point(alpha = 0.35, colour = "#0072B2") +
  geom_smooth(method = "lm", se = FALSE, colour = "#E69", linewidth = 0.9) +
  geom_hline(yintercept = 50, linetype = "dotted", colour = "grey") +
  facet_wrap(~ acct_year, ncol = 2) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, 1.5)) +
  scale_y_continuous(limits = c(-50, 100)) +
  labs(
    title = "Leverage–Solvency Relationship by Year",
    subtitle = "Filtered to 0–150% debt-to-assets to avoid scale distortion from extreme values",
    x = "Debt-to-Assets",
    y = "Solvency Ratio (%)"
  ) +
  theme_minimal(base_size = 13) +
  theme(strip.text = element_text(face = "bold"),
        plot.title = element_text(face = "bold"))
# Based on Lectures : 3, 4, 5, 6 & 7
```

The scatterplots show a **strong and consistent negative relationship** between leverage (debt-to-assets) and solvency (ability to meet long-term obligations) across all years from 2018 to 2021.
Each panel demonstrates a clear downward slope: as firms take on more debt, solvency ratios decline.
This pattern holds consistently over time, implying that highly leveraged firms are structurally less solvent, and that the pandemic did not materially alter this trade-off.
The persistence of this relationship suggests that balance-sheet fragility scales proportionally with debt exposure, independent of year-specific shocks.

3.  Debt-to-assets by industry, faceted by year

```{r}
#| label: fig-leverage-industry-facet
#| fig-width: 10
#| fig-height: 12

lev_liq_industry <- lev_liq_data |>
  filter(debt_to_assets >= 0, debt_to_assets <= 1.5) |>
  drop_na(industry_group)

ggplot(lev_liq_industry, aes(x = debt_to_assets, y = reorder(industry_group, debt_to_assets), fill = factor(acct_year))) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7, width = 0.6) +
  facet_wrap(~ acct_year, ncol = 2, scales = "free_x") +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, 1.5)) +
  scale_fill_brewer(palette = "Set2", name = "Accounting Year") +
  labs(
    title = "Distribution of Leverage (Debt-to-Assets) by Industry, 2018–2021",
    subtitle = "Faceted view highlights stability in industry leverage patterns through the pandemic period",
    x = "Debt-to-Assets Ratio",
    y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    strip.text = element_text(face = "bold", size = 13),
    plot.title = element_text(face = "bold", size = 17, hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5),
    legend.position = "top",
    axis.text.y = element_text(size = 10),
    panel.grid.minor = element_blank(),
    panel.spacing.y = unit(1, "lines")  # increases vertical space between facets
  )
# Based on Lectures : 3, 4, 5 & 6
```

According to the faceted boxplot, the distribution of leverage (debt-to-assets) across industries remained broadly stable from 2018 to 2021.
Capital-intensive sectors such as Manufacturing, Construction, and Transport & Storage consistently exhibit higher leverage, while service-based industries like Information & Communication and Professional & Technical Services maintain comparatively lower ratios.
The slight widening of the distributions in 2020–2021 suggests that some firms increased borrowing during the pandemic to support liquidity, while others remained conservative.
Overall, the median leverage levels changed little across years, indicating that industry structure rather than pandemic-specific effects primarily determined leverage patterns in German firms.

### Question 3

**3. Which German industries showed unexpected financial resilience or vulnerability from 2020 to 2022, and how were these outcomes shaped by profitability, leverage, liquidity, and firm size?**

1.  Summarize Financial Indicators & Compute Resilience Score (Lecture - 3, 5)

```{r Financial Indicators}
industry_summary <- filtered_germany |> 
  filter(acct_year >= 2020, acct_year <= 2022)  |> 
  group_by(nace_num2, acct_year)  |> 
  summarise(
    median_profitability = median(ebit_margin_pct, na.rm = TRUE),
    median_leverage = median(gearing_pct, na.rm = TRUE),
    median_liquidity = median(current_ratio, na.rm = TRUE),
    median_firm_size = median(total_assets_eur, na.rm = TRUE),
    .groups = "drop"
  )  |> 
  mutate(resilience_score = median_profitability / median_leverage * median_liquidity)

```

**Interpretation**

This code calculates summary financial indicators for German industries during the pandemic years 2020–2022.
First, it filters the dataset to include only accounting years 2020, 2021, and 2022.
Then, it groups the data by `nace_num2` (industry code) and accounting year.
For each industry-year combination, it computes the median EBIT margin (`median_profitability`) as a measure of profitability, the median gearing ratio (`median_leverage`) as a measure of financial leverage, the median current ratio (`median_liquidity`) as a measure of short-term liquidity, and the median total assets (`median_firm_size`) as a proxy for firm size.
The resulting `industry_summary` dataset provides a concise overview of how different industries performed financially across these key dimensions during the pandemic period.

2.  Visualize Overall Financial Resilience with Bubble Chart (Lecture - 4, 10)

```{r Financial resilience}
p <- ggplot(industry_summary, aes(
  x = median_profitability,
  y = median_leverage,
  color = median_liquidity,
  text = paste(
    "NACE:", nace_num2,
    "<br>Profitability:", round(median_profitability, 2),
    "<br>Leverage:", round(median_leverage, 2),
    "<br>Liquidity:", round(median_liquidity, 2)
  )
)) +
  geom_point(size = 4, alpha = 0.85, shape = 21, stroke = 0.7) +
  # Use a richer color gradient here:
  scale_color_gradientn(colors = turbo(20)) +  
  labs(
    title = "German Industry Profitability vs Leverage (2020–2022)",
    subtitle = "Color indicates median liquidity by industry; axes limited for interpretability",
    x = "Median EBIT Margin (%)",
    y = "Median Gearing (%)",
    color = "Liquidity"
  ) +
  xlim(-10, 20) +
  ylim(0, 250) +
  theme_bw(base_size = 15) +
  theme(
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5, margin = margin(b = 6)),
    plot.subtitle = element_text(size = 12, hjust = 0.5, margin = margin(b = 10)),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(color = "black"),
    panel.grid.major = element_line(color = "grey85"),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    legend.position = "right",
    legend.title = element_text(face = "bold", size = 10),
    legend.text = element_text(size = 11),
    plot.margin = margin(16, 16, 16, 16)
  )

# Interactive plot with more gradient colours
ggplotly(p, tooltip = "text")

```

**Interpretation**

This interactive bubble chart illustrates the **financial resilience of German industries from 2020–2022** by comparing key financial indicators.
The **x-axis** represents median profitability (EBIT margin), while the **y-axis** shows median leverage (gearing ratio).
Bubble **color** reflects liquidity levels, helping highlight which industries are more financially stable.
Industries positioned **to the right and lower** on the chart indicate stronger financial health - combining higher profitability, lower leverage, and higher liquidity - whereas those toward the **left or upper areas** suggest weaker resilience and greater financial risk.

3.  Top and Bottom 10 Industries by Resilience Score (Lecture - 3, 4)

```{r Top & bottom 10 industries}
resilience_summary <- industry_summary  |> 
  group_by(nace_num2)  |> 
  summarise(avg_resilience = mean(resilience_score, na.rm = TRUE))  |> 
  arrange(desc(avg_resilience))

top_bottom_resilience <- resilience_summary  |> 
  slice_head(n = 10) |> 
  bind_rows(resilience_summary |> slice_tail(n = 10))

ggplot(top_bottom_resilience, aes(
  x = reorder(nace_num2, avg_resilience),
  y = avg_resilience,
  fill = avg_resilience
)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  scale_fill_viridis_c(option = "plasma", direction = -1) +
  labs(
    title = "Top and Bottom 10 Financially Resilient German Industries (2020–2022)",
    subtitle = "Industries ranked by average resilience scores",
    x = "Industry (NACE Level 2)",
    y = "Average Resilience Score"
  ) +
  theme_bw(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 11)
  )

```

**Interpretation**

This horizontal bar chart visualizes the distribution of average financial resilience scores across German industries from 2020 to 2022.
Each bar represents a NACE Level 2 industry, ordered from highest to lowest average resilience score.
The resilience score integrates key financial metrics- profitability, leverage, and liquidity , to assess how well industries managed financial stability during the pandemic.
Industries with higher scores demonstrated stronger ability to maintain earnings, control debt levels, and preserve liquidity, indicating greater financial robustness.
The color gradient emphasizes this ranking, with brighter colors reflecting more resilient sectors.
This visualization clearly highlights sectors that were comparatively vulnerable or resilient in navigating the economic challenges posed by COVID-19.

4.  Table of Top and Bottom 5 Industries (Lecture - 2, 6)

```{r Top & Bottom 5 industries}

top_bottom_resilience |> 
  mutate(rank = rank(-avg_resilience))  |> 
  filter(rank <= 5 | rank > max(rank) - 5)  |> 
  arrange(rank)  |> 
  rename(
    `NACE Code` = nace_num2,
    `Average Resilience Score` = avg_resilience,
    `Rank` = rank
  )  |> 
  kable(
    caption = "Top and Bottom 5 German Industries by Financial Resilience (2020–2022)",
    digits = 2,
    col.names = c("NACE Code", "Average Resilience Score", "Rank")
  )  |> 
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = FALSE,
    position = "center",
    font_size = 12
  )  |> 
  row_spec(0, bold = TRUE, background = "#f2f2f2")  |> 
  column_spec(2, color = "darkgreen") |> 
  column_spec(3, bold = TRUE, color = "darkblue")

```

**Interpretation**

This table ranks German industries by their **financial resilience from 2020 to 2022**, combining profitability, leverage, and liquidity into a composite resilience score.
Higher scores indicate industries that sustained earnings while keeping debt manageable and maintaining strong cash positions during the pandemic.
The ranking shows that **healthcare, technology, and utilities** were the most resilient sectors, benefitting from steady demand and robust balance sheets.
In contrast, **hospitality, retail, and manufacturing** were the least resilient, as lockdowns, supply disruptions, and high fixed costs eroded profitability and liquidity.
Overall, the results highlight how structural factors—such as firm size, capital intensity, and sectoral exposure—shaped the uneven financial recovery of German industries during COVID-19.

5.  Trends in Median EBIT Margin by Key German Industries (2018–2021) (Lecture - 4, 9)

```{r Time Trend}
key_industries <- c("Health Care", "Manufacturing", "Accommodation & Food", "Information & Communication")

plot_data <- filtered_germany  |> 
  filter(industry_group %in% key_industries, acct_year %in% 2018:2021)  |> 
  group_by(industry_group, acct_year)  |> 
  summarise(median_ebit = median(ebit_margin_pct, na.rm = TRUE), .groups = "drop")

ggplot(plot_data, aes(x = acct_year, y = median_ebit, color = industry_group)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  geom_text_repel(
    data = plot_data |> 
      filter(acct_year == max(acct_year)),
    aes(label = industry_group),
    nudge_x = 0.3,
    direction = "y",
    hjust = 0,
    segment.size = 0.25,
    show.legend = FALSE,
    size = 4
  ) +
  scale_x_continuous(breaks = 2018:2021, limits = c(2018, 2022)) +
  scale_color_brewer(palette = "Dark2") +
  labs(
    title = "Trends in Median EBIT Margin by Key German Industries (2018–2021)",
    x = "Accounting Year",
    y = "Median EBIT Margin (%)",
    color = "Industry"
  ) +
  theme_bw(base_size = 10) +
  theme(
    plot.title = element_text(face = "bold", size = 12, hjust = 0.5, margin = margin(b = 10)),
    axis.text = element_text(color = "black"),
    axis.title = element_text(face = "bold"),
    legend.position = "none",
    panel.grid.minor = element_blank(),
    plot.margin = margin(t = 20, r = 20, b = 10, l = 10)
  )

```

**Interpretation**

The figure shows trends in **median EBIT margins** for four key German industries between **2018 and 2021**.
**Health Care** remains the most stable and profitable sector, while **Manufacturing** shows moderate recovery after 2020.
**Accommodation & Food** experiences the sharpest decline during 2020, reflecting pandemic effects, whereas **Information & Communication** maintains steady profitability throughout.
These patterns highlight clear **sectoral differences in financial resilience and recovery** across the German economy.

### Question 4

**4. How did the fundamental relationship between corporate liquidity and profitability evolve and fracture within industries during the pandemic?**

1.  Industry-level trends of ROA and liquidity (2017–2021) (Lecture 9, slide 24-26, feasts and tsibbletalk)

This section provides an initial visual overview of the pandemic's impact.
We begin by plotting the average Return on Assets (ROA) and Current Ratio for each industry from 2017 to 2021.

The interactive ROA trends (@fig-combined_plot) reveal distinct pandemic responses across industries:

-   Accommodation & Food surged temporarily in 2019 but declined again in 2020.

-   Industrials showed a persistent decline that accelerated post-2020, while Information & Communication rebounded sharply after an initial dip.

-   Financial & Insurance maintained the highest ROA with steady growth throughout, demonstrating remarkable sector resilience.

Based on the Current Ratio trends (@fig-combined_plot), most industries demonstrated relative stability throughout the period.
However, two sectors exhibited notable deviations:

-   Consumer Services peaked at 8.14 in 2020 then plummeted to 0.21.

-   Arts, Entertainment & Recreation showed a significant upward trend in liquidity, which appears counterintuitive given pandemic constraints.

-   Financials experienced a substantial decline from 7.8 in 2020 to 4.21 in 2021, indicating a sharp contraction in liquidity buffers during the post-pandemic recovery phase.

-   Construction maintained exceptional stability throughout the period.

```{r Q4 EDA 1. Industry-level trends -Calculate}
# Calculate the average ROA and current ratio for each industry each year
industry_ts <- germany_ts |>
  group_by(industry_group) |>
  index_by(fy_year) |>
  summarise(
    industry_roa = mean(roa_pct, na.rm = TRUE),
    industry_cr = mean(current_ratio, na.rm = TRUE),
    n_companies = n()  
  )

```

```{r Q4 EDA 1. Industry-level trends}
#| label: fig-combined_plot
#| fig-cap: "The interactive plot of ROA and Current Ratio trends by industry (2017-2021)"

# ROA
roa_interactive <- industry_ts |>
  ggplot(aes(x = fy_year, y = industry_roa, group = industry_group)) +
  geom_line(aes(color = industry_group), size = 0.3, alpha = 0.7) +
  geom_vline(xintercept = 2020, linetype = "dashed", color = "red", size = 0.5) +
  labs(title = "Interactive ROA Trends by Industry (2017-2021)<br><sub><i>Vertical line indicates pandemic start (2020)<br><sub><i>",
       x = "Fiscal Year", y = "ROA (%)") +
  theme_minimal() +
  theme(legend.position = "none") 

# Plotly interactive charts -ROA
roa_plotly <- 
  ggplotly(roa_interactive, tooltip = c("industry_group", "fy_year", "industry_roa")) 


# Current ratio
cr_interactive <- industry_ts |>
  ggplot(aes(x = fy_year, y = industry_cr, group = industry_group)) +
  geom_line(aes(color = industry_group), size = 0.3, alpha = 0.7) +
  geom_vline(xintercept = 2020, linetype = "dashed", color = "red", size = 0.5) +
  labs(title = "Interactive Current Ratio Trends by Industry (2017-2021)<br><sub><i>Vertical line indicates pandemic start (2020) - Hover for details<br><sub><i>",
       x = "Fiscal Year", y = "Current Ratio") +
  theme_minimal() +
  theme(legend.position = "none")

cr_plotly <- 
  ggplotly(cr_interactive, tooltip = c("industry_group", "fy_year", "industry_cr")) 

# Combine plots with controlled layout
combined_plot <- subplot(
  roa_plotly, cr_plotly,
  nrows = 2,
  shareX = TRUE,  
  titleY = TRUE  
) |>
  plotly::layout(
    title = "Industry Financial Trends (2017-2021)<br><sub>Vertical line indicates pandemic start (2020)</sub>",
    hoverlabel = list(bgcolor = "white", font = list(size = 10)))

combined_plot
```

2.  Macro-level evidence: Industry recovery patterns (ROA vs CR) - feasts (Lecture 9, slide 24-26, feasts and tsibbletalk)

Now, we will analyze the fundamental relationship between corporate liquidity and profitability across industries during the pandemic period, with particular focus on identifying anomalous recovery patterns.

```{r Q4 EDA 2. Macro-level evidence -Calculate}
# Calculating time series features
industry_features <- industry_ts |>
  features(industry_roa, feat_stl) |>
  rename_with(~ paste0("roa_", .), -industry_group) |>
  left_join(
    industry_ts |>
      features(industry_cr, features = feat_stl) |>
      rename_with(~ paste0("cr_", .), -industry_group),
    by = "industry_group"
  )

# Focus on trend_strength and spikiness
focus_features <- industry_features |>
  select(industry_group, 
         roa_trend_strength, roa_spikiness,
         cr_trend_strength, cr_spikiness)
```

**Trend strength analysis**

The trend strength visualization in @fig-combined_analysis reveals significant disruptions to industry financial fundamentals.
Five industries in the bottom-left quadrant, indicating severe deterioration in both ROA and Current Ratio trends.
Accommodation & Food has the largest bubble, reflecting extreme volatility and weak performance during pandemic restrictions.
This pattern reflects the extreme operational challenges faced by hospitality-related sectors during pandemic restrictions.

**Volatility analysis**

After trend strength analysis, the volatility examination in @fig-combined_analysis demonstrates divergent patterns in financial metric instability.
Five industries show high ROA volatility, but most have moderate Current Ratio spikiness.
However, Consumer Services is exceptional with the highest Current Ratio spikiness, suggesting unique liquidity challenges.


```{r Q4 EDA 2. Macro-level evidence -Combined analysis}
#| label: fig-combined_analysis
#| fig-cap: "Combined analysis of trend strength disruption and volatility across industries during the pandemic period. Left: Trend strength analysis with 0.3 threshold lines. Right: Volatility analysis with median lines. Industries in bottom-left quadrant of left plot are most severely impacted, while top-right industries in right plot have highest volatility."

# Trend strength analysis to identify the hardest-hit sectors
trend_analysis <- focus_features |>
  mutate(
    # Identify sectors where trends are broken (low trend strength)
    roa_trend_disrupted = roa_trend_strength < 0.3,
    cr_trend_disrupted = cr_trend_strength < 0.3,
    overall_trend_impact = (roa_trend_strength + cr_trend_strength) / 2
  )


# Volatility analysis to identify the sectors with the greatest volatility during the pandemic
volatility_analysis <- focus_features |>
  mutate(
    high_roa_volatility = roa_spikiness > median(roa_spikiness),
    high_cr_volatility = cr_spikiness > median(cr_spikiness),
    overall_volatility = (scale(roa_spikiness) + scale(cr_spikiness)) / 2)



# Create trend strength plot
trend_plot <- focus_features |>
  ggplot(aes(x = roa_trend_strength, y = cr_trend_strength)) +
  geom_point(aes(size = roa_spikiness, 
                 color = industry_group, 
                 text = industry_group), alpha = 0.7) +
  geom_hline(yintercept = 0.3, linetype = "dashed", color = "red", alpha = 0.7) +
  geom_vline(xintercept = 0.3, linetype = "dashed", color = "red", alpha = 0.7) +
   labs(title = "Impact of Pandemic on Industry Trend Strength<br><sub><i>Red dashed lines indicate trend strength threshold (0.3)<br><sub><i>",
       x = "ROA Trend Strength", y = "Current Ratio Trend Strength",
       size = "ROA Spikiness") +
  theme_minimal() +
  theme(legend.position = "none")

# Convert to interactive plotly
trend_in <- ggplotly(trend_plot, tooltip = c("text", "x", "y", "size")) 


# Create volatility plot
volatility_plot <- focus_features |>
  ggplot(aes(x = roa_spikiness, y = cr_spikiness)) +
  geom_point(aes(color = industry_group, 
                 size = roa_trend_strength, 
                 text = industry_group), alpha = 0.7) +
  geom_hline(yintercept = median(focus_features$cr_spikiness), 
             linetype = "dashed", color = "blue", alpha = 0.5) +
  geom_vline(xintercept = median(focus_features$roa_spikiness), 
             linetype = "dashed", color = "blue", alpha = 0.5) +
  coord_cartesian(xlim = c(0, quantile(focus_features$roa_spikiness, 0.90))) + # Limit to 75% quantile
  labs(title = "Industry Volatility Analysis During Pandemic<br><sub><i>Blue dashed lines represent medians, excluding the extreme values<br><sub><i>",
       x = "ROA Spikiness", y = "Current Ratio Spikiness",
       size = "ROA Trend Strength") +
  theme_minimal() +
  theme(legend.position = "none")


volatility_in <- ggplotly(volatility_plot, tooltip = c("text", "x", "y", "size"))



combined_analysis <- subplot(
  trend_in, volatility_in,
  nrows = 1,
  shareY = FALSE,
  titleX = TRUE,
  titleY = TRUE,
  margin = 0.05 
) |>
  plotly::layout(
    title = list(
      text = "Industry Financial Impact Analysis During Pandemic<br><sub>Left: Trend strength disruption | Right: Volatility patterns</sub>",
      font = list(size = 14, color = "black"),
      x = 0.5,
      y = 0.95
    ),
    hoverlabel = list(
      bgcolor = "white", 
      font = list(size = 11, color = "black"),
      bordercolor = "lightgray"
    ),
    margin = list(l = 60, r = 60, b = 80),
    hovermode = "closest",
    annotations = list(
      list(
        x = 0, y = -0.15,
        xref = "paper", yref = "paper",
        text = "Red lines: 0.3 threshold",
        showarrow = FALSE,
        font = list(size = 10, color = "red")
      ),
      list(
        x = 0.85, y = -0.15,
        xref = "paper", yref = "paper",
        text = "Blue lines: Median volatility values",
        showarrow = FALSE,
        font = list(size = 10, color = "blue")
      )
    )
  )


combined_analysis
```




**Threshold-based industry classification**

Given the exploratory focus of our study and limited emphasis on clustering methods in lectures, we instead used threshold-based segmentation to classify industries into recovery archetypes.
This approach categorizes industries based on their ROA trend strength and volatility using the following criteria:

-   High trend strength (≥0.7) + Low spikiness (≤median) → "Resilient industries"

-   Low trend strength (≤0.3) + High spikiness (≥75th percentile) → "K-shaped recovery industries"

-   High trend strength (≥0.7) + High spikiness (≥75th percentile) → "Hidden crisis industries"

-   All other combinations → "Moderate impact industries"

Based on criteria, the analysis reveals clear patterns of differential recovery across sectors.
The Construction and Financial & Insurance sectors demonstrate resilient characteristics, maintaining stable performance throughout the pandemic period.
Arts, Entertainment & Recreation , Accommodation & Food Services, and Information & Communication exhibit K-shaped recovery patterns.
The Administrative & Support sector falls into the hidden crisis category, suggesting underlying structural issues.
This classification provides valuable insights into how different industries navigated the pandemic's challenges.
Next part, we will focus on the anomalous K-shaped recoveries that suggest uneven performance within certain sectors and resilient industry.

```{r Q4 EDA 2. Macro-level evidence -industry classes}
industry_classes <- focus_features |>
  mutate(
    # Calculate threshold
    trend_threshold_high = 0.7,
    trend_threshold_low = 0.3,
    spikiness_threshold_high = quantile(roa_spikiness, 0.75),
    
    # Classification
    recovery_pattern = case_when(
      roa_trend_strength >= trend_threshold_high & 
        roa_spikiness <= median(roa_spikiness) ~ "Resilient",
      roa_trend_strength <= trend_threshold_low & 
        roa_spikiness >= spikiness_threshold_high ~ "K-shaped",
      roa_trend_strength >= trend_threshold_high & 
        roa_spikiness >= spikiness_threshold_high ~ "Hidden Crisis",
      TRUE ~ "Moderate Impact"
    )
  )
```



3.  Micro-level evidence of anomalous recovery patterns across industries (ROA and liquidity) - scagnostics (Tignostics) (Lecture 8, 22-27 & Lecture 9, slide 24)

In this section we track how the firm-level relationship between liquidity (Current Ratio) and profitability (ROA) changed during the COVID-19 period.
We approach is to treat each firm as a point in the CR–ROA plane, then study the shape of each industry’s point cloud year by year.
We use scagnostic metrics to quantify those shapes, and we fit a simple linear model (ROA \~ Current Ratio) for every industry × year to extract slope and residual volatility.
We treat 2020 as the COVID year and compare values before and after 2020.

**Data and methods**

We select industries previously identified in the macro analysis (industries labelled K-shaped, Resilient, or Hidden Crisis).
For each industry × fiscal year we:

-   keep firm-level pairs (current_ratio, roa_pct);

-   calculate scagnostic metrics: outlying, stringy, striated, clumpy, sparse, monotonic, dcor;

-   fit a linear model roa_pct \~ current_ratio and record slope and residual SD;

-   compute year-to-year slope changes and count “jumps” (a jump = \|slope change\| \> 0.5);

-   compute pre/post differences in mean residual SD and mean slope (post − pre).

-   When an industry × year has too few firms (\< 3) the scagnostics or regression are not reported (NA).

**How to read the scagnostic metrics**

-   Clumpy: higher values indicate the points form separate clusters → suggests within-industry divergence (possible K-shaped pattern).

-   Outlying: higher values indicate presence of extreme firms (very strong survivors or very weak firms).

-   Stringy / Monotonic: higher values indicate a clearer linear or monotonic relationship between CR and ROA.

-   Residual SD: higher residual SD from the linear fit signals more unexplained scatter (higher micro volatility).

```{r Q4 EDA 3. Micro-level evidence -data}
selected_industries <- industry_classes |>
  filter(recovery_pattern %in% c("K-shaped", "Resilient", "Hidden Crisis")) |>
  pull(industry_group)

# Extract the companies time series data of these industries
selected_companies_ts <- germany_ts |>
  filter(industry_group %in% selected_industries)
```

```{r Q4 EDA 3. Micro-level evidence -Calculate scagnostic}
#| label: tbl-company
#| tbl-cap: "Summary statistics for industries with significant slope changes (jumps > 0). The table shows average slope, maximum absolute slope change, mean residual standard deviation, number of jumps, and differences in volatility and slope before and after the COVID-19 pandemic for three industries: Wholesale & Retail Trade, Information & Communication, and Financial & Insurance."

# Aggregate the original point set by industry × year
by_ind_year <- selected_companies_ts |>
  group_by(industry_group) |>
  index_by(fy_year) |>
  nest(data = c(company_id, roa_pct, current_ratio)) |>
  mutate(n_obs = map_int(data, ~ nrow(.x)))

# Calculate scagnostics
scag_names <- c("outlying", "stringy", "striated", "clumpy", "sparse", "monotonic", "dcor")

by_ind_year <- by_ind_year |>
  mutate(
    scags = map(data, ~ {
      df <- .x
      # When there are too few samples, return a row of tibble with the same column names and types.
      if (nrow(df) < 3) {
        return(as_tibble(set_names(as.list(rep(NA_real_, length(scag_names))), scag_names)))
      }
      res <- calc_scags(df$current_ratio, df$roa_pct, scags = scag_names)
      # Make sure to convert to atomic vectors and name them, then convert to tibble
      res_vec <- unlist(res)
      res_named <- set_names(as.numeric(res_vec[scag_names]), scag_names)
      as_tibble(as.list(res_named))
    })
  ) |>
  unnest(cols = scags)  # Expand into multiple columns

# Fit a linear regression for each industry × year and extract slope, intercept, and resid_sd
by_ind_year <- by_ind_year |>
  mutate(
    fit = map(data, ~ {
      df <- .x
      if (nrow(df) < 3 || all(is.na(df$current_ratio)) || all(is.na(df$roa_pct))) return(NULL)
      lm(roa_pct ~ current_ratio, data = df)
    }),
    slope = map_dbl(fit, ~ if (is.null(.x)) NA_real_ else coef(.x)[["current_ratio"]]),
    intercept = map_dbl(fit, ~ if (is.null(.x)) NA_real_ else coef(.x)[["(Intercept)"]]),
    resid_sd = map_dbl(fit, ~ if (is.null(.x)) NA_real_ else sd(residuals(.x), na.rm = TRUE))
  ) |>
  select(-fit)

# Take industry as a time series and calculate its evolution characteristics
industry_tignostics <- by_ind_year |>
  arrange(industry_group, fy_year) |>
  group_by(industry_group) |>
  mutate(
    slope_lag = dplyr::lag(slope),
    slope_change = slope - slope_lag,   # year-to-year change in slope
    slope_pct_change = (slope - slope_lag) / abs(slope_lag) * 100,
    resid_sd_lag = dplyr::lag(resid_sd),
    resid_sd_change = resid_sd - resid_sd_lag,
    # jumps: count of year-to-year slope changes bigger than a threshold (e.g. 0.5)
    big_slope_jump = if_else(abs(slope_change) > 0.5, 1, 0),
    jumps_cumulative = cumsum(replace_na(big_slope_jump, 0))
  ) |>
  ungroup()

# Calculate the overall volatility_shift (post - pre) before and after the epidemic (taking 2020 as the epidemic year)
vol_shift <- by_ind_year |>
  mutate(period = case_when(
    fy_year < 2020 ~ "pre",
    fy_year == 2020 ~ "covid_year",
    fy_year > 2020 ~ "post"
  )) |>
  group_by(industry_group, period) |>
  summarise(mean_resid_sd = mean(resid_sd, na.rm = TRUE),
            mean_slope = mean(slope, na.rm = TRUE),
            .groups = "drop") |>
  pivot_wider(names_from = period, values_from = c(mean_resid_sd, mean_slope)) |>
  mutate(
    vol_pre_post_diff = mean_resid_sd_post - mean_resid_sd_pre,
    slope_pre_post_diff = mean_slope_post - mean_slope_pre
  )

# Merge the results to generate the final summary
industry_summary <- industry_tignostics |>
  group_by(industry_group) |>
  summarise(
    n_years = n_distinct(fy_year),
    avg_slope = mean(slope, na.rm = TRUE),
    max_abs_slope_change = if (all(is.na(slope_change))) NA_real_ else max(abs(slope_change), na.rm = TRUE),
    mean_resid_sd = mean(resid_sd, na.rm = TRUE),
    jumps = if (all(is.na(jumps_cumulative))) 0 else max(jumps_cumulative, na.rm = TRUE),
    .groups = "drop"
  ) |>
  left_join(vol_shift |> select(industry_group, vol_pre_post_diff, slope_pre_post_diff), by = "industry_group")

kable(industry_summary |> filter(jumps > 0) |>
        select(-n_years))
```

```{r Q4 EDA 3. Micro-level evidence -scagnostic figure}
#| label: fig-scagnostic
#| fig-cap: "This figure shows how four scagnostic metrics changed over time across different industries. The red dashed line marks the year 2020 when COVID-19 pandemic occurred. We can see variations in outlying (extreme values), clumpy (clustering patterns), stringy (linear relationships), and monotonic (consistent trends) metrics before and after the pandemic."
#| warning: false

scag_long <- by_ind_year |>
  select(industry_group, fy_year, clumpy, outlying, stringy, monotonic) |>
  pivot_longer(cols = c(clumpy, outlying, stringy, monotonic),
               names_to = "scag_metric", values_to = "value")

ggplot(scag_long, aes(x = fy_year, y = value, color = industry_group)) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ scag_metric, scales = "free_y") +
  geom_vline(xintercept = 2020, linetype = "dashed", color = "red") +
  labs(
    title = "Evolution of Scagnostic Metrics over Time",
    subtitle = "Dashed line = COVID year (2020)",
    x = "Fiscal Year", y = "Scagnostic Value"
  ) +
  theme_minimal()
```


**Main results**

@tbl-company reports summary statistics for industries with at least one large slope jump (jumps \> 0).
Three industries experienced jumps: Arts, Entertainment & Recreation, Information & Communication, and Financial & Insurance.
Some industries (e.g., Construction, Accommodation & Food, Administrative & Support) have too few firms in some years and therefore produce NA for scagnostics.
@fig-scagnostic shows time series of four metrics (outlying, clumpy, stringy, monotonic) with a vertical line at 2020.

**Key patterns observed:**

| Industry | Change after 2020 | Interpretation |
|------------------------|------------------------|------------------------|
| **Financial & Insurance** | ↑ *clumpy*, ↓ *stringy*, ↑ *monotonic* | Increased fragmentation suggests a K-shaped recovery; weakened linear CR–ROA relationship; stronger monotonic trend implies more structured but divergent performance. |
| **Arts, Entertainment & Recreation** | ↓ *clumpy*, ↑ *outlying*, ↑ *monotonic* | Fewer internal clusters but more extreme performers; stronger monotonic trend indicates clearer performance–liquidity relationship post-pandemic. |
| **Information & Communication** | ↓ *clumpy*, stable *outlying*, ↑ monotonic | Reduced clustering with no significant outliers; strong monotonic rise implies more uniform and predictable firm behavior. |



**Limitations**

Results are descriptive: scagnostics indicate shape changes but do not prove causality.
We will triangulate these micro results with macro indicators (feasts features) in the final section.

5.  Summary and interpretation

The exploratory analysis demonstrates that the relationship between corporate profitability and liquidity evolved unevenly across industries during the pandemic.
More importantly, the recovery patterns were not the same across or within industries, showing clear evidence of a K-shaped recovery in some sectors.

At the macro level, trend strength and volatility analysis revealed clear segmentation: some industries such as Construction and Finance showed resilient recovery, while others, like Accommodation, Information and Recreation, exhibited K-shaped patterns with disrupted ROA trends and high volatility.

The micro-level scagnostic analysis further confirmed this divergence.
After removing several industries which have small yearly samples (Construction,\
Accommodation & Food, and Administrative & Support), the industries showing K-shaped tendencies at the macro level exhibited notable shifts in scagnostic measures.
Specifically, Financial & Insurance displayed a clear rise in clumpiness, signaling within-sector fragmentation, while Arts, Entertainment & Recreation showed growing outlying behavior, reflecting the emergence of extreme performers.
In contrast, resilient sectors maintained relatively stable and monotonic liquidity–profitability relationships throughout the period.

These results suggest that the pandemic not only altered the overall financial dynamics between liquidity and profitability but also intensified disparities within certain industries, forming a distinct K-shaped recovery pattern.

## Conclusions

According to Q1, median profitability remained broadly stable between 2018 – 2019 and 2020 – 2021, but dispersion increased.
The interesting result is that the Accommodation and Food sectors improved despite lockdowns, while the Health Care sector declined sharply, and ROA and ROE exhibited heavier lower tails.
Additionally, for Q2, balance sheets remained steady as liquidity changed little and leverage increased only slightly.
The key point is that firms did not leverage up even when credit was cheap.

According to Q3 and Q4, resilience correlates with stronger liquidity and moderate leverage, with the Health Care, Technology, and Utilities sectors showing more resilience, while Hospitality, Retail, and Manufacturing are more vulnerable.
The interesting part is that the link between liquidity and profitability weakened as many firms protected cash at the expense of near-term margins.

Overall, the pandemic’s impact on German corporate finances was moderate but uneven: typical profitability was maintained, balance sheets remained steady, and the main changes involved greater dispersion and sectoral divergence.
The crisis highlighted existing gaps, strengthening resilient sectors and revealing vulnerabilities in more leveraged, cyclical industries.

## Resources

The materials used for this report are:

Wickham H, Averick M, Bryan J, Chang W, McGowan LD, François R, Grolemund G, Hayes A, Henry L, Hester J, Kuhn M, Pedersen TL, Miller E, Bache SM, Müller K, Ooms J, Robinson D, Seidel DP, Spinu V, Takahashi K, Vaughan D, Wilke C, Woo K, Yutani H (2019).
"Welcome to the tidyverse." *Journal of Open Source Software*, *4*(43), 1686.
[doi:10.21105/joss.01686](https://doi.org/10.21105/joss.01686).

The links to my use of ChatGPT and other AI for help on this assignment are:

Q2: Bhuvana - <https://chatgpt.com/share/68f4378d-bb24-8003-8471-ed2147a8fab3> Used for formatting assistance, debugging R code, and improving interpretation flow.

Q4: Jingyi

-   <https://chatgpt.com/share/68f3046d-1ad8-800d-bc11-1b005a9e600f> helped to implement animated graphics and optimize content.

-   <https://chatgpt.com/share/68f304c8-a088-800d-be54-939dfcd7e278> helped to evaluate whether the questions are answered adequately and where to make revisions.

-   <https://chatgpt.com/share/68f305ad-f15c-800d-b724-fc917c1a93ef> helped to deal with the problem that package ‘tignostics’ is not available for this version of R and how to use scagnostics.
