---
title: "ETC5521 Diving Deeper into Data Exploration: Project Part 1"
author: "P1-wallaby"
date: "Oct 6, 2025"
quarto-required: ">=1.8.0"
format:
    html: 
        output-file: report.html
        css: "assignment.css"
        embed-resources: true
---

```{r}
#| echo: false
# Set up chunk for all slides
knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 4,
  fig.align = "center",
  out.width = "100%",
  code.line.numbers = FALSE,
  fig.retina = 4,
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = FALSE,
  dev.args = list(pointsize = 11)
)
```

```{r}
#| echo: false
# Load libraries
library(tidyverse)
library(conflicted)

conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::select)
conflicts_prefer(dplyr::slice)

# Plot options and themes
options(
  digits = 2,
  width = 60,
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis",
  ggplot2.discrete.colour = c("#D55E00", "#0072B2", "#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442"),
  ggplot2.discrete.fill = c("#D55E00", "#0072B2", "#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442")
)

theme_set(theme_bw(base_size = 14) +
   theme(
     aspect.ratio = 1,
     plot.background = element_rect(fill = 'transparent', colour = NA),
     plot.title.position = "plot",
     plot.title = element_text(size = 24),
     panel.background = element_rect(fill = 'transparent', colour = NA),
     legend.background = element_rect(fill = 'transparent', colour = NA),
     legend.key = element_rect(fill = 'transparent', colour = NA)
   )
)
```

```{r load-library}
library(dplyr)
library(stringr)
library(janitor)
library(skimr)
```


## Introduction 

The COVID-19 pandemic caused significant disruptions across global markets, testing the financial resilience of firms in every sector. In Germany, the first confirmed case was recorded on 27 January 2020. Nationwide contact restrictions began on 22 March 2020. Most nationwide measures were lifted between 20 March and 2 April 2022, and the final federal provisions expired on 7 April 2023.

This project examines how German companies responded to these shocks during 2020 – 2021, using financial data from the Bureau van Dijk OSIRIS Database — a reliable global source of harmonised corporate information. The dataset available for this study covers 2018 – 2021. The period from 2018–2019 provides the pre-pandemic baseline, while 2020 – 2021 captures the pandemic period. Data for 2022 are not available in this extract.

Our analysis focuses on how industries differed in their financial outcomes during this period, investigating changes in profitability, leverage, liquidity, and firm size. By identifying which sectors showed unexpected resilience or vulnerability, we aim to understand how structural and strategic differences shaped recovery from the pandemic’s economic impacts. All analytical procedures follow the ETC5521 workflow, ensuring reproducibility, transparency, and alignment with the course framework for exploratory data analysis.


## Data description


```{r read-data}
#define the folder
data_directory <- "osiris"

#filter only germany
germany_file_paths <- list.files(
  data_directory,
  pattern = "^osiris_?Germany_(2018|2019|2020|2021)\\.rda$",
  full.names = TRUE,
  ignore.case = TRUE
)

#name the list by file stems
germany_file_stems <- tools::file_path_sans_ext(basename(germany_file_paths))
osiris_germany <- setNames(vector("list", length(germany_file_paths)), germany_file_stems)

#load each file
for (file_index in seq_along(germany_file_paths)) {
  temporary_environment <- new.env(parent = emptyenv())
  load(germany_file_paths[file_index], envir = temporary_environment)
  loaded_objects <- as.list(temporary_environment)

  #If the file holds one object, store that object; otherwise store the sub-list
  osiris_germany[[file_index]] <- if (length(loaded_objects) == 1) loaded_objects[[1]] else loaded_objects
}

#quick check
str(osiris_germany, max.level = 1)
```

```{r read-data-2}
#build a single data frame with a 'year' column taken from the filename
combined_germany <- {
  germany_parts <- vector("list", length(osiris_germany))
  names(germany_parts) <- names(osiris_germany)

  for (file_stem in names(osiris_germany)) {
    data_frame_from_file <- osiris_germany[[file_stem]]
    extracted_year <- as.integer(str_extract(file_stem, "\\d{4}"))

    germany_parts[[file_stem]] <- mutate(data_frame_from_file, year = extracted_year)
  }

  #bind and sort by year (optional)
  bind_rows(germany_parts) |>
    arrange(year)
}
```

```{r data-cleaning}
filtered_germany <- combined_germany |>
  clean_names() |>
  
  rename(
    #IDs / scope (used in all questions)
    company_name = company_name_name, #label
    company_id = name_id, #unique firm id
    acct_year = year, #reporting year
    country = country_country,
    city = city_city_city, #city/HQ
    consolidation_code = consolidation_code_consol_code, #consolidated vs standalone
    
    #industry codes (Q1, Q2, Q4)
    nace_code = nace_rev_1_core_code_cnacecd, #EU industry code (primary)
    icb_code = industrial_classification_benchmark_icb, #ICB classification (fallback)
    sic_code = us_sic_core_code_csicuscde, #SIC code (fallback)
    
    #scale / balance sheet (Q2, Q3)
    total_assets_eur = total_assets_data13077, #base for ratios
    total_liabilities_eur = total_liabilities_and_debt_data14022, #debt measure
    total_equity_eur = total_shareholders_equity_data14041, #equity capital
    
    #revenue & profit (Q1, Q3)
    total_revenue_eur = total_revenues_data13004, #denominator for margins
    net_sales_eur = net_sales_data13002, #sales denominator
    gross_sales_eur = gross_sales_data13000, #top-line revenue
    net_income_eur = net_income_starting_line_data15500, #used to calculate ROA/ROE
    
    #profitability ratios (Q1, Q4)
    ebit_margin_pct = ebit_margin_percent_data31055, #profitability metric (EBIT %)
    ebitda_margin_pct = ebitda_margin_percent_data31060, #profitability metric (EBITDA %)
    roa_pct = return_on_total_assets_percent_data31015, #return on assets
    roe_pct = roe_percent_data31065, #return on equity
    
    #efficiency (supporting for Q1/Q4)
    net_assets_turnover = net_assets_turnover_data31225, #asset efficiency
    stock_turnover = stock_turnover_data31220, #inventory management efficiency
    
    #leverage, solvency, and liquidity (Q2)
    solvency_pct = solvency_ratio_percent_data31310, #solvency strength
    gearing_pct = gearing_percent_data31315, #leverage level
    current_ratio = current_ratio_data31105, #liquidity ratio (short-term)
    liquidity_ratio = liquidity_ratio_data31110, #alternate liquidity ratio
    shareholders_liq_pct = shareholders_liquidity_ratio_data31305, #shareholder liquidity
    interest_cover = interest_cover_data31115 #interest
  ) |>
  
  mutate(
    fy_end_date = ymd(as.character(company_fiscal_year_end_date_closdate)),
    fy_year     = year(fy_end_date)
  ) |>
  
  #select only relevant columns to answer the questions
  select(
    company_id, company_name, fy_end_date, fy_year, acct_year, country, city, consolidation_code,
    nace_code, icb_code, sic_code,
    total_assets_eur, total_liabilities_eur, total_equity_eur,
    total_revenue_eur, net_sales_eur, gross_sales_eur, net_income_eur,
    ebit_margin_pct, ebitda_margin_pct, roa_pct, roe_pct,
    net_assets_turnover, stock_turnover,
    solvency_pct, gearing_pct, current_ratio, liquidity_ratio,
    shareholders_liq_pct, interest_cover
  )
  
```

```{r table-checking}

num_cols <- c(
  "total_assets_eur","total_liabilities_eur","total_equity_eur",
  "total_revenue_eur","net_sales_eur","gross_sales_eur","net_income_eur",
  "ebit_margin_pct","ebitda_margin_pct","roa_pct","roe_pct",
  "net_assets_turnover","stock_turnover",
  "solvency_pct","gearing_pct","current_ratio","liquidity_ratio",
  "shareholders_liq_pct","interest_cover"
)

#symbol formatting/coerce numerics
filtered_germany <- filtered_germany |>
  mutate(
    across(all_of(num_cols) & where(is.character), readr::parse_number),
    country = stringr::str_to_title(country),
    city    = stringr::str_to_title(city)
  )

#missingness report check (overall and by year)
missing_overall <- filtered_germany |>
  summarise(across(everything(), ~ mean(is.na(.)))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "missing_rate")

missing_by_year <- filtered_germany |>
  group_by(fy_year) |>
  summarise(across(everything(), ~ mean(is.na(.)))) |>
  pivot_longer(-fy_year, names_to = "variable", values_to = "missing_rate")

#sanity checks (optional but recommended)
dup_firm_year <- filtered_germany |>
  count(company_id, fy_year) |>
  filter(n > 1)

impossible_assets <- filtered_germany |>
  filter(!is.na(total_assets_eur) & total_assets_eur < 0)

consol_levels <- filtered_germany |>
  count(consolidation_code, sort = TRUE)

```

```{r NACE-column}
#clean 2-digit NACE as integer
filtered_germany <- filtered_germany |>
  mutate(
    nace_num2 = suppressWarnings(as.integer(str_sub(readr::parse_number(as.character(nace_code)), 1, 2)))
  )

#map NACE divisions to broad section names
industry_from_nace <- function(x){
  case_when(
    x %in% 1:3 ~ "Agriculture, Forestry & Fishing",
    x %in% 5:9 ~ "Mining & Quarrying",
    x %in% 10:33 | x %in% 15:37 ~ "Manufacturing",
    x %in% 35 ~ "Electricity, Gas, Steam",
    x %in% 36:39 ~ "Water Supply & Waste",
    x %in% 41:43 ~ "Construction",
    x %in% 45:47 ~ "Wholesale & Retail Trade",
    x %in% 49:53 ~ "Transport & Storage",
    x %in% 55:56 ~ "Accommodation & Food",
    x %in% 58:63 ~ "Information & Communication",
    x %in% 64:66 ~ "Financial & Insurance",
    x %in% 68 ~ "Real Estate",
    x %in% 69:75 ~ "Professional, Scientific & Technical",
    x %in% 77:82 ~ "Administrative & Support",
    x %in% 84 ~ "Public Administration",
    x %in% 85 ~ "Education",
    x %in% 86:88 ~ "Human Health & Social Work",
    x %in% 90:93 ~ "Arts, Entertainment & Recreation",
    x %in% 94:96 ~ "Other Service Activities",
    x %in% 97:98 ~ "Household Activities",
    x %in% 99 ~ "Extraterritorial Organizations",
    TRUE ~ NA_character_
  )
}

#ICB top-level sector labels
icb_map <- c(
  "0001" = "Oil & Gas",
  "1000" = "Basic Materials",
  "2000" = "Industrials",
  "3000" = "Consumer Goods",
  "4000" = "Health Care",
  "5000" = "Consumer Services",
  "6000" = "Telecommunications",
  "7000" = "Utilities",
  "8000" = "Financials",
  "9000" = "Technology"
)

#normalize ICB to a 4-digit “bucket” (example: 2573 to 2000)
normalize_icb <- function(x){
  x_chr <- str_extract(as.character(x), "\\d+")
  ifelse(is.na(x_chr), NA_character_,
         sprintf("%04d", as.integer(floor(as.numeric(x_chr) / 1000) * 1000)))
}

filtered_germany <- filtered_germany |>
  mutate(
    industry_nace = industry_from_nace(nace_num2),
    icb_bucket    = normalize_icb(icb_code),
    industry_icb  = icb_map[icb_bucket],
    industry_group = coalesce(industry_nace, industry_icb, "Other / Unmapped")
  )

#check
unmapped_sample <- filtered_germany |>
  filter(industry_group == "Other / Unmapped") |>
  select(company_id, fy_year, nace_code, icb_code) |>
  head(15)

#factor order for plots
ordered_levels <- c(
  "Agriculture, Forestry & Fishing","Mining & Quarrying","Manufacturing",
  "Electricity, Gas, Steam","Water Supply & Waste","Construction",
  "Wholesale & Retail Trade","Transport & Storage","Accommodation & Food",
  "Information & Communication","Financial & Insurance","Real Estate",
  "Professional, Scientific & Technical","Administrative & Support",
  "Public Administration","Education","Human Health & Social Work",
  "Arts, Entertainment & Recreation","Other Service Activities",
  "Household Activities","Extraterritorial Organizations",
  "Oil & Gas","Basic Materials","Industrials","Consumer Goods","Health Care",
  "Consumer Services","Telecommunications","Utilities","Financials","Technology",
  "Other / Unmapped"
)
filtered_germany <- filtered_germany |>
  mutate(industry_group = factor(industry_group, levels = ordered_levels))
```

### Research Question

**Which German industries showed unexpected financial resilience or vulnerability from 2020 to 2022, and how were these outcomes shaped by profitability, leverage, liquidity, and firm size?**

During 2020–2022, industries such as healthcare, information technology, and utilities demonstrated notable financial resilience, supported by strong liquidity and stable profitability. In contrast, sectors like hospitality, transportation, and manufacturing showed greater vulnerability due to higher leverage and lower profit margins. Overall, firm size and liquidity emerged as key buffers against financial stress, while high debt levels amplified vulnerability across more capital-intensive industries.


## Conclusions



## Resources

The materials used for this report are:

Wickham H, Averick M, Bryan J, Chang W, McGowan LD, François R,
  Grolemund G, Hayes A, Henry L, Hester J, Kuhn M, Pedersen TL,
  Miller E, Bache SM, Müller K, Ooms J, Robinson D, Seidel DP, Spinu
  V, Takahashi K, Vaughan D, Wilke C, Woo K, Yutani H (2019).
  "Welcome to the tidyverse." _Journal of Open Source Software_,
  *4*(43), 1686. [doi:10.21105/joss.01686](https://doi.org/10.21105/joss.01686).

The links to my use of ChatGPT for help on this assignment are:

- [https://chatgpt.com/share/68857a22-debc-8001-bfcc-10552befba4c]() helped to XXX.

