---
title: "ETC5521 Diving Deeper into Data Exploration: Project Part 1"
author: "P1-wallaby"
date: "Oct 6, 2025"
quarto-required: ">=1.8.0"
format:
    html: 
        output-file: report.html
        css: "assignment.css"
        embed-resources: true
---

```{r}
#| echo: false
# Set up chunk for all slides
knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 4,
  fig.align = "center",
  out.width = "100%",
  code.line.numbers = FALSE,
  fig.retina = 4,
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = FALSE,
  dev.args = list(pointsize = 11)
)
```

```{r}
#| echo: false
# Load libraries
library(tidyverse)
library(conflicted)

conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::select)
conflicts_prefer(dplyr::slice)

# Plot options and themes
options(
  digits = 2,
  width = 60,
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis",
  ggplot2.discrete.colour = c("#D55E00", "#0072B2", "#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442"),
  ggplot2.discrete.fill = c("#D55E00", "#0072B2", "#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442")
)

theme_set(theme_bw(base_size = 14) +
   theme(
     aspect.ratio = 1,
     plot.background = element_rect(fill = 'transparent', colour = NA),
     plot.title.position = "plot",
     plot.title = element_text(size = 24),
     panel.background = element_rect(fill = 'transparent', colour = NA),
     legend.background = element_rect(fill = 'transparent', colour = NA),
     legend.key = element_rect(fill = 'transparent', colour = NA)
   )
)
```

```{r load-library}
library(dplyr)
library(stringr)
library(janitor)
library(skimr)
library(tsibble)
library(knitr)
library(naniar)
library(patchwork)
library(plotly)
library(tsibbletalk)
library(feasts)
library(broom)
library(purrr)
library(cassowaryr)
library(gganimate)
library(gifski)
```



## Introduction 

The COVID-19 pandemic caused significant disruptions across global markets, testing the financial resilience of firms in every sector. In Germany, the first confirmed case was recorded on 27 January 2020. Nationwide contact restrictions began on 22 March 2020. Most nationwide measures were lifted between 20 March and 2 April 2022, and the final federal provisions expired on 7 April 2023.

This project examines how German companies responded to these shocks during 2020 – 2021, using financial data from the Bureau van Dijk OSIRIS Database — a reliable global source of harmonised corporate information. The dataset available for this study covers 2018 – 2021. The period from 2018–2019 provides the pre-pandemic baseline, while 2020 – 2021 captures the pandemic period. Data for 2022 are not available in this extract.

Our analysis focuses on how industries differed in their financial outcomes during this period, investigating changes in profitability, leverage, liquidity, and firm size. By identifying which sectors showed unexpected resilience or vulnerability, we aim to understand how structural and strategic differences shaped recovery from the pandemic’s economic impacts. All analytical procedures follow the ETC5521 workflow, ensuring reproducibility, transparency, and alignment with the course framework for exploratory data analysis.


## Data description


```{r read-data}
#define the folder
data_directory <- "osiris"

#filter only germany
germany_file_paths <- list.files(
  data_directory,
  pattern = "^osiris_?Germany_(2018|2019|2020|2021)\\.rda$",
  full.names = TRUE,
  ignore.case = TRUE
)

#name the list by file stems
germany_file_stems <- tools::file_path_sans_ext(basename(germany_file_paths))
osiris_germany <- setNames(vector("list", length(germany_file_paths)), germany_file_stems)

#load each file
for (file_index in seq_along(germany_file_paths)) {
  temporary_environment <- new.env(parent = emptyenv())
  load(germany_file_paths[file_index], envir = temporary_environment)
  loaded_objects <- as.list(temporary_environment)

  #If the file holds one object, store that object; otherwise store the sub-list
  osiris_germany[[file_index]] <- if (length(loaded_objects) == 1) loaded_objects[[1]] else loaded_objects
}

#quick check
str(osiris_germany, max.level = 1)
```

```{r read-data-2}
#build a single data frame with a 'year' column taken from the filename
combined_germany <- {
  germany_parts <- vector("list", length(osiris_germany))
  names(germany_parts) <- names(osiris_germany)

  for (file_stem in names(osiris_germany)) {
    data_frame_from_file <- osiris_germany[[file_stem]]
    extracted_year <- as.integer(str_extract(file_stem, "\\d{4}"))

    germany_parts[[file_stem]] <- mutate(data_frame_from_file, year = extracted_year)
  }

  #bind and sort by year (optional)
  bind_rows(germany_parts) |>
    arrange(year)
}
```

```{r data-cleaning}
filtered_germany <- combined_germany |>
  clean_names() |>
  
  rename(
    #IDs / scope (used in all questions)
    company_name = company_name_name, #label
    company_id = name_id, #unique firm id
    acct_year = year, #reporting year
    country = country_country,
    city = city_city_city, #city/HQ
    consolidation_code = consolidation_code_consol_code, #consolidated vs standalone
    
    #industry codes (Q1, Q2, Q4)
    nace_code = nace_rev_1_core_code_cnacecd, #EU industry code (primary)
    icb_code = industrial_classification_benchmark_icb, #ICB classification (fallback)
    sic_code = us_sic_core_code_csicuscde, #SIC code (fallback)
    
    #scale / balance sheet (Q2, Q3)
    total_assets_eur = total_assets_data13077, #base for ratios
    total_liabilities_eur = total_liabilities_and_debt_data14022, #debt measure
    total_equity_eur = total_shareholders_equity_data14041, #equity capital
    
    #revenue & profit (Q1, Q3)
    total_revenue_eur = total_revenues_data13004, #denominator for margins
    net_sales_eur = net_sales_data13002, #sales denominator
    gross_sales_eur = gross_sales_data13000, #top-line revenue
    net_income_eur = net_income_starting_line_data15500, #used to calculate ROA/ROE
    
    #profitability ratios (Q1, Q4)
    ebit_margin_pct = ebit_margin_percent_data31055, #profitability metric (EBIT %)
    ebitda_margin_pct = ebitda_margin_percent_data31060, #profitability metric (EBITDA %)
    roa_pct = return_on_total_assets_percent_data31015, #return on assets
    roe_pct = roe_percent_data31065, #return on equity
    
    #efficiency (supporting for Q1/Q4)
    net_assets_turnover = net_assets_turnover_data31225, #asset efficiency
    stock_turnover = stock_turnover_data31220, #inventory management efficiency
    
    #leverage, solvency, and liquidity (Q2)
    solvency_pct = solvency_ratio_percent_data31310, #solvency strength
    gearing_pct = gearing_percent_data31315, #leverage level
    current_ratio = current_ratio_data31105, #liquidity ratio (short-term)
    liquidity_ratio = liquidity_ratio_data31110, #alternate liquidity ratio
    shareholders_liq_pct = shareholders_liquidity_ratio_data31305, #shareholder liquidity
    interest_cover = interest_cover_data31115 #interest
  ) |>
  
  mutate(
    fy_end_date = ymd(as.character(company_fiscal_year_end_date_closdate)),
    fy_year     = year(fy_end_date)
  ) |>
  
  #select only relevant columns to answer the questions
  select(
    company_id, company_name, fy_end_date, fy_year, acct_year, country, city, consolidation_code,
    nace_code, icb_code, sic_code,
    total_assets_eur, total_liabilities_eur, total_equity_eur,
    total_revenue_eur, net_sales_eur, gross_sales_eur, net_income_eur,
    ebit_margin_pct, ebitda_margin_pct, roa_pct, roe_pct,
    net_assets_turnover, stock_turnover,
    solvency_pct, gearing_pct, current_ratio, liquidity_ratio,
    shareholders_liq_pct, interest_cover
  )
  
```

```{r table-checking}

num_cols <- c(
  "total_assets_eur","total_liabilities_eur","total_equity_eur",
  "total_revenue_eur","net_sales_eur","gross_sales_eur","net_income_eur",
  "ebit_margin_pct","ebitda_margin_pct","roa_pct","roe_pct",
  "net_assets_turnover","stock_turnover",
  "solvency_pct","gearing_pct","current_ratio","liquidity_ratio",
  "shareholders_liq_pct","interest_cover"
)

#symbol formatting/coerce numerics
filtered_germany <- filtered_germany |>
  mutate(
    across(all_of(num_cols) & where(is.character), readr::parse_number),
    country = stringr::str_to_title(country),
    city    = stringr::str_to_title(city)
  )

#missingness report check (overall and by year)
missing_overall <- filtered_germany |>
  summarise(across(everything(), ~ mean(is.na(.)))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "missing_rate")

missing_by_year <- filtered_germany |>
  group_by(fy_year) |>
  summarise(across(everything(), ~ mean(is.na(.)))) |>
  pivot_longer(-fy_year, names_to = "variable", values_to = "missing_rate")

#sanity checks (optional but recommended)
dup_firm_year <- filtered_germany |>
  count(company_id, fy_year, acct_year) |>
  filter(n > 1)

impossible_assets <- filtered_germany |>
  filter(!is.na(total_assets_eur) & total_assets_eur < 0)

consol_levels <- filtered_germany |>
  count(consolidation_code, sort = TRUE)

```

```{r NACE-column}
#clean 2-digit NACE as integer
filtered_germany <- filtered_germany |>
  mutate(
    nace_num2 = suppressWarnings(as.integer(str_sub(readr::parse_number(as.character(nace_code)), 1, 2)))
  )

#map NACE divisions to broad section names
industry_from_nace <- function(x){
  case_when(
    x %in% 1:3 ~ "Agriculture, Forestry & Fishing",
    x %in% 5:9 ~ "Mining & Quarrying",
    x %in% 10:33 | x %in% 15:37 ~ "Manufacturing",
    x %in% 35 ~ "Electricity, Gas, Steam",
    x %in% 36:39 ~ "Water Supply & Waste",
    x %in% 41:43 ~ "Construction",
    x %in% 45:47 ~ "Wholesale & Retail Trade",
    x %in% 49:53 ~ "Transport & Storage",
    x %in% 55:56 ~ "Accommodation & Food",
    x %in% 58:63 ~ "Information & Communication",
    x %in% 64:66 ~ "Financial & Insurance",
    x %in% 68 ~ "Real Estate",
    x %in% 69:75 ~ "Professional, Scientific & Technical",
    x %in% 77:82 ~ "Administrative & Support",
    x %in% 84 ~ "Public Administration",
    x %in% 85 ~ "Education",
    x %in% 86:88 ~ "Human Health & Social Work",
    x %in% 90:93 ~ "Arts, Entertainment & Recreation",
    x %in% 94:96 ~ "Other Service Activities",
    x %in% 97:98 ~ "Household Activities",
    x %in% 99 ~ "Extraterritorial Organizations",
    TRUE ~ NA_character_
  )
}

#ICB top-level sector labels
icb_map <- c(
  "0001" = "Oil & Gas",
  "1000" = "Basic Materials",
  "2000" = "Industrials",
  "3000" = "Consumer Goods",
  "4000" = "Health Care",
  "5000" = "Consumer Services",
  "6000" = "Telecommunications",
  "7000" = "Utilities",
  "8000" = "Financials",
  "9000" = "Technology"
)

#normalize ICB to a 4-digit “bucket” (example: 2573 to 2000)
normalize_icb <- function(x){
  x_chr <- str_extract(as.character(x), "\\d+")
  ifelse(is.na(x_chr), NA_character_,
         sprintf("%04d", as.integer(floor(as.numeric(x_chr) / 1000) * 1000)))
}

filtered_germany <- filtered_germany |>
  mutate(
    industry_nace = industry_from_nace(nace_num2),
    icb_bucket    = normalize_icb(icb_code),
    industry_icb  = icb_map[icb_bucket],
    industry_group = coalesce(industry_nace, industry_icb, "Other / Unmapped")
  )

#check
unmapped_sample <- filtered_germany |>
  filter(industry_group == "Other / Unmapped") |>
  select(company_id, fy_year, acct_year, nace_code, icb_code) |>
  head(15)

#factor order for plots
ordered_levels <- c(
  "Agriculture, Forestry & Fishing","Mining & Quarrying","Manufacturing",
  "Electricity, Gas, Steam","Water Supply & Waste","Construction",
  "Wholesale & Retail Trade","Transport & Storage","Accommodation & Food",
  "Information & Communication","Financial & Insurance","Real Estate",
  "Professional, Scientific & Technical","Administrative & Support",
  "Public Administration","Education","Human Health & Social Work",
  "Arts, Entertainment & Recreation","Other Service Activities",
  "Household Activities","Extraterritorial Organizations",
  "Oil & Gas","Basic Materials","Industrials","Consumer Goods","Health Care",
  "Consumer Services","Telecommunications","Utilities","Financials","Technology",
  "Other / Unmapped"
)
filtered_germany <- filtered_germany |>
  mutate(industry_group = factor(industry_group, levels = ordered_levels))
```


```{r}
write.csv(filtered_germany, file = "filtered_germany.csv", row.names = FALSE)
```


### Research Question

### Question 4

**How did the fundamental relationship between corporate liquidity and profitability evolve and fracture within industries during the pandemic? Are there anomalous recovery patterns (e.g., K-shaped) within specific industries?**

During 2020–2022, industries such as healthcare, information technology, and utilities demonstrated notable financial resilience, supported by strong liquidity and stable profitability. In contrast, sectors like hospitality, transportation, and manufacturing showed greater vulnerability due to higher leverage and lower profit margins. Overall, firm size and liquidity emerged as key buffers against financial stress, while high debt levels amplified vulnerability across more capital-intensive industries.


#### How did the fundamental relationship between corporate liquidity and profitability evolve and fracture within industries during the pandemic?				


The COVID-19 pandemic from 2020 to 2022 severely disrupted global business activity, placing significant stress on corporate financial structures. One critical dimension of this disruption lies in how firms managed their liquidity—the ability to meet short-term obligations—while maintaining profitability during the crisis. Financial theory typically suggests that maintaining sufficient liquidity supports profitability by ensuring operational stability and investment flexibility. However, during the pandemic, this fundamental relationship may have fractured across industries due to uneven economic shocks and varying policy responses.

This study investigates how the relationship between corporate liquidity and profitability evolved and diverged across industries during the pandemic period. Using data from the OSIRIS database for Germany between 2017 and 2022, I focus on two key indicators: 

- **Return on Assets (ROA)** as a measure of profitability 

- **Current Ratio** as a measure of liquidity. 

I aim to uncover distinct recovery patterns—such as “K-shaped” rebounds, resilient sectors, and hidden crises—that reveal how the financial health of different industries adapted or deteriorated during the pandemic.


 

##### IDA

1. Data Cleaning on Consolidation Codes

The raw data from Orbis contained financial statements under different consolidation codes (C1, C2, C\*, U1, U2). To ensure consistency and comparability, I prioritized consolidated accounts which represent the entire economic entity. For each firm-year observation, I retained the highest-priority available consolidated statement following the order: C1 > C2 > C*. Unconsolidated statements (U1, U2) were excluded from the analysis.


```{r}
germany_q4 <- filtered_germany %>%
  mutate(
    consolidation_priority = case_when(
      consolidation_code == "C1" ~ 1,  # highest priority
      consolidation_code == "C2" ~ 2,
      consolidation_code == "C*" ~ 3,
      consolidation_code == "U1" ~ 99, # Very low priority, about to be filtered
      consolidation_code == "U2" ~ 99,
      TRUE ~ 99  # Handle any unexpected code
    )
  ) %>%
  
  # Group by company and year, and keep only the highest priority records
  group_by(company_id, fy_year) %>%
  arrange(consolidation_priority, .by_group = TRUE) %>%
  slice(1) %>% # Take the first row of each group, which has the highest priority
  ungroup() %>%
  # Filter out the remaining non-consolidated records
  filter(consolidation_priority %in% c(1, 2, 3))
```

2. Select the columns to be analyzed

```{r}
germany_q4 <- germany_q4 %>%
  select(company_id, fy_year, roa_pct, current_ratio, industry_group)
```

3. Convert data into a tsibble object using the tsibble package (Lecture 9, slide 5 & Reading week 9)


Each company was defined as a key and the financial year as the index. Basic screening was performed to check for missing values, outliers, and duplicates.


```{r}
germany_ts <- germany_q4 %>%
  as_tsibble(key = company_id, index = fy_year) %>%
  arrange(company_id, fy_year)
```


4. Summary

The trends in ROA and Current Ratio from 2017 to 2021 reveal a distinct pattern, potentially linked to the pandemic. ROA showed significant volatility: it was stable pre-pandemic, fell to a negative value in 2020, but then recovered strongly to a peak of 4.87% in 2021. In contrast, the Current Ratio remained relatively stable throughout the period, indicating consistent short-term liquidity despite the economic disruption. This suggests that while profitability was heavily impacted, the company's ability to meet its short-term obligations was maintained.

```{r}
#| label: tbl-summary
#| tbl-cap: "Trends in Mean Return on Assets (ROA) and Current Ratio from Pre-Pandemic to Post-Pandemic Years (2017-2021)"

kable(germany_ts %>%
  index_by(fy_year) %>%
  summarise(mean_roa = mean(roa_pct, na.rm = TRUE),
            mean_cr = mean(current_ratio, na.rm = TRUE)))
```

##### EDA

###### 1.. Data preparation

**Check for missing value (Tutorial 8)**


```{r}
#| label: fig-missing_plots
#| fig-cap: "The plots of Missing Value Distribution"

# Check for missing ROA and CR
miss_var <- 
  gg_miss_var(germany_ts, show_pct = TRUE) + 
  labs(title = "Distribution of Missing Values", 
       subtitle = "For all variables")

# Check for missing values by industry group
miss_industry <- germany_ts |> 
  group_by(industry_group) |> 
  gg_miss_var(show_pct = TRUE) + 
  labs(title = "Distribution of Missing Values", 
       subtitle = "By industry group")


# Missing case pattern
miss_case <- gg_miss_case(germany_ts, show_pct = TRUE) + 
  labs(title = "Distribution of Missing Case Pattern")
  
miss_var / miss_industry / miss_case
```


```{r}
#| label: fig-missing_clu
#| fig-cap: "The plot of clustering to visualize missing patterns"

m3 <- vis_miss(germany_ts, cluster = TRUE, sort_miss = TRUE) +
  theme(axis.text.x = element_text(size = 8, angle = 45, hjust = 1),
        axis.text.y = element_text(size = 8))

m3
```

The initial examination of missing values revealed a relatively low proportion of incomplete data in the dataset. As shown in @fig-missing_clu, the missing value analysis showed that only 0.7% of the total data was missing, while 99.3% of observations were complete. ROA contained between 1% and 1.5% missing values across the dataset, while the current ratio had a slightly higher proportion of missing data, ranging from 2% to 2.5% (@fig-missing_plots).


**Address missingness issues**

The 'Other / Unmapped' category was removed from the study as it comprises companies without clear industry classification.


```{r}
germany_ts <- germany_ts %>%
  as_tibble() %>%   # Convert to a normal tibble
  # Remove observations with missing ROA or CR
  filter(!is.na(roa_pct) & !is.na(current_ratio)) %>%
   # Remove industries that don't have data in 2019-2021
  group_by(industry_group) %>%
  filter(all(2019:2021 %in% fy_year)) %>%
  filter(industry_group != "Other / Unmapped") %>%
  ungroup() %>%
  # Convert back to tsibble format
  as_tsibble(key = company_id, index = fy_year) %>%
  arrange(company_id, fy_year)
```

###### 2. Industry-level trends of ROA and liquidity (2017–2021) (Lecture 9, slide 24-26, feasts and tsibbletalk)  

```{r}
# Calculate the average ROA and current ratio for each industry each year
industry_ts <- germany_ts %>%
  group_by(industry_group) %>%
  index_by(fy_year) %>%
  summarise(
    industry_roa = mean(roa_pct, na.rm = TRUE),
    industry_cr = mean(current_ratio, na.rm = TRUE),
    n_companies = n()  
  )

```


```{r}
#| label: fig-ROA_ploty
#| fig-cap: "The interactive plot of ROA trends by industry (2017-2021)"

# ROA
roa_interactive <- industry_ts %>%
  ggplot(aes(x = fy_year, y = industry_roa, group = industry_group)) +
  geom_line(aes(color = industry_group), size = 0.8, alpha = 0.7) +
  geom_vline(xintercept = 2020, linetype = "dashed", color = "red", size = 1) +
  labs(title = "Interactive ROA Trends by Industry (2017-2021)<br><sub>Vertical line indicates pandemic start (2020) - Hover for details",
       x = "Year", y = "ROA (%)") +
  theme_minimal() +
  theme(legend.position = "none") 

# Plotly interactive charts -ROA
roa_plotly <- 
  ggplotly(roa_interactive, tooltip = c("industry_group", "fy_year", "industry_roa")) %>%
  plotly::layout(hoverlabel = list(bgcolor = "white", font = list(size = 12)))

roa_plotly
```



```{r}
#| label: fig-CR_ploty
#| tbl-cap: "The interactive plot of Current Ratio trends by industry (2017-2021)"

# Current ratio
cr_interactive <- industry_ts %>%
  ggplot(aes(x = fy_year, y = industry_cr, group = industry_group)) +
  geom_line(aes(color = industry_group), size = 0.8, alpha = 0.7) +
  geom_vline(xintercept = 2020, linetype = "dashed", color = "red", size = 1) +
  labs(title = "Interactive Current Ratio Trends by Industry (2017-2021) <br><sub>Vertical line indicates pandemic start (2020) - Hover for details",
       x = "Year", y = "Current Ratio") +
  theme_minimal() +
  theme(legend.position = "none")


cr_plotly <- 
  ggplotly(cr_interactive, tooltip = c("industry_group", "fy_year", "industry_cr")) %>%
  plotly::layout(hoverlabel = list(bgcolor = "white", font = list(size = 12)))


cr_plotly
```


The interactive ROA trends (@fig-ROA_ploty) reveal distinct pandemic responses across industries:

- Accommodation & Food surged temporarily in 2020 but declined again in 2021.

- Industrials showed a persistent decline that accelerated post-2020, while Information & Communication rebounded sharply after an initial dip. 
-
- Manufacturing and Transport & Storage transformed from negative to positive ROA during the pandemic period. 

- Financial & Insurance maintained the highest ROA with steady growth throughout, demonstrating remarkable sector resilience.

Based on the Current Ratio trends (@fig-CR_ploty), most industries demonstrated relative stability throughout the period. However, two sectors exhibited notable deviations: 

- Consumer Services peaked at 8.14 in 2020 then plummeted to 0.21.

- Arts, Entertainment & Recreation showed a significant upward trend in liquidity, which appears counterintuitive given pandemic constraints. 

- Financials experienced a substantial decline from 7.8 in 2020 to 4.2 in 2021, indicating a sharp contraction in liquidity buffers during the post-pandemic recovery phase.

-  Construction maintained exceptional stability throughout the period.

###### 3. Macro-level evidence: Industry recovery patterns (ROA vs CR) - feasts (Lecture 9, slide 24-26, feasts and tsibbletalk)  

Now, I will analyze the fundamental relationship between corporate liquidity and profitability across industries during the pandemic period, with particular focus on identifying anomalous recovery patterns.

```{r}
# Calculating time series features
industry_features <- industry_ts %>%
  features(industry_roa, feat_stl) %>%
  rename_with(~ paste0("roa_", .), -industry_group) %>%
  left_join(
    industry_ts %>%
      features(industry_cr, features = feat_stl) %>%
      rename_with(~ paste0("cr_", .), -industry_group),
    by = "industry_group"
  )

# Focus on trend_strength and spikiness
focus_features <- industry_features %>%
  select(industry_group, 
         roa_trend_strength, roa_spikiness,
         cr_trend_strength, cr_spikiness)
```

**Trend strength analysis**

The trend strength visualization in @fig-trend_in reveals significant disruptions to industry financial fundamentals. Five industries in the bottom-left quadrant, indicating severe deterioration in both ROA and Current Ratio trends. Accommodation & Food has the largest bubble, reflecting extreme volatility and weak performance during pandemic restrictions. This pattern reflects the extreme operational challenges faced by hospitality-related sectors during pandemic restrictions.

```{r}
#| label: fig-trend_in
#| fig-cap: "Scatter plot visualizing trend strength disruption across industries during the pandemic period. The 0.3 threshold lines demarcate sectors experiencing significant breakdown in ROA (x-axis) and Current Ratio (y-axis) trends, with bubble size representing ROA spikiness as a measure of financial instability. Industries in bottom-left corner are most severely impacted."

# Trend strength analysis to identify the hardest-hit sectors
trend_analysis <- focus_features %>%
  mutate(
    # Identify sectors where trends are broken (low trend strength)
    roa_trend_disrupted = roa_trend_strength < 0.3,
    cr_trend_disrupted = cr_trend_strength < 0.3,
    overall_trend_impact = (roa_trend_strength + cr_trend_strength) / 2
  )

# Create plot
trend_plot <- focus_features %>%
  ggplot(aes(x = roa_trend_strength, y = cr_trend_strength)) +
  geom_point(aes(size = roa_spikiness, 
                 color = industry_group, 
                 text = industry_group), alpha = 0.7) +
  geom_hline(yintercept = 0.3, linetype = "dashed", color = "red", alpha = 0.7) +
  geom_vline(xintercept = 0.3, linetype = "dashed", color = "red", alpha = 0.7) +
   labs(title = "Impact of Pandemic on Industry Trend Strength<br><sub>Red dashed lines indicate trend strength threshold (0.3).",
       x = "ROA Trend Strength", y = "Current Ratio Trend Strength",
       size = "ROA Spikiness") +
  theme_minimal() +
  theme(legend.position = "none")

# Convert to interactive plotly
trend_in <- ggplotly(trend_plot, tooltip = c("text", "x", "y", "size")) %>%
  plotly::layout(
    hoverlabel = list(bgcolor = "white", font = list(size = 12)),
    annotations = list(
      list(
        x = 0.3, y = 1,
        xref = "x", yref = "y",
        text = "Threshold (0.3)",
        showarrow = FALSE,
        font = list(color = "red", size = 10)
      )
    )
  )

trend_in
```


**Volatility analysis**

After trend strength analysis, the volatility examination in @fig-volatility_in demonstrates divergent patterns in financial metric instability. Five industries show high ROA volatility, but most have moderate Current Ratio spikiness. However, Consumer Services is exceptional with the highest Current Ratio spikiness, suggesting unique liquidity challenges.


```{r}
#| label: fig-volatility_in
#| fig-cap: "Volatility analysis of industry financial metrics during pandemic. Plot displays ROA spikiness (x-axis) against Current Ratio spikiness (y-axis), with industry groups color-coded and bubble sizes indicating ROA trend strength. Median values marked with dashed blue lines. Top-right industries have highest volatility. The ROA spikiness of the Accommodation industry exceeds 20,000 and is not shown in the figure."

# Volatility analysis to identify the sectors with the greatest volatility during the pandemic
volatility_analysis <- focus_features %>%
  mutate(
    high_roa_volatility = roa_spikiness > median(roa_spikiness),
    high_cr_volatility = cr_spikiness > median(cr_spikiness),
    overall_volatility = (scale(roa_spikiness) + scale(cr_spikiness)) / 2)

volatility_plot <- focus_features %>%
  ggplot(aes(x = roa_spikiness, y = cr_spikiness)) +
  geom_point(aes(color = industry_group, 
                 size = roa_trend_strength, 
                 text = industry_group), alpha = 0.7) +
  geom_hline(yintercept = median(focus_features$cr_spikiness), 
             linetype = "dashed", color = "blue", alpha = 0.5) +
  geom_vline(xintercept = median(focus_features$roa_spikiness), 
             linetype = "dashed", color = "blue", alpha = 0.5) +
  coord_cartesian(xlim = c(0, quantile(focus_features$roa_spikiness, 0.90))) + # Limit to 75% quantile
  labs(title = "Industry Volatility Analysis During Pandemic<br><sub>Blue dashed lines represent medians, excluding the extreme values",
       x = "ROA Spikiness", y = "Current Ratio Spikiness",
       size = "ROA Trend Strength") +
  theme_minimal() +
  theme(legend.position = "none")


volatility_in <- ggplotly(volatility_plot, tooltip = c("text", "x", "y", "size")) %>%
  plotly::layout(
    hoverlabel = list(bgcolor = "white", font = list(size = 12)),
    annotations = list(
      list(
        x = median(focus_features$roa_spikiness),
        y = 0.9,
        xref = "x", yref = "paper",
        text = "Median ROA Spikiness",
        showarrow = FALSE,
        font = list(color = "blue", size = 10)
      ),
      list(
        x = 0.5,
        y = median(focus_features$cr_spikiness),
        xref = "paper", yref = "y",
        text = "Median CR Spikiness",
        showarrow = FALSE,
        font = list(color = "blue", size = 10)
      )
    )
  )

volatility_in
```

**Threshold-based industry classification**

Given the exploratory focus of our study and limited emphasis on clustering methods in lectures, we instead used threshold-based segmentation to classify industries into recovery archetypes. This approach categorizes industries based on their ROA trend strength and volatility using the following criteria:


- High trend strength (≥0.7) + Low spikiness (≤median) → "Resilient industries"

- Low trend strength (≤0.3) + High spikiness (≥75th percentile) → "K-shaped recovery industries"

- High trend strength (≥0.7) + High spikiness (≥75th percentile) → "Hidden crisis industries"

- All other combinations → "Moderate impact industries"


As shown in @fig-classification, the analysis reveals clear patterns of differential recovery across sectors. The Construction and Financial & Insurance sectors demonstrate resilient characteristics, maintaining stable performance throughout the pandemic period. Wholesale & Retail Trade, Accommodation & Food Services, and Information & Communication exhibit K-shaped recovery patterns. No industry is hidden crisis. The Administrative & Support sector falls into the hidden crisis category, suggesting underlying structural issues. This classification provides valuable insights into how different industries navigated the pandemic's challenges. Next part, we will focus on the anomalous K-shaped recoveries that suggest uneven performance within certain sectors and resilient industry.


```{r}
industry_classes <- focus_features %>%
  mutate(
    # Calculate threshold
    trend_threshold_high = 0.7,
    trend_threshold_low = 0.3,
    spikiness_threshold_high = quantile(roa_spikiness, 0.75),
    
    # Classification
    recovery_pattern = case_when(
      roa_trend_strength >= trend_threshold_high & 
        roa_spikiness <= median(roa_spikiness) ~ "Resilient",
      roa_trend_strength <= trend_threshold_low & 
        roa_spikiness >= spikiness_threshold_high ~ "K-shaped",
      roa_trend_strength >= trend_threshold_high & 
        roa_spikiness >= spikiness_threshold_high ~ "Hidden Crisis",
      TRUE ~ "Moderate Impact"
    )
  )
```

```{r}
#| label: fig-classification
#| fig-cap: "Classification of industry recovery patterns based on ROA trend strength and volatility. Four distinct patterns emerge: Resilient (high trend strength, low volatility), K-shaped (low trend strength, high volatility), Hidden Crisis (high trend strength but high volatility), and Moderate Impact (all others). Dashed lines represent classification thresholds."

classification_plot <- industry_classes %>%
  ggplot(aes(x = roa_trend_strength, y = roa_spikiness, 
             color = recovery_pattern, 
             text = paste("Industry:", industry_group,
                         "<br>Recovery Pattern:", recovery_pattern,
                         "<br>ROA Trend Strength:", round(roa_trend_strength, 3),
                         "<br>ROA Spikiness:", round(roa_spikiness, 3)))) +
  geom_point(size = 2, alpha = 0.7) +
  geom_vline(xintercept = c(0.3, 0.7), linetype = "dashed", alpha = 0.5) +
  geom_hline(yintercept = median(industry_classes$roa_spikiness), 
             linetype = "dashed", alpha = 0.5) +
  labs(title = "Industry Recovery Patterns Classification",
       x = "ROA Trend Strength", y = "ROA Spikiness",
       color = "Recovery Pattern") +
  theme_minimal() +
  theme(legend.position = "bottom")


classification_plotly <- ggplotly(classification_plot, tooltip = "text") %>%
  plotly::layout(
    hoverlabel = list(bgcolor = "white", font = list(size = 12)),
    title = list(
      text = "Industry Recovery Patterns Classification<br><sub><i>Based on ROA trend strength and volatility analysis</i></sub>",
      x = 0.05
    ),
    margin = list(t = 100)
  )

classification_plotly
```


###### 4. Micro-level evidence of anomalous recovery patterns across industries (ROA and liquidity) - scagnostics (Tignostics) (Lecture 8, 22-27 & Lecture 9, slide 24)

In this section I track how the firm-level relationship between liquidity (Current Ratio) and profitability (ROA) changed during the COVID-19 period. I approach is to treat each firm as a point in the CR–ROA plane, then study the shape of each industry’s point cloud year by year. I use scagnostic metrics to quantify those shapes, and I fit a simple linear model (ROA ~ Current Ratio) for every industry × year to extract slope and residual volatility. I treat 2020 as the COVID year and compare values before and after 2020.

**Data and methods**

I select industries previously identified in the macro analysis (industries labelled K-shaped, Resilient, or Hidden Crisis). For each industry × fiscal year I:

- keep firm-level pairs (current_ratio, roa_pct);

- calculate scagnostic metrics: outlying, stringy, striated, clumpy, sparse, monotonic, dcor;

- fit a linear model roa_pct ~ current_ratio and record slope and residual SD;

- compute year-to-year slope changes and count “jumps” (a jump = |slope change| > 0.5);

- compute pre/post differences in mean residual SD and mean slope (post − pre).

- When an industry × year has too few firms (< 3) the scagnostics or regression are not reported (NA).

**How to read the scagnostic metrics**

- Clumpy: higher values indicate the points form separate clusters → suggests within-industry divergence (possible K-shaped pattern).

- Outlying: higher values indicate presence of extreme firms (very strong survivors or very weak firms).

- Stringy / Monotonic: higher values indicate a clearer linear or monotonic relationship between CR and ROA.

- Residual SD: higher residual SD from the linear fit signals more unexplained scatter (higher micro volatility).



```{r}
selected_industries <- industry_classes %>%
  filter(recovery_pattern %in% c("K-shaped", "Resilient", "Hidden Crisis")) %>%
  pull(industry_group)

# Extract the companies time series data of these industries
selected_companies_ts <- germany_ts %>%
  filter(industry_group %in% selected_industries)
```


```{r}
#| label: tbl-company
#| tbl-cap: "Summary statistics for industries with significant slope changes (jumps > 0). The table shows average slope, maximum absolute slope change, mean residual standard deviation, number of jumps, and differences in volatility and slope before and after the COVID-19 pandemic for three industries: Wholesale & Retail Trade, Information & Communication, and Financial & Insurance."

# Aggregate the original point set by industry × year
by_ind_year <- selected_companies_ts %>%
  group_by(industry_group) %>%
  index_by(fy_year) %>%
  nest(data = c(company_id, roa_pct, current_ratio)) %>%
  mutate(n_obs = map_int(data, ~ nrow(.x)))

# Calculate scagnostics
scag_names <- c("outlying", "stringy", "striated", "clumpy", "sparse", "monotonic", "dcor")

by_ind_year <- by_ind_year %>%
  mutate(
    scags = map(data, ~ {
      df <- .x
      # When there are too few samples, return a row of tibble with the same column names and types.
      if (nrow(df) < 3) {
        return(as_tibble(set_names(as.list(rep(NA_real_, length(scag_names))), scag_names)))
      }
      res <- calc_scags(df$current_ratio, df$roa_pct, scags = scag_names)
      # Make sure to convert to atomic vectors and name them, then convert to tibble
      res_vec <- unlist(res)
      res_named <- set_names(as.numeric(res_vec[scag_names]), scag_names)
      as_tibble(as.list(res_named))
    })
  ) %>%
  unnest(cols = scags)  # Expand into multiple columns

# Fit a linear regression for each industry × year and extract slope, intercept, and resid_sd
by_ind_year <- by_ind_year %>%
  mutate(
    fit = map(data, ~ {
      df <- .x
      if (nrow(df) < 3 || all(is.na(df$current_ratio)) || all(is.na(df$roa_pct))) return(NULL)
      lm(roa_pct ~ current_ratio, data = df)
    }),
    slope = map_dbl(fit, ~ if (is.null(.x)) NA_real_ else coef(.x)[["current_ratio"]]),
    intercept = map_dbl(fit, ~ if (is.null(.x)) NA_real_ else coef(.x)[["(Intercept)"]]),
    resid_sd = map_dbl(fit, ~ if (is.null(.x)) NA_real_ else sd(residuals(.x), na.rm = TRUE))
  ) %>%
  select(-fit)

# Take industry as a time series and calculate its evolution characteristics
industry_tignostics <- by_ind_year %>%
  arrange(industry_group, fy_year) %>%
  group_by(industry_group) %>%
  mutate(
    slope_lag = dplyr::lag(slope),
    slope_change = slope - slope_lag,   # year-to-year change in slope
    slope_pct_change = (slope - slope_lag) / abs(slope_lag) * 100,
    resid_sd_lag = dplyr::lag(resid_sd),
    resid_sd_change = resid_sd - resid_sd_lag,
    # jumps: count of year-to-year slope changes bigger than a threshold (e.g. 0.5)
    big_slope_jump = if_else(abs(slope_change) > 0.5, 1, 0),
    jumps_cumulative = cumsum(replace_na(big_slope_jump, 0))
  ) %>%
  ungroup()

# Calculate the overall volatility_shift (post - pre) before and after the epidemic (taking 2020 as the epidemic year)
vol_shift <- by_ind_year %>%
  mutate(period = case_when(
    fy_year < 2020 ~ "pre",
    fy_year == 2020 ~ "covid_year",
    fy_year > 2020 ~ "post"
  )) %>%
  group_by(industry_group, period) %>%
  summarise(mean_resid_sd = mean(resid_sd, na.rm = TRUE),
            mean_slope = mean(slope, na.rm = TRUE),
            .groups = "drop") %>%
  pivot_wider(names_from = period, values_from = c(mean_resid_sd, mean_slope)) %>%
  mutate(
    vol_pre_post_diff = mean_resid_sd_post - mean_resid_sd_pre,
    slope_pre_post_diff = mean_slope_post - mean_slope_pre
  )

# Merge the results to generate the final summary
industry_summary <- industry_tignostics %>%
  group_by(industry_group) %>%
  summarise(
    n_years = n_distinct(fy_year),
    avg_slope = mean(slope, na.rm = TRUE),
    max_abs_slope_change = if (all(is.na(slope_change))) NA_real_ else max(abs(slope_change), na.rm = TRUE),
    mean_resid_sd = mean(resid_sd, na.rm = TRUE),
    jumps = if (all(is.na(jumps_cumulative))) 0 else max(jumps_cumulative, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  left_join(vol_shift %>% select(industry_group, vol_pre_post_diff, slope_pre_post_diff), by = "industry_group")

kable(industry_summary %>% filter(jumps > 0) %>%
        select(-n_years))
```



```{r}
#| label: fig-scagnostic
#| fig-cap: "This figure shows how four scagnostic metrics changed over time across different industries. The red dashed line marks the year 2020 when COVID-19 pandemic occurred. We can see variations in outlying (extreme values), clumpy (clustering patterns), stringy (linear relationships), and monotonic (consistent trends) metrics before and after the pandemic."

scag_long <- by_ind_year %>%
  select(industry_group, fy_year, clumpy, outlying, stringy, monotonic) %>%
  pivot_longer(cols = c(clumpy, outlying, stringy, monotonic),
               names_to = "scag_metric", values_to = "value")

ggplot(scag_long, aes(x = fy_year, y = value, color = industry_group)) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ scag_metric, scales = "free_y") +
  geom_vline(xintercept = 2020, linetype = "dashed", color = "red") +
  labs(
    title = "Evolution of Scagnostic Metrics over Time",
    subtitle = "Dashed line = COVID year (2020)",
    x = "Fiscal Year", y = "Scagnostic Value"
  ) +
  theme_minimal()

```


```{r}
#| label: fig-animated_plot
#| fig-cap: "The animated plot (appendix) shows the yearly scatter of firms in the CR–ROA plane and fitted lines."

p <- ggplot(selected_companies_ts, aes(x = current_ratio, y = roa_pct, color = industry_group)) +
  geom_point(size = 2, alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Year: {frame_time}", x = "Current Ratio", y = "ROA (%)") +
  scale_color_viridis_d() +
  transition_time(fy_year) +
  ease_aes("linear")
anim <- animate(p, nframes = 200, fps = 10,
                device = "ragg_png", 
                renderer = gifski_renderer("images/my_animation.gif"))


print(anim)
```

**Main results**

Table @tbl-company reports summary statistics for industries with at least one large slope jump (jumps > 0). Three industries experienced jumps: Wholesale & Retail Trade, Information & Communication, and Financial & Insurance. Some industries (e.g., Construction, Accommodation & Food, Administrative & Support) have too few firms in some years and therefore produce NA for scagnostics. Figure @fig-scagnostic shows time series of four metrics (outlying, clumpy, stringy, monotonic) with a vertical line at 2020.

Key patterns observed:

- **Financial & Insurance**: clumpy increases markedly after 2020, while stringy and monotonic fall. Interpretation: the industry becomes more divided into subgroups (clusters of firms) and the previous linear/monotonic relation between CR and ROA weakens. This is consistent with a K-shaped fragmentation: some firms maintain or improve profitability while others decline.

- **Wholesale & Retail Trade**: clumpy decreases after 2020 and monotonic and stringy show relative stability or an increase. Interpretation: the CR–ROA relationship remains more coherent and linear for firms that remain in the sample; less evidence of large within-industry clustering. No strong rise in outlying, so fewer extreme single-firm survivors/losers were detected.

- **Information & Communication**: clumpy falls but outlying increases after 2020. Interpretation: the industry shows fewer multiple clusters but a greater presence of extreme firms — a small set of firms perform very differently from the rest (either much better or much worse), which also suggests micro-level unevenness.


The animated plot (@fig-animated_plot) shows the yearly scatter of firms in the CR–ROA plane and fitted lines. This animation helps visualise how point clouds split, shift, or tighten around a trend line.

**Limitations**

Results are descriptive: scagnostics indicate shape changes but do not prove causality. I will triangulate these micro results with macro indicators (feasts features) in the final section.



###### 5. Summary and interpretation

The exploratory analysis demonstrates that the relationship between corporate profitability and liquidity evolved unevenly across industries during the pandemic. More importantly, the recovery patterns were not the same across or within industries, showing clear evidence of a K-shaped recovery in some sectors.

At the macro level, trend strength and volatility analysis revealed clear segmentation: some industries such as Construction and Finance showed resilient recovery, while others, like Accommodation and Retail, exhibited K-shaped patterns with disrupted ROA trends and high volatility.

The micro-level scagnostic analysis further confirmed this divergence. After removing several industries which have small yearly samples (Construction, 	
Accommodation & Food, and Administrative & Support), industries identified as K-shaped in the macro stage displayed higher clumpiness and outlying scores, indicating growing firm-level inequality in recovery. In contrast, resilient industries maintained stable and monotonic liquidity–profitability relationships over time.

Together, these results suggest that the pandemic not only altered the overall financial dynamics between liquidity and profitability but also intensified disparities within certain industries, forming a distinct K-shaped recovery pattern.






## Conclusions




#### How did the fundamental relationship between corporate liquidity and profitability evolve and fracture within industries during the pandemic?





## Resources

The materials used for this report are:

Wickham H, Averick M, Bryan J, Chang W, McGowan LD, François R,
  Grolemund G, Hayes A, Henry L, Hester J, Kuhn M, Pedersen TL,
  Miller E, Bache SM, Müller K, Ooms J, Robinson D, Seidel DP, Spinu
  V, Takahashi K, Vaughan D, Wilke C, Woo K, Yutani H (2019).
  "Welcome to the tidyverse." _Journal of Open Source Software_,
  *4*(43), 1686. [doi:10.21105/joss.01686](https://doi.org/10.21105/joss.01686).

The links to my use of ChatGPT for help on this assignment are:

- [https://chatgpt.com/share/68857a22-debc-8001-bfcc-10552befba4c]() helped to XXX.

