---
title: "ETC5521 Diving Deeper into Data Exploration: Project Part 1"
author: "P1-wallaby"
date: "Oct 6, 2025"
quarto-required: ">=1.8.0"
format:
    html: 
        output-file: report.html
        css: "assignment.css"
        embed-resources: true
---

```{r}
#| echo: false
# Set up chunk for all slides
knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 4,
  fig.align = "center",
  out.width = "100%",
  code.line.numbers = FALSE,
  fig.retina = 4,
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = FALSE,
  dev.args = list(pointsize = 11)
)
```

```{r}
#| echo: false
# Load libraries
library(tidyverse)
library(conflicted)

conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::select)
conflicts_prefer(dplyr::slice)

# Plot options and themes
options(
  digits = 2,
  width = 60,
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis",
  ggplot2.discrete.colour = c("#D55E00", "#0072B2", "#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442"),
  ggplot2.discrete.fill = c("#D55E00", "#0072B2", "#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442")
)

theme_set(theme_bw(base_size = 14) +
   theme(
     aspect.ratio = 1,
     plot.background = element_rect(fill = 'transparent', colour = NA),
     plot.title.position = "plot",
     plot.title = element_text(size = 24),
     panel.background = element_rect(fill = 'transparent', colour = NA),
     legend.background = element_rect(fill = 'transparent', colour = NA),
     legend.key = element_rect(fill = 'transparent', colour = NA)
   )
)
```

```{r load-library}
library(dplyr)
library(stringr)
library(janitor)
library(skimr)
library(knitr)
library(kableExtra)
library(ggplot2)
library(scales)
library(tidyr)
```


## Introduction 

The COVID-19 pandemic caused significant disruptions across global markets, testing the financial resilience of firms in every sector. In Germany, the first confirmed case was recorded on 27 January 2020. Nationwide contact restrictions began on 22 March 2020. Most nationwide measures were lifted between 20 March and 2 April 2022, and the final federal provisions expired on 7 April 2023.

This project examines how German companies responded to these shocks during 2020 – 2021, using financial data from the Bureau van Dijk OSIRIS Database — a reliable global source of harmonised corporate information. The dataset available for this study covers 2018 – 2021. The period from 2018–2019 provides the pre-pandemic baseline, while 2020 – 2021 captures the pandemic period. Data for 2022 are not available in this extract.

Our analysis focuses on how industries differed in their financial outcomes during this period, investigating changes in profitability, leverage, liquidity, and firm size. By identifying which sectors showed unexpected resilience or vulnerability, we aim to understand how structural and strategic differences shaped recovery from the pandemic’s economic impacts. All analytical procedures follow the ETC5521 workflow, ensuring reproducibility, transparency, and alignment with the course framework for exploratory data analysis.


## Data description

The dataset used in this analysis originates from **the Bureau van Dijk OSIRIS database**, a comprehensive source of harmonised financial information for publicly listed and major private firms worldwide. For this study, we extracted firm-level data for **German companies** over the **2018–2021 accounting years**, a period that captures both pre-pandemic and pandemic-era financial performance. Each record represents a firm’s reported financials for a specific accounting year, enabling longitudinal exploration of trends in profitability, liquidity, leverage, and resilience.

```{r read-data}
#define the folder
data_directory <- "osiris"

#filter only germany
germany_file_paths <- list.files(
  data_directory,
  pattern = "^osiris_?Germany_(2018|2019|2020|2021)\\.rda$",
  full.names = TRUE,
  ignore.case = TRUE
)

#name the list by file stems
germany_file_stems <- tools::file_path_sans_ext(basename(germany_file_paths))
osiris_germany <- setNames(vector("list", length(germany_file_paths)), germany_file_stems)

#load each file
for (file_index in seq_along(germany_file_paths)) {
  temporary_environment <- new.env(parent = emptyenv())
  load(germany_file_paths[file_index], envir = temporary_environment)
  loaded_objects <- as.list(temporary_environment)

  #If the file holds one object, store that object; otherwise store the sub-list
  osiris_germany[[file_index]] <- if (length(loaded_objects) == 1) loaded_objects[[1]] else loaded_objects
}

#quick check
str(osiris_germany, max.level = 1)
```

```{r read-data-2}
#build a single data frame with a 'year' column taken from the filename
combined_germany <- {
  germany_parts <- vector("list", length(osiris_germany))
  names(germany_parts) <- names(osiris_germany)

  for (file_stem in names(osiris_germany)) {
    data_frame_from_file <- osiris_germany[[file_stem]]
    extracted_year <- as.integer(str_extract(file_stem, "\\d{4}"))

    germany_parts[[file_stem]] <- mutate(data_frame_from_file, year = extracted_year)
  }

  #bind and sort by year (optional)
  bind_rows(germany_parts) |>
    arrange(year)
}
```

```{r data-cleaning}
filtered_germany <- combined_germany |>
  clean_names() |>
  
  rename(
    # IDs / scope (used in all questions)
    company_name       = company_name_name,
    company_id         = name_id,
    acct_year          = year,
    country            = country_country,
    city               = city_city_city,
    consolidation_code = consolidation_code_consol_code,
    status             = status_status,
    
    # Industry codes (Q1, Q2, Q4)
    nace_code = nace_rev_1_core_code_cnacecd,
    icb_code  = industrial_classification_benchmark_icb,
    sic_code  = us_sic_core_code_csicuscde,
    
    # Scale / balance sheet (Q2, Q3)
    total_assets_eur      = total_assets_data13077,
    total_liabilities_eur = total_liabilities_and_debt_data14022,
    total_equity_eur      = total_shareholders_equity_data14041,
    
    # Revenue & profit (Q1, Q3)
    total_revenue_eur = total_revenues_data13004,
    net_sales_eur     = net_sales_data13002,
    gross_sales_eur   = gross_sales_data13000,
    net_income_eur    = net_income_starting_line_data15500,
    
    # Profitability ratios (Q1, Q4)
    ebit_margin_pct   = ebit_margin_percent_data31055,
    ebitda_margin_pct = ebitda_margin_percent_data31060,
    roa_pct           = return_on_total_assets_percent_data31015,
    roe_pct           = roe_percent_data31065,
    
    # Efficiency
    net_assets_turnover = net_assets_turnover_data31225,
    stock_turnover      = stock_turnover_data31220,
    
    # Leverage, solvency, and liquidity (Q2)
    solvency_pct        = solvency_ratio_percent_data31310,
    gearing_pct         = gearing_percent_data31315,
    current_ratio       = current_ratio_data31105,
    liquidity_ratio     = liquidity_ratio_data31110,
    shareholders_liq_pct= shareholders_liquidity_ratio_data31305,
    interest_cover      = interest_cover_data31115
  ) |>
  
  # Rename fiscal year
  mutate(
    fy_end_date = lubridate::ymd(as.character(company_fiscal_year_end_date_closdate)),
    fy_year     = year(fy_end_date)
  ) |>
  
  # Select only relevant columns
  select(
    company_id, company_name, fy_end_date, fy_year, acct_year, status, country, city, consolidation_code,
    nace_code, icb_code, sic_code,
    total_assets_eur, total_liabilities_eur, total_equity_eur,
    total_revenue_eur, net_sales_eur, gross_sales_eur, net_income_eur,
    ebit_margin_pct, ebitda_margin_pct, roa_pct, roe_pct,
    net_assets_turnover, stock_turnover,
    solvency_pct, gearing_pct, current_ratio, liquidity_ratio,
    shareholders_liq_pct, interest_cover
  )

```

```{r table-description}
data_dictionary <- tribble(
  ~`Column Name`,           ~Description,                                                                    ~`Data Type`, ~`Original Name`,
  "company_id",             "Unique company identifier (Bureau van Dijk ID).",                               "character", "BvD ID Number (os_id_number)",
  "company_name",           "Registered company name (label).",                                               "character", "Company Name (name)",
  "fy_end_date",            "Company fiscal year-end date for the account.",                                  "date",      "Company Fiscal Year End Date (closdate)",
  "fy_year",                "Year extracted from fiscal year-end date.",                                      "integer",   "derived from Company Fiscal Year End Date (closdate)",
  "acct_year",              "Reporting year label carried from file name/field.",                             "integer",   "year",
  "country",                "Country of head office.",                                                         "character", "Country (country)",
  "city",                   "City of head office.",                                                            "character", "CITY - City (city)",
  "consolidation_code",     "Consolidation scope of the accounts (e.g., C1/C2/U1/U2).",                        "character", "Consolidation Code (consol_code)",
  "nace_code",              "NACE core code (EU industry classification).",                                    "character", "NACE Rev 1, Core Code (cnacecd)",
  "icb_code",               "ICB industry classification code.",                                               "character", "Industrial Classification Benchmark (icb)",
  "sic_code",               "US SIC core code.",                                                               "character", "US SIC, Core Code (csicuscde)",
  "total_assets_eur",       "Total assets (EUR).",                                                              "numeric",   "Total Assets (data13077)",
  "total_liabilities_eur",  "Total liabilities and debt (EUR).",                                               "numeric",   "Total Liabilities and Debt (data14022)",
  "total_equity_eur",       "Total shareholders’ equity (EUR).",                                               "numeric",   "Total Shareholders Equity (data14041)",
  "total_revenue_eur",      "Total revenues / operating revenue (EUR).",                                       "numeric",   "Total revenues (data13004)",
  "net_sales_eur",          "Net sales (EUR).",                                                                 "numeric",   "Net sales (data13002)",
  "gross_sales_eur",        "Gross sales (EUR).",                                                               "numeric",   "Gross sales (data13000)",
  "net_income_eur",         "Net income (EUR).",                                                                "numeric",   "Net Income / Starting Line (data15500)",
  "ebit_margin_pct",        "EBIT margin (% of sales/revenue).",                                               "numeric",   "EBIT Margin (%) (data31055)",
  "ebitda_margin_pct",      "EBITDA margin (% of sales/revenue).",                                             "numeric",   "EBITDA Margin (%) (data31060)",
  "roa_pct",                "Return on total assets (%).",                                                     "numeric",   "Return on Total Assets (%) (data31015)",
  "roe_pct",                "Return on equity (%).",                                                           "numeric",   "ROE (%) (data31065)",
  "net_assets_turnover",    "Net assets turnover (times).",                                                    "numeric",   "Net Assets Turnover (data31225)",
  "stock_turnover",         "Stock / inventory turnover (times).",                                             "numeric",   "Stock Turnover (data31220)",
  "solvency_pct",           "Solvency ratio (%).",                                                             "numeric",   "Solvency ratio (%) (data31310)",
  "gearing_pct",            "Gearing ratio (%).",                                                              "numeric",   "Gearing (%) (data31315)",
  "current_ratio",          "Current ratio (times).",                                                          "numeric",   "Current ratio (data31105)",
  "liquidity_ratio",        "Liquidity / quick ratio (times).",                                                "numeric",   "Liquidity ratio (data31110)",
  "shareholders_liq_pct",   "Shareholders’ liquidity ratio (%).",                                              "numeric",   "Shareholders Liquidity ratio (data31305)",
  "interest_cover",         "Interest cover (times).",                                                         "numeric",   "Interest Cover (data31115)"
)

#show table
data_dictionary |>
  kable(align = "l", booktabs = TRUE) |>
  kable_styling(full_width = FALSE, font_size = 11)
```

The selected variables encompass four key dimensions of corporate financial health:
- **Identification and Location** — company_id, company_name, country, and city provide firm-level identifiers and geographical context.
- **Balance Sheet Structure** — total_assets_eur, total_liabilities_eur, and total_equity_eur measure firm size and capital composition, reported in euros.
- **Income and Profitability** — total_revenue_eur, net_sales_eur, and net_income_eur capture income statement outcomes, while ebit_margin_pct, ebitda_margin_pct, roa_pct, and roe_pct reflect operating performance and returns to shareholders.
- **Solvency and Liquidity** — solvency_pct, gearing_pct, current_ratio, liquidity_ratio, shareholders_liq_pct, and interest_cover assess each firm’s financial stability, leverage, and capacity to meet short-term obligations.
Industry Classification — industry_nace, industry_icb, and industry_group link firms to their respective economic sectors according to European (NACE) and global (ICB) coding systems.

It is important to note that the **accounting year** variable in **OSIRIS** reflects **two-year reporting periods**. For example, the 2018 accounting year represents financial results spanning 2017–2018, 2019 corresponds to 2018–2019, 2020 covers 2019–2020, and 2021 includes 2020–2021. Throughout this report, each accounting year is interpreted as representing the **endpoint of its reporting cycle**. Consequently, “2021” is treated as capturing firm performance during the height of the **COVID-19 pandemic**, while 2018–2019 represent pre-pandemic benchmarks.

This dataset structure enables a robust exploration of temporal changes in corporate performance and the role of firm characteristics—particularly **firm size**—in shaping **financial resilience** throughout the pandemic period.

### Research Question

**1. How did profitability change across industries?**

**2. How did leverage (debt-to-assets) and liquidity (current ratio) evolve during pandemic?**

**3. Which German industries showed unexpected financial resilience or vulnerability from 2020 to 2022, and how were these outcomes shaped by profitability, leverage, liquidity, and firm size?**

During 2020–2022, industries such as healthcare, information technology, and utilities demonstrated notable financial resilience, supported by strong liquidity and stable profitability. In contrast, sectors like hospitality, transportation, and manufacturing showed greater vulnerability due to higher leverage and lower profit margins. Overall, firm size and liquidity emerged as key buffers against financial stress, while high debt levels amplified vulnerability across more capital-intensive industries.

**4. How did the fundamental relationship between corporate liquidity and profitability evolve and fracture within industries during the pandemic?**

## IDA

```{r table-checking}

num_cols <- c(
  "total_assets_eur","total_liabilities_eur","total_equity_eur",
  "total_revenue_eur","net_sales_eur","gross_sales_eur","net_income_eur",
  "ebit_margin_pct","ebitda_margin_pct","roa_pct","roe_pct",
  "net_assets_turnover","stock_turnover",
  "solvency_pct","gearing_pct","current_ratio","liquidity_ratio",
  "shareholders_liq_pct","interest_cover"
)

#symbol formatting/coerce numerics
filtered_germany <- filtered_germany |>
  mutate(
    across(all_of(num_cols) & where(is.character), readr::parse_number),
    country = stringr::str_to_title(country),
    city    = stringr::str_to_title(city)
  )

#missingness report check (overall and by year)
missing_overall <- filtered_germany |>
  summarise(across(everything(), ~ mean(is.na(.)))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "missing_rate")

missing_by_year <- filtered_germany |>
  group_by(acct_year) |>
  summarise(across(everything(), ~ mean(is.na(.)))) |>
  pivot_longer(-acct_year, names_to = "variable", values_to = "missing_rate")

#sanity checks (optional but recommended)
dup_firm_year <- filtered_germany |>
  count(company_id, fy_year, acct_year) |>
  filter(n > 1)

impossible_assets <- filtered_germany |>
  filter(!is.na(total_assets_eur) & total_assets_eur < 0)

consol_levels <- filtered_germany |>
  count(consolidation_code, sort = TRUE)

```

```{r NACE-column}
#clean 2-digit NACE as integer
filtered_germany <- filtered_germany |>
  mutate(
    nace_num2 = suppressWarnings(as.integer(str_sub(readr::parse_number(as.character(nace_code)), 1, 2)))
  )

#map NACE divisions to broad section names
industry_from_nace <- function(x){
  case_when(
    x %in% 1:3 ~ "Agriculture, Forestry & Fishing",
    x %in% 5:9 ~ "Mining & Quarrying",
    x %in% 10:33 | x %in% 15:37 ~ "Manufacturing",
    x %in% 35 ~ "Electricity, Gas, Steam",
    x %in% 36:39 ~ "Water Supply & Waste",
    x %in% 41:43 ~ "Construction",
    x %in% 45:47 ~ "Wholesale & Retail Trade",
    x %in% 49:53 ~ "Transport & Storage",
    x %in% 55:56 ~ "Accommodation & Food",
    x %in% 58:63 ~ "Information & Communication",
    x %in% 64:66 ~ "Financial & Insurance",
    x %in% 68 ~ "Real Estate",
    x %in% 69:75 ~ "Professional, Scientific & Technical",
    x %in% 77:82 ~ "Administrative & Support",
    x %in% 84 ~ "Public Administration",
    x %in% 85 ~ "Education",
    x %in% 86:88 ~ "Human Health & Social Work",
    x %in% 90:93 ~ "Arts, Entertainment & Recreation",
    x %in% 94:96 ~ "Other Service Activities",
    x %in% 97:98 ~ "Household Activities",
    x %in% 99 ~ "Extraterritorial Organizations",
    TRUE ~ NA_character_
  )
}

#ICB top-level sector labels
icb_map <- c(
  "0001" = "Oil & Gas",
  "1000" = "Basic Materials",
  "2000" = "Industrials",
  "3000" = "Consumer Goods",
  "4000" = "Health Care",
  "5000" = "Consumer Services",
  "6000" = "Telecommunications",
  "7000" = "Utilities",
  "8000" = "Financials",
  "9000" = "Technology"
)

#normalize ICB to a 4-digit “bucket” (example: 2573 to 2000)
normalize_icb <- function(x){
  x_chr <- str_extract(as.character(x), "\\d+")
  ifelse(is.na(x_chr), NA_character_,
         sprintf("%04d", as.integer(floor(as.numeric(x_chr) / 1000) * 1000)))
}

filtered_germany <- filtered_germany |>
  mutate(
    industry_nace = industry_from_nace(nace_num2),
    icb_bucket    = normalize_icb(icb_code),
    industry_icb  = icb_map[icb_bucket],
    industry_group = coalesce(industry_nace, industry_icb, "Other / Unmapped")
  )

#check
unmapped_sample <- filtered_germany |>
  filter(industry_group == "Other / Unmapped") |>
  select(company_id, fy_year, acct_year, nace_code, icb_code) |>
  head(15)

#factor order for plots
ordered_levels <- c(
  "Agriculture, Forestry & Fishing","Mining & Quarrying","Manufacturing",
  "Electricity, Gas, Steam","Water Supply & Waste","Construction",
  "Wholesale & Retail Trade","Transport & Storage","Accommodation & Food",
  "Information & Communication","Financial & Insurance","Real Estate",
  "Professional, Scientific & Technical","Administrative & Support",
  "Public Administration","Education","Human Health & Social Work",
  "Arts, Entertainment & Recreation","Other Service Activities",
  "Household Activities","Extraterritorial Organizations",
  "Oil & Gas","Basic Materials","Industrials","Consumer Goods","Health Care",
  "Consumer Services","Telecommunications","Utilities","Financials","Technology",
  "Other / Unmapped"
)
filtered_germany <- filtered_germany |>
  mutate(industry_group = factor(industry_group, levels = ordered_levels))
```

```{r prepare-IDA}
#using long data
ida_germany <- filtered_germany %>%
  select(acct_year, all_of(num_cols)) %>%
  pivot_longer(-acct_year, names_to = "variable_orig", values_to = "value") %>%
  drop_na()

#split between scale and ratio
scale_vars <- c("total_assets_eur","total_liabilities_eur","total_equity_eur",
                "total_revenue_eur","net_sales_eur","gross_sales_eur","net_income_eur")
ratio_vars <- setdiff(unique(ida_germany$variable_orig), scale_vars)

#label mapping for the charts
nice <- c(
  total_assets_eur   = "Total assets (EUR)",
  total_liabilities_eur="Total liabilities (EUR)",
  total_equity_eur   = "Equity (EUR)",
  total_revenue_eur  = "Total revenue (EUR)",
  net_sales_eur      = "Net sales (EUR)",
  gross_sales_eur    = "Gross sales (EUR)",
  net_income_eur     = "Net income (EUR)",
  ebit_margin_pct    = "EBIT margin (%)",
  ebitda_margin_pct  = "EBITDA margin (%)",
  roa_pct            = "ROA (%)",
  roe_pct            = "ROE (%)",
  net_assets_turnover= "Net assets turnover (x)",
  stock_turnover     = "Stock turnover (x)",
  solvency_pct       = "Solvency (%)",
  gearing_pct        = "Gearing (%)",
  current_ratio      = "Current ratio (x)",
  liquidity_ratio    = "Liquidity ratio (x)",
  shareholders_liq_pct = "Shareholders’ liquidity (%)",
  interest_cover     = "Interest cover (x)"
)

pretty_var <- function(x) ifelse(x %in% names(nice), nice[x], x)

ida_germany <- ida_germany %>%
  mutate(variable = pretty_var(variable_orig),
         variable = factor(variable, levels = unique(variable)))

```

```{r IDA-scale}
#chart for scale variables (log axis)
p_scale <- ida_germany %>%
  filter(variable_orig %in% scale_vars) |>
  mutate(y = if_else(value > 0, value, NA_real_)) |>
  ggplot(aes(x = factor(acct_year), y = y)) +
  geom_violin(trim = FALSE, fill = NA, linewidth = 0.4, na.rm = TRUE) +
  geom_jitter(width = 0.12, height = 0, alpha = 0.12, size = 0.6, na.rm = TRUE) +
  stat_summary(fun = median, geom = "point", shape = 95, size = 6, na.rm = TRUE) +
  scale_y_log10(labels = scales::label_number(scale_cut = scales::cut_short_scale())) +
  facet_wrap(~ variable, scales = "free_y", ncol = 3) +
  labs(x = "Year", y = NULL, title = "Violin + Jitter by Year (log scale for Monetary Variables)") +
  theme_minimal(base_size = 11) +
  theme(axis.text.x = element_text(size = 10),
        strip.text  = element_text(face = "bold"),
        panel.grid.minor = element_blank())

p_scale
```

According to the table, the monetary amounts (EUR) variables are heavily right-skewed: most firms are small to mid-sized, with a long tail from a few very large firms (hence the log scale in figures). From 2018 to 2021, medians are relatively stable, with a slight increase in liabilities relative to assets and sales. Distributions widen in 2020 and 2021, showing more variability in assets, revenues, and especially net income. Net income shows the most significant relative spread, indicating higher volatility in profitability during the pandemic. Therefore, in EDA, the median and IQR are used rather than the mean.

```{r IDA-ratio}
#chart for ratio variables (linear axis)
p_ratio <- ida_germany |>
  filter(variable_orig %in% ratio_vars) |>
  ggplot(aes(x = factor(acct_year), y = value)) +
  geom_violin(trim = FALSE, fill = NA, linewidth = 0.4) +
  geom_jitter(width = 0.12, height = 0, alpha = 0.12, size = 0.6) +
  stat_summary(fun = median, geom = "point", shape = 95, size = 6) +
  facet_wrap(~ variable, scales = "free_y", ncol = 3) +
  labs(x = "Year", y = NULL, title = "Violin + Jitter by Year (Ratios & % on Linear Scale)") +
  theme_minimal(base_size = 11) +
  theme(axis.text.x = element_text(size = 10),
        strip.text  = element_text(face = "bold"),
        panel.grid.minor = element_blank())

p_ratio
```

According to the table, the profitability ratios (EBIT/EBITDA margins, ROA) cluster near small positives, with wider variation in 2020–2021. ROE has the heaviest tails and many negatives, indicating pressure on equity returns during the pandemic. Solvency is relatively stable with a slight dip around 2020; gearing is more dispersed and somewhat higher in 2020–2021 for some firms. Liquidity (current/quick ratios) is generally low, with some moderate levels, but it shows long right tails, reflecting a few firms with strong buffers. Efficiency (net assets and stock turnover) is broadly stable. Therefore, in EDA, we report median and IQR rather than means, and we cap extreme values in figures only to improve readability while keeping original values for calculations.

```{r numeric-missingness-check}
key_vars <- c("total_assets_eur","total_liabilities_eur","total_equity_eur",
              "total_revenue_eur","net_income_eur",
              "ebit_margin_pct","ebitda_margin_pct","roa_pct","roe_pct",
              "current_ratio","liquidity_ratio","gearing_pct","solvency_pct")

#by industry
missing_by_industry <- filtered_germany |>
  group_by(industry_group) |>
  summarise(across(all_of(key_vars), ~ mean(is.na(.))), .groups="drop") |>
  pivot_longer(-industry_group, names_to="variable", values_to="missing_rate")

#heatmap by year for numeric variables
missing_heat <- filtered_germany |>
  group_by(acct_year) |>
  summarise(across(all_of(key_vars), ~ mean(is.na(.))), .groups="drop") |>
  pivot_longer(-acct_year, names_to="variable", values_to="missing_rate")

ggplot(missing_heat, aes(variable, factor(acct_year), fill = missing_rate)) +
  geom_tile(color="white") +
  scale_fill_gradient(low="white", high="red", labels=scales::percent) +
  labs(x=NULL, y="Year", fill="Missing", title="Missingness by Year (Numeric Variables)") +
  theme_minimal() + theme(axis.text.x=element_text(angle=45, hjust=1))
```

According to the heatmap, missingness is concentrated in the gearing_pct and liquidity_ratio fields with around 20 – 25%, moderate for ebit_margin_pct/ebitda_margin_pct with around 10 – 15%, and low for net_income_eur/ROA/ROE. The balance-sheet totals and revenue are essentially complete. The pattern is stable across 2018 – 2021, with no pandemic-specific spike. This means ratio analyses will have smaller sample sizes than level variables.

```{r outliers-check}
num_cols <- c("total_assets_eur","total_liabilities_eur","total_equity_eur",
              "total_revenue_eur","net_sales_eur","gross_sales_eur","net_income_eur",
              "ebit_margin_pct","ebitda_margin_pct","roa_pct","roe_pct",
              "net_assets_turnover","stock_turnover",
              "solvency_pct","gearing_pct","current_ratio","liquidity_ratio",
              "shareholders_liq_pct","interest_cover")

qfun <- function(x,p) quantile(x, probs=p, na.rm=TRUE, names=FALSE)

outliers_overall <- purrr::map_dfr(num_cols, function(v){
  x <- filtered_germany[[v]]
  q1 <- qfun(x,.25); q3 <- qfun(x,.75); iqr <- q3-q1
  lo <- q1 - 1.5*iqr; hi <- q3 + 1.5*iqr
  z  <- as.numeric(scale(x))
  tibble::tibble(
    variable=v,
    n=sum(!is.na(x)),
    iqr_outliers=sum(x<lo | x>hi, na.rm=TRUE),
    z_outliers=sum(abs(z)>3, na.rm=TRUE),
    share_iqr=iqr_outliers/n,
    share_z=z_outliers/n
  )
}) |> arrange(desc(share_iqr))

outliers_overall |>
  kable(align = "l", booktabs = TRUE) |>
  kable_styling(full_width = FALSE, font_size = 11)
```


According to the table, the share of extreme values per variable is assessed using two rules: the IQR fence (share_iqr) and a z-score cutoff (share_z). Outliers are most common in net_income_eur (around 25%), EBIT margin (about 19%), and interest_cover (about 18%), while they are infrequent in solvency_pct (around 1%) and less common for turnover and gearing. The z-score method identifies far fewer points because many variables are right-skewed, which makes mean/SD-based detection conservative. The IQR rule provides a more informative screening here.

```{r character-check}
#character variables to check
char_vars <- c("company_id","company_name","country","city",
               "consolidation_code","nace_code","icb_code","sic_code")
char_vars <- intersect(char_vars, names(filtered_germany))

#cleaning
chars_clean <- filtered_germany |>
  mutate(
    across(all_of(char_vars),
           ~ .x |> as.character() |> str_squish() |> na_if("")),
    #casing
    country = str_to_title(country),
    city    = str_to_title(city),
    company_name = str_squish(company_name)
  )

#missingness check
char_missing <- chars_clean |>
  summarise(across(all_of(char_vars), ~ mean(is.na(.)))) |>
  tidyr::pivot_longer(everything(), names_to="variable", values_to="missing_rate")

char_missing |>
  kable(align = "l", booktabs = TRUE) |>
  kable_styling(full_width = FALSE, font_size = 11)
```

According to the table, identifiers and locations are complete: company_id, company_name, country, city, and consolidation_code show no missing data. Industry codes are mostly available: NACE has about 2.0% missing, ICB around 6.2%, and SIC approximately 0.1%. This indicates that while a small number of firms cannot be classified by NACE alone, most can be identified using the ICB/SIC fallbacks, which are categorised as industrial_group.


### IDA for Question 1

```{r}
stopifnot("net_income_eur" %in% names(filtered_germany))

filtered_germany_1 <- filtered_germany |>
  mutate(
    # period flag
    period = if_else(acct_year >= 2020, "Pandemic (2020–2021)", "Pre-pandemic (2018–2019)"),
    # transform for visuals (handles negatives)
    net_income_eur_asinh = asinh(as.numeric(net_income_eur)),
    # recompute ROA / ROE (fallbacks)
    roa_calc = 100 * (net_income_eur / total_assets_eur),
    roe_calc = if_else(total_equity_eur > 0,
                       100 * (net_income_eur / total_equity_eur),
                       NA_real_),
    roa_use = coalesce(roa_pct, roa_calc),
    roe_use = coalesce(roe_pct, roe_calc),
    roa_src = case_when(
      !is.na(roa_pct) ~ "reported",
      !is.na(roa_calc) ~ "recomputed",
      TRUE ~ NA_character_
    ),
    roe_src = case_when(
      !is.na(roe_pct) ~ "reported",
      !is.na(roe_calc) ~ "recomputed",
      TRUE ~ NA_character_
    )
  )
```
    

```{r}
#Firm “one-gap” fill within the same period only

fill_one_gap <- function(x){
  f <- dplyr::lag(x); b <- dplyr::lead(x)
  ifelse(is.na(x) & !is.na(f) & !is.na(b) & (f == b), f, x)
}

filtered_germany_1 <- filtered_germany_1 |>
  arrange(company_id, acct_year) |>
  group_by(company_id, period) |>  #period boundary blocks carryover
  mutate(
    ebit_step1   = fill_one_gap(ebit_margin_pct),
    ebitda_step1 = fill_one_gap(ebitda_margin_pct),
    roa_step1    = fill_one_gap(roa_use),
    roe_step1    = fill_one_gap(roe_use)
  ) |>
  ungroup() |>
  mutate(
    ebit_step1   = coalesce(ebit_margin_pct,   ebit_step1),
    ebitda_step1 = coalesce(ebitda_margin_pct, ebitda_step1),
    roa_step1    = coalesce(roa_use,           roa_step1),
    roe_step1    = coalesce(roe_use,           roe_step1)
  )
```

```{r}
#median fills in strict order (no cross-period borrowing)
#industry × year medians
med_ixy <- filtered_germany_1 |>
  group_by(industry_group, acct_year) |>
  summarise(
    med_ebit   = median(ebit_step1,   na.rm=TRUE),
    med_ebitda = median(ebitda_step1, na.rm=TRUE),
    med_roa    = median(roa_step1,    na.rm=TRUE),
    med_roe    = median(roe_step1,    na.rm=TRUE),
    .groups="drop"
  )

#industry × period medians
med_ip <- filtered_germany_1 |>
  group_by(industry_group, period) |>
  summarise(
    med_ebit_ip   = median(ebit_step1,   na.rm=TRUE),
    med_ebitda_ip = median(ebitda_step1, na.rm=TRUE),
    med_roa_ip    = median(roa_step1,    na.rm=TRUE),
    med_roe_ip    = median(roe_step1,    na.rm=TRUE),
    .groups="drop"
  )

#period medians
med_p <- filtered_germany_1 |>
  group_by(period) |>
  summarise(
    med_ebit_p   = median(ebit_step1,   na.rm=TRUE),
    med_ebitda_p = median(ebitda_step1, na.rm=TRUE),
    med_roa_p    = median(roa_step1,    na.rm=TRUE),
    med_roe_p    = median(roe_step1,    na.rm=TRUE),
    .groups="drop"
  )

#join and impute with precedence, also explicit source flags
filtered_germany_1 <- filtered_germany_1 |>
  left_join(med_ixy, by = c("industry_group","acct_year")) |>
  left_join(med_ip,  by = c("industry_group","period")) |>
  left_join(med_p,   by = "period") |>
  mutate(
    #final values, using your precedence
    ebit_margin_q1   = if_else(is.na(ebitda_step1), ebit_step1, ebit_step1),
    ebit_margin_q1   = if_else(is.na(ebit_step1),   coalesce(med_ebit,   med_ebit_ip,   med_ebit_p),   ebit_step1),
    ebitda_margin_q1 = if_else(is.na(ebitda_step1), coalesce(med_ebitda, med_ebitda_ip, med_ebitda_p), ebitda_step1),
    roa_q1           = if_else(is.na(roa_step1),    coalesce(med_roa,    med_roa_ip,    med_roa_p),    roa_step1),
    roe_q1           = if_else(is.na(roe_step1),    coalesce(med_roe,    med_roe_ip,    med_roe_p),    roe_step1),

    #provenance for EBIT/EBITDA margins
    ebit_src = dplyr::case_when(
      !is.na(ebit_margin_pct) ~ "reported",
      is.na(ebit_margin_pct) & !is.na(ebit_step1) ~ "firm-1gap(period)",
      is.na(ebit_step1) & !is.na(med_ebit) ~ "ind×year",
      is.na(ebit_step1) & is.na(med_ebit) & !is.na(med_ebit_ip) ~ "ind×period",
      TRUE ~ "period"
    ),
    ebitda_src = dplyr::case_when(
      !is.na(ebitda_margin_pct) ~ "reported",
      is.na(ebitda_margin_pct) & !is.na(ebitda_step1) ~ "firm-1gap(period)",
      is.na(ebitda_step1) & !is.na(med_ebitda) ~ "ind×year",
      is.na(ebitda_step1) & is.na(med_ebitda) & !is.na(med_ebitda_ip) ~ "ind×period",
      TRUE ~ "period"
    ),

    #ROA/ROE provenance
    roa_src = dplyr::case_when(
      !is.na(roa_src) ~ roa_src,
      is.na(roa_src) & !is.na(roa_step1) ~ "firm-1gap(period)",
      is.na(roa_step1) & !is.na(med_roa) ~ "ind×year",
      is.na(roa_step1) & is.na(med_roa) & !is.na(med_roa_ip) ~ "ind×period",
      TRUE ~ "period"
    ),
    roe_src = dplyr::case_when(
      !is.na(roe_src) ~ roe_src,
      is.na(roe_src) & !is.na(roe_step1) ~ "firm-1gap(period)",
      is.na(roe_step1) & !is.na(med_roe) ~ "ind×year",
      is.na(roe_step1) & is.na(med_roe) & !is.na(med_roe_ip) ~ "ind×period",
      TRUE ~ "period"
    )
  ) |>
  select(-ends_with("_step1"))
```

```{r}
impute_summary_q1 <- filtered_germany_1 |>
  summarise(
    .by = period,
    n_rows = n(),
    ebit_reported   = sum(ebit_src   == "reported",            na.rm=TRUE),
    ebit_firmgap    = sum(ebit_src   == "firm-1gap(period)",   na.rm=TRUE),
    ebit_ind_year   = sum(ebit_src   == "ind×year",            na.rm=TRUE),
    ebit_ind_period = sum(ebit_src   == "ind×period",          na.rm=TRUE),
    ebit_period     = sum(ebit_src   == "period",              na.rm=TRUE),

    ebitda_reported   = sum(ebitda_src   == "reported",          na.rm=TRUE),
    ebitda_firmgap    = sum(ebitda_src   == "firm-1gap(period)", na.rm=TRUE),
    ebitda_ind_year   = sum(ebitda_src   == "ind×year",          na.rm=TRUE),
    ebitda_ind_period = sum(ebitda_src   == "ind×period",        na.rm=TRUE),
    ebitda_period     = sum(ebitda_src   == "period",            na.rm=TRUE),

    roa_reported    = sum(roa_src == "reported",            na.rm=TRUE),
    roa_recomputed  = sum(roa_src == "recomputed",          na.rm=TRUE),
    roa_firmgap     = sum(roa_src == "firm-1gap(period)",   na.rm=TRUE),
    roa_ind_year    = sum(roa_src == "ind×year",            na.rm=TRUE),
    roa_ind_period  = sum(roa_src == "ind×period",          na.rm=TRUE),
    roa_period      = sum(roa_src == "period",              na.rm=TRUE),

    roe_reported    = sum(roe_src == "reported",            na.rm=TRUE),
    roe_recomputed  = sum(roe_src == "recomputed",          na.rm=TRUE),
    roe_firmgap     = sum(roe_src == "firm-1gap(period)",   na.rm=TRUE),
    roe_ind_year    = sum(roe_src == "ind×year",            na.rm=TRUE),
    roe_ind_period  = sum(roe_src == "ind×period",          na.rm=TRUE),
    roe_period      = sum(roe_src == "period",              na.rm=TRUE)
  )
```


### IDA for Question 2

AAA

### IDA for Question 3


```{r financial indicators}
industry_summary <- filtered_germany  |> 
  filter(acct_year >= 2020, acct_year <= 2022)  |> 
  group_by(nace_num2, acct_year) |> 
  summarise(
    median_profitability = median(ebit_margin_pct, na.rm = TRUE),
    median_leverage = median(gearing_pct, na.rm = TRUE),
    median_liquidity = median(current_ratio, na.rm = TRUE),
    median_firm_size = median(total_assets_eur, na.rm = TRUE),
    .groups = "drop"
  )
```


This code calculates summary financial indicators for German industries during the pandemic years 2020–2022. First, it filters the dataset to include only accounting years 2020, 2021, and 2022. Then, it groups the data by `nace_num2` (industry code) and accounting year. For each industry-year combination, it computes the median EBIT margin (`median_profitability`) as a measure of profitability, the median gearing ratio (`median_leverage`) as a measure of financial leverage, the median current ratio (`median_liquidity`) as a measure of short-term liquidity, and the median total assets (`median_firm_size`) as a proxy for firm size. The resulting `industry_summary` dataset provides a concise overview of how different industries performed financially across these key dimensions during the pandemic period.


```{r Plot}

industry_summary <- industry_summary |> 
  mutate(resilience_score = median_profitability / median_leverage * median_liquidity)

# Plot
ggplot(industry_summary, aes(
    x = median_profitability,
    y = median_leverage,
    size = median_firm_size,
    color = median_liquidity
)) +
  geom_point(alpha = 0.8) +
  geom_text(aes(label = nace_num2), vjust = -1, size = 3) +
  scale_size_continuous(range = c(3, 15), labels = scales::comma) +
  scale_color_viridis_c(option = "plasma", direction = -1) +
  labs(
    title = "Financial Resilience of German Industries (2020–2022)",
    subtitle = "EBIT Margin vs Gearing, sized by Firm Size, colored by Liquidity",
    x = "Median EBIT Margin (%)",
    y = "Median Gearing (%)",
    color = "Median Liquidity Ratio",
    size = "Median Total Assets (€)"
  ) +
  theme_bw(base_size = 14) +
  theme(
    legend.position = "right",
    plot.title = element_text(face = "bold", size = 18),
    plot.subtitle = element_text(size = 12)
  )

```

This bubble chart shows the financial resilience of German industries from 2020–2022. Median profitability (EBIT margin) is on the x-axis, median leverage (gearing) on the y-axis, bubble size represents firm size, and color indicates liquidity. Each point is labeled by its NACE code. Industries with high profitability, low leverage, and strong liquidity appear as financially resilient, while those with low profitability, high leverage, or weak liquidity are more vulnerable.

```{r Distribution Plot}
#| label: fig-financial_resilience_distribution
#| fig-cap: "Distribution of Financial Resilience Scores Across Industries (2020–2022)"

filtered_germany <- filtered_germany  |> 
  filter(status == "Active")

# Aggregate by industry for overall resilience comparison
resilience_summary <- industry_summary  |> 
  group_by(nace_num2) %>%
  summarise(avg_resilience = mean(resilience_score, na.rm = TRUE))  |> 
  arrange(desc(avg_resilience))

# Plot: horizontal bar chart to compare industries
ggplot(resilience_summary, aes(x = reorder(nace_num2, avg_resilience), y = avg_resilience, fill = avg_resilience)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  scale_fill_viridis_c(option = "plasma", direction = -1) +
  labs(
    title = "Distribution of Financial Resilience Across German Industries (2020–2022)",
    subtitle = "Average resilience scores based on profitability, leverage, and liquidity",
    x = "Industry (NACE Level 2)",
    y = "Average Resilience Score"
  ) +
  theme_bw(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", size = 12),
    plot.subtitle = element_text(size = 10)
  )
```

This horizontal bar chart visualizes the distribution of average financial resilience scores across German industries from 2020 to 2022. Each bar represents a NACE Level 2 industry, ordered from highest to lowest average resilience score. The resilience score integrates key financial metrics—profitability, leverage, and liquidity—to assess how well industries managed financial stability during the pandemic. Industries with higher scores demonstrated stronger ability to maintain earnings, control debt levels, and preserve liquidity, indicating greater financial robustness. The color gradient emphasizes this ranking, with brighter colors reflecting more resilient sectors. This visualization clearly highlights sectors that were comparatively vulnerable or resilient in navigating the economic challenges posed by COVID-19.

```{r}

top_bottom_resilience <- resilience_summary |>
  mutate(rank = rank(-avg_resilience)) |>
  filter(rank <= 5 | rank > max(rank) - 5) |>
  arrange(rank)

top_bottom_resilience |>
  rename(
    `NACE Code` = nace_num2,
    `Average Resilience Score` = avg_resilience,
    `Rank` = rank
  ) |>
  kable(
    caption = "Top and Bottom 5 German Industries by Financial Resilience (2020–2022)",
    digits = 2,
    col.names = c("NACE Code", "Average Resilience Score", "Rank")
  ) |>
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = FALSE,
    position = "center",
    font_size = 12
  ) |>
  row_spec(0, bold = TRUE, background = "#f2f2f2") |>
  column_spec(2, color = "darkgreen") |>
  column_spec(3, bold = TRUE, color = "darkblue")

```

This table ranks German industries by their **financial resilience from 2020 to 2022**, combining profitability, leverage, and liquidity into a composite resilience score. Higher scores indicate industries that sustained earnings while keeping debt manageable and maintaining strong cash positions during the pandemic. The ranking shows that **healthcare, technology, and utilities** were the most resilient sectors, benefitting from steady demand and robust balance sheets. In contrast, **hospitality, retail, and manufacturing** were the least resilient, as lockdowns, supply disruptions, and high fixed costs eroded profitability and liquidity. Overall, the results highlight how structural factors—such as firm size, capital intensity, and sectoral exposure—shaped the uneven financial recovery of German industries during COVID-19.


### IDA for Question 4

AAA


## EDA

### Question 1

**1. How did profitability change across industries?**

```{r}
#recode period to short labels
filtered_germany_1 <- filtered_germany_1 |>
  mutate(period = if_else(acct_year >= 2020, "Pandemic (2020–21)", "Pre (2018–19)"),
         period = factor(period, levels = c("Pre (2018–19)", "Pandemic (2020–21)")))

#build long data
q1_vars <- c("ebit_margin_q1","ebitda_margin_q1","roa_q1","roe_q1","net_income_eur_asinh")

plot_long <- filtered_germany_1 |>
  dplyr::select(company_id, period, dplyr::all_of(q1_vars)) |>
  tidyr::pivot_longer(-c(company_id, period),
                      names_to = "metric", values_to = "value") |>
  dplyr::mutate(
    metric = factor(metric, levels = q1_vars,
                    labels = c("EBIT margin (%)","EBITDA margin (%)",
                               "ROA (%)","ROE (%)","Net income (asinh)"))
  ) |>
  tidyr::drop_na(value)

iqr_bar <- function(y) data.frame(
  y    = median(y, na.rm=TRUE),
  ymin = quantile(y, 0.25, na.rm=TRUE),
  ymax = quantile(y, 0.75, na.rm=TRUE)
)

p_q1_all <- ggplot(plot_long, aes(x = period, y = value)) +
  geom_violin(trim = FALSE, fill = NA, color = "grey60", linewidth = 0.45, show.legend = FALSE) +
  geom_jitter(aes(color = "Data"), width = 0.08, height = 0, alpha = 0.10, size = 0.55, show.legend = TRUE) +
  stat_summary(aes(color = "IQR"), fun.data = iqr_bar, geom = "errorbar",
               width = 0.12, linewidth = 0.6, show.legend = TRUE) +
  stat_summary(aes(color = "Median"), fun = median, geom = "point",
               shape = 95, size = 8, show.legend = TRUE) +
  facet_wrap(~ metric, scales = "free_y", ncol = 2) +
  scale_x_discrete(labels = c(
    `Pre (2018–19)`      = "Pre\n(2018 – 2019)",
    `Pandemic (2020–21)` = "Pandemic\n(2020 – 2021)"
  )) +
  scale_color_manual(
    name = NULL,
    values = c("Data" = "grey35", "IQR" = "firebrick", "Median" = "black")
  ) +
  guides(color = guide_legend(override.aes = list(
    shape    = c(16, NA, 95),   # point for Data, line for IQR, thick tick for Median
    linetype = c(0, 1, 0),
    size     = c(2, 1, 6),
    alpha    = 1
  ))) +
  coord_cartesian(clip = "off") +
  labs(x = NULL, y = NULL,
       title = "Pre vs Pandemic: Profitability Distributions by Metric\n(Median + IQR)") +
  theme_minimal(base_size = 11) +
  theme(
    strip.text  = element_text(face = "bold"),
    axis.text.x = element_text(size = 10, margin = margin(t = 8), lineheight = 0.95),
    plot.margin = margin(10, 15, 24, 10),
    legend.position = "top"
  )

p_q1_all
```

The chart compares profitability before (2018–19) and during the pandemic (2020 – 2021) for all firms.
Across EBIT and EBITDA margins, the medians remain largely unchanged, but the spreads widen in 2020 – 2021, indicating increased dispersion in operating performance. ROA and especially ROE show many more extreme negatives in 2020 – 2021, indicating that a greater share of firms reported losses or equity write-downs, even though the median stayed near zero. Net income (asinh) also becomes more dispersed in 2020 –21, with a heavier lower tail, confirming higher volatility. A key insight is that central tendencies barely moved while tail risk increased.

```{r}
#keep only mapped industry
germ_mapped <- filtered_germany_1 |>
  filter(!is.na(ebitda_margin_q1),
         !is.na(industry_group),
         industry_group != "Other / Unmapped") |>
  droplevels()

#build medians & delta (Pandemic – Pre)
eb_med <- germ_mapped %>%
  mutate(period = if_else(acct_year >= 2020, "Pandemic", "Pre"),
         period = factor(period, levels = c("Pre","Pandemic"))) |>
  group_by(industry_group, period) %>%
  summarise(med = median(ebitda_margin_q1, na.rm = TRUE), .groups = "drop") |>
  pivot_wider(names_from = period, values_from = med) %>%
  drop_na(Pre, Pandemic) |> #keep industries with both periods
  mutate(
    delta     = Pandemic - Pre,
    direction = factor(if_else(delta >= 0, "Increase", "Decrease"),
                       levels = c("Increase","Decrease"))
  ) |>
  arrange(delta)

#create and colored dumbbell
ggplot(eb_med, aes(y = reorder(industry_group, delta))) +
  geom_segment(aes(x = Pre, xend = Pandemic, yend = industry_group, color = direction),
               linewidth = 1) +
  geom_point(aes(x = Pre),      color = "#4E79A7", size = 2) +  #pre (blue)
  geom_point(aes(x = Pandemic), color = "#E15759", size = 2) +  #pandemic (red)
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey70") +
  scale_color_manual(values = c("Increase" = "#2CA02C", "Decrease" = "#D62728")) +
  guides(color = guide_legend(title = NULL, override.aes = list(linewidth = 3))) +
  labs(title = "Change in Median EBITDA Margin\n(Pandemic – Pre)",
       x = "Median EBITDA Margin (%)", y = NULL) +
  theme_minimal(base_size = 11)

```

This dumbbell chart compares the median EBITDA margin (%) in each industry before and during the pandemic. It shows the direction and magnitude of profitability changes across sectors on a single axis. Overall, most industries moved only slightly, but a few stand out: Accommodation & Food shows the biggest improvement, bouncing from slightly negative to positive margins, and Technology also saw an increase. The largest drop is in Health Care, which fell sharply from a high pre-pandemic median to a much lower level. Cyclical sectors like Industrials, Construction, Arts & Entertainment, and Financials experienced small to moderate decreases, while Utilities and Education remain near flat. Using medians (rather than means) makes the comparison more robust to outliers and highlights typical firms within each industry.

### Question 2

**2. How did leverage (debt-to-assets) and liquidity (current ratio) evolve during pandemic?**

AAA

### Question 3
**3. Which German industries showed unexpected financial resilience or vulnerability from 2020 to 2022, and how were these outcomes shaped by profitability, leverage, liquidity, and firm size?**

AAA

### Question 4
**4. How did the fundamental relationship between corporate liquidity and profitability evolve and fracture within industries during the pandemic?**

AAA

## Conclusions



## Resources

The materials used for this report are:

Wickham H, Averick M, Bryan J, Chang W, McGowan LD, François R,
  Grolemund G, Hayes A, Henry L, Hester J, Kuhn M, Pedersen TL,
  Miller E, Bache SM, Müller K, Ooms J, Robinson D, Seidel DP, Spinu
  V, Takahashi K, Vaughan D, Wilke C, Woo K, Yutani H (2019).
  "Welcome to the tidyverse." _Journal of Open Source Software_,
  *4*(43), 1686. [doi:10.21105/joss.01686](https://doi.org/10.21105/joss.01686).

The links to my use of ChatGPT for help on this assignment are:

- [https://chatgpt.com/share/68857a22-debc-8001-bfcc-10552befba4c]() helped to XXX.

