---
title: "ETC5521 Diving Deeper into Data Exploration: Project Part 2"
author: "P2-wallaby"
date: "Oct 25, 2025"
quarto-required: ">=1.8.0"
format:
    html: 
        output-file: report_p2.html
        css: "assignment.css"
        theme: cerulean
    
        code-fold: true
        code-summary: "Show code"
        code-line-numbers: false
        code-tools: false
        code-copy: false
    
        #big size save
        embed-resources: false
        self-contained: false
    
        #trim JS helpers
        toc: true
        toc-depth: 3
        toc-location: left
        anchor-sections: false
        smooth-scroll: false
    
        #smaller images
        fig-width: 6
        fig-height: 4
        fig-dpi: 72
    
        #lighter code highlighting
        highlight-style: github
        df-print: paged
editor: 
  markdown: 
    wrap: sentence
---

```{r}
#| echo: false
# Set up chunk for all slides
knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 4,
  fig.align = "center",
  out.width = "100%",
  code.line.numbers = FALSE,
  fig.retina = 4,
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = FALSE,
  dev.args = list(pointsize = 11)
)
```

```{r}
#| echo: false
# Load libraries
library(tidyverse)
library(conflicted)

conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::select)
conflicts_prefer(dplyr::slice)

# Plot options and themes
options(
  digits = 2,
  width = 60,
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis",
  ggplot2.discrete.colour = c("#D55E00", "#0072B2", "#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442"),
  ggplot2.discrete.fill = c("#D55E00", "#0072B2", "#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442")
)

theme_set(theme_bw(base_size = 14) +
   theme(
     aspect.ratio = 1,
     plot.background = element_rect(fill = 'transparent', colour = NA),
     plot.title.position = "plot",
     plot.title = element_text(size = 24),
     panel.background = element_rect(fill = 'transparent', colour = NA),
     legend.background = element_rect(fill = 'transparent', colour = NA),
     legend.key = element_rect(fill = 'transparent', colour = NA)
   )
)
```

This is the detail library that we used for this project.

```{r load-library}
library(dplyr)
library(stringr)
library(janitor)
library(skimr)
library(tsibble)
library(knitr)
library(naniar)
library(patchwork)
library(plotly)
library(tsibbletalk)
library(feasts)
library(broom)
library(purrr)
library(cassowaryr)
library(gganimate)
library(gifski)
library(kableExtra)
library(ggplot2)
library(scales)
library(tidyr)
library(ggrepel)
library(viridisLite)
```

## Introduction

This report provides a comprehensive examination of financial performance across German and Swedish industries during the period 2018 to 2021. Drawing on firm-level data from the OSIRIS database, the analysis applies established financial ratios and summary statistics to assess profitability, liquidity, leverage, and firm size across a diverse range of sectors. The primary aim is to identify structural strengths and vulnerabilities within each country’s corporate environment, with a particular focus on how different industries responded to the onset and progression of the COVID-19 pandemic. By combining descriptive analytics with comparative industry insights, the report aims to inform understanding of financial resilience under conditions of economic uncertainty, highlighting notable patterns and sector-specific trends that emerged over this critical period.


## Data description

The dataset used in this analysis originates from **the Bureau van Dijk OSIRIS database**, a comprehensive source of harmonised financial information for publicly listed and major private firms worldwide (Bureau van Dijk, 2025).
For this study, we extracted firm-level data for **German companies** over the **2018–2021 accounting years**, a period that captures both pre-pandemic and pandemic-era financial performance.
Each record represents a firm’s reported financials for a specific accounting year, enabling longitudinal exploration of trends in profitability, liquidity, leverage, and resilience.
These are the code that we used to load the data and combine them:

Loading the data.

```{r read-data}
#define the folder
data_directory <- "osiris"

#filter germany and sweden
germany_sweden_file_paths <- list.files(
  data_directory,
  pattern = "^osiris_?(Germany|Sweden)_(2018|2019|2020|2021)\\.rda$",
  full.names = TRUE,
  ignore.case = TRUE
)

#name the list by file stems
germany_sweden_file_stems <- tools::file_path_sans_ext(basename(germany_sweden_file_paths))
osiris_germany_sweden <- setNames(vector("list", length(germany_sweden_file_paths)), germany_sweden_file_stems)

#load each file
for (file_index in seq_along(germany_sweden_file_paths)) {
  temporary_environment <- new.env(parent = emptyenv())
  load(germany_sweden_file_paths[file_index], envir = temporary_environment)
  loaded_objects <- as.list(temporary_environment)

  #If the file holds one object, store that object, otherwise store the sub-list
  osiris_germany_sweden[[file_index]] <- if (length(loaded_objects) == 1) loaded_objects[[1]] else loaded_objects
}

```

Combining the data.

```{r read-data-2}
#build a single data frame with 'year' and 'source_country' taken from the filename
combined_germany_sweden <- {
  parts <- vector("list", length(osiris_germany_sweden))
  names(parts) <- names(osiris_germany_sweden)

  for (file_stem in names(osiris_germany_sweden)) {
    data_frame_from_file <- osiris_germany_sweden[[file_stem]]

    extracted_year     <- as.integer(str_extract(file_stem, "\\d{4}"))
    extracted_country  <- str_extract(file_stem, "(?i)Germany|Sweden") |> stringr::str_to_title()

    parts[[file_stem]] <- mutate(
      data_frame_from_file,
      year = extracted_year,
      source_country = extracted_country #avoid clashing with the raw 'country'
    )
  }

  #bind and sort
  bind_rows(parts) |>
    arrange(source_country, year)
}

#unique country–year pairs
combined_germany_sweden |>
  dplyr::distinct(source_country, year) |>
  dplyr::arrange(source_country, year) |>
  kable(align = "l", booktabs = TRUE) |>
  kable_styling(full_width = FALSE, font_size = 11)

```

After the data is combined, we standardise the raw file into analysis-ready fields.
It normalises column names, maps long vendor headers to clear variables by topic, and creates a fiscal-year label from the year-end date.
Finally, it keeps only the identifiers, industry codes, and financial variables needed for the analysis.

```{r data-cleaning}
filtered_germany_sweden <- combined_germany_sweden |>
  clean_names() |>
  
  rename(
    # IDs / scope (used in all questions)
    company_name       = company_name_name,
    company_id         = name_id,
    acct_year          = year,
    country            = country_country,
    city               = city_city_city,
    consolidation_code = consolidation_code_consol_code,
    status             = status_status,
    
    # Industry codes (Q1, Q2, Q4)
    nace_code = nace_rev_1_core_code_cnacecd,
    icb_code  = industrial_classification_benchmark_icb,
    sic_code  = us_sic_core_code_csicuscde,
    
    # Scale / balance sheet (Q2, Q3)
    total_assets_eur      = total_assets_data13077,
    total_liabilities_eur = total_liabilities_and_debt_data14022,
    total_equity_eur      = total_shareholders_equity_data14041,
    
    # Revenue & profit (Q1, Q3)
    total_revenue_eur = total_revenues_data13004,
    net_sales_eur     = net_sales_data13002,
    gross_sales_eur   = gross_sales_data13000,
    net_income_eur    = net_income_starting_line_data15500,
    
    # Profitability ratios (Q1, Q4)
    ebit_margin_pct   = ebit_margin_percent_data31055,
    ebitda_margin_pct = ebitda_margin_percent_data31060,
    roa_pct           = return_on_total_assets_percent_data31015,
    roe_pct           = roe_percent_data31065,
    
    # Efficiency
    net_assets_turnover = net_assets_turnover_data31225,
    stock_turnover      = stock_turnover_data31220,
    
    # Leverage, solvency, and liquidity (Q2)
    solvency_pct        = solvency_ratio_percent_data31310,
    gearing_pct         = gearing_percent_data31315,
    current_ratio       = current_ratio_data31105,
    liquidity_ratio     = liquidity_ratio_data31110,
    shareholders_liq_pct= shareholders_liquidity_ratio_data31305,
    interest_cover      = interest_cover_data31115
  ) |>
  
  # Rename fiscal year
  mutate(
    fy_end_date = lubridate::ymd(as.character(company_fiscal_year_end_date_closdate)),
    fy_year     = year(fy_end_date)
  ) |>
  
  # Select only relevant columns
  select(
    company_id, company_name, fy_end_date, fy_year, acct_year, status, country, city, consolidation_code,
    nace_code, icb_code, sic_code,
    total_assets_eur, total_liabilities_eur, total_equity_eur,
    total_revenue_eur, net_sales_eur, gross_sales_eur, net_income_eur,
    ebit_margin_pct, ebitda_margin_pct, roa_pct, roe_pct,
    net_assets_turnover, stock_turnover,
    solvency_pct, gearing_pct, current_ratio, liquidity_ratio,
    shareholders_liq_pct, interest_cover
  )

```

This is the table describing the variables we used in this project:

```{r table-description, echo=FALSE}
data_dictionary <- tribble(
  ~`Column Name`,           ~Description,                                                                    ~`Data Type`, ~`Original Name`,
  "company_id",             "Unique company identifier (Bureau van Dijk ID).",                               "character", "BvD ID Number (os_id_number)",
  "company_name",           "Registered company name (label).",                                               "character", "Company Name (name)",
  "fy_end_date",            "Company fiscal year-end date for the account.",                                  "date",      "Company Fiscal Year End Date (closdate)",
  "fy_year",                "Year extracted from fiscal year-end date.",                                      "integer",   "derived from Company Fiscal Year End Date (closdate)",
  "acct_year",              "Reporting year label carried from file name/field.",                             "integer",   "year",
  "country",                "Country of head office.",                                                         "character", "Country (country)",
  "city",                   "City of head office.",                                                            "character", "CITY - City (city)",
  "consolidation_code",     "Consolidation scope of the accounts (e.g., C1/C2/U1/U2).",                        "character", "Consolidation Code (consol_code)",
  "nace_code",              "NACE core code (EU industry classification).",                                    "character", "NACE Rev 1, Core Code (cnacecd)",
  "icb_code",               "ICB industry classification code.",                                               "character", "Industrial Classification Benchmark (icb)",
  "sic_code",               "US SIC core code.",                                                               "character", "US SIC, Core Code (csicuscde)",
  "total_assets_eur",       "Total assets (EUR).",                                                              "numeric",   "Total Assets (data13077)",
  "total_liabilities_eur",  "Total liabilities and debt (EUR).",                                               "numeric",   "Total Liabilities and Debt (data14022)",
  "total_equity_eur",       "Total shareholders’ equity (EUR).",                                               "numeric",   "Total Shareholders Equity (data14041)",
  "total_revenue_eur",      "Total revenues / operating revenue (EUR).",                                       "numeric",   "Total revenues (data13004)",
  "net_sales_eur",          "Net sales (EUR).",                                                                 "numeric",   "Net sales (data13002)",
  "gross_sales_eur",        "Gross sales (EUR).",                                                               "numeric",   "Gross sales (data13000)",
  "net_income_eur",         "Net income (EUR).",                                                                "numeric",   "Net Income / Starting Line (data15500)",
  "ebit_margin_pct",        "EBIT margin (% of sales/revenue).",                                               "numeric",   "EBIT Margin (%) (data31055)",
  "ebitda_margin_pct",      "EBITDA margin (% of sales/revenue).",                                             "numeric",   "EBITDA Margin (%) (data31060)",
  "roa_pct",                "Return on total assets (%).",                                                     "numeric",   "Return on Total Assets (%) (data31015)",
  "roe_pct",                "Return on equity (%).",                                                           "numeric",   "ROE (%) (data31065)",
  "net_assets_turnover",    "Net assets turnover (times).",                                                    "numeric",   "Net Assets Turnover (data31225)",
  "stock_turnover",         "Stock / inventory turnover (times).",                                             "numeric",   "Stock Turnover (data31220)",
  "solvency_pct",           "Solvency ratio (%).",                                                             "numeric",   "Solvency ratio (%) (data31310)",
  "gearing_pct",            "Gearing ratio (%).",                                                              "numeric",   "Gearing (%) (data31315)",
  "current_ratio",          "Current ratio (times).",                                                          "numeric",   "Current ratio (data31105)",
  "liquidity_ratio",        "Liquidity / quick ratio (times).",                                                "numeric",   "Liquidity ratio (data31110)",
  "shareholders_liq_pct",   "Shareholders’ liquidity ratio (%).",                                              "numeric",   "Shareholders Liquidity ratio (data31305)",
  "interest_cover",         "Interest cover (times).",                                                         "numeric",   "Interest Cover (data31115)"
)

#show table
data_dictionary |>
  kable(align = "l", booktabs = TRUE) |>
  kable_styling(full_width = FALSE, font_size = 11)
```

The selected variables encompass four key dimensions of corporate financial health:

-   **Identification and Location** - company_id, company_name, country, and city provide firm-level identifiers and geographical context.

<!-- -->

-   **Balance Sheet Structure** - total_assets_eur, total_liabilities_eur, and total_equity_eur measure firm size and capital composition, reported in euros.

<!-- -->

-   **Income and Profitability** - total_revenue_eur, net_sales_eur, and net_income_eur capture income statement outcomes, while ebit_margin_pct, ebitda_margin_pct, roa_pct, and roe_pct reflect operating performance and returns to shareholders.

<!-- -->

-   **Solvency and Liquidity** - solvency_pct, gearing_pct, current_ratio, liquidity_ratio, shareholders_liq_pct, and interest_cover assess each firm’s financial stability, leverage, and capacity to meet short-term obligations.

-   **Industry Classification** - industry_nace, industry_icb, and industry_group link firms to their respective economic sectors according to European (NACE) and global (ICB) coding systems.

It is important to note that the **accounting year** variable in **OSIRIS** reflects **two-year reporting periods**.
For example, the 2018 accounting year represents financial results spanning 2017–2018, 2019 corresponds to 2018–2019, 2020 covers 2019–2020, and 2021 includes 2020–2021.
Throughout this report, each accounting year is interpreted as representing the **endpoint of its reporting cycle**.
Consequently, “2021” is treated as capturing firm performance during the height of the **COVID-19 pandemic**, while 2018–2019 represent pre-pandemic benchmarks.

This dataset structure enables a robust exploration of temporal changes in corporate performance and the role of firm characteristics , particularly **firm size** , in shaping **financial resilience** throughout the pandemic period.

### Research Question

Here are the research questions (sub-questions) that guide us in analysing and answering the primary question: How did the pandemic affect corporate financials in Germany and Sweden?

**1. How did profitability change across industries?**

This question investigates how profitability variables such as ROA, EBITDA margin, and EBIT margin shifted across industries between the pre-pandemic period (2018 to 2019) and the pandemic period (2020 to 2021) in Germany and Sweden.
It aims to identify which sectors saw profit declines, remained stable, or grew during the pandemic.
By comparing profitability trends across industries, the analysis highlights which sectors proved more resilient to the economic impacts of the pandemic.
These insights help explain how the pandemic affected overall corporate performance and provide a clearer understanding of the differences in financial recovery among industries.

**2. How did leverage (debt-to-assets) and liquidity (current ratio) evolve during pandemic?**

This question extends the Part 1 analysis by comparing **how leverage (debt-to-assets)** and **liquidity (current ratio)** evolved between **Germany and Sweden** from 2018 to 2021.\
The goal is to determine whether firms in the two economies displayed similar balance-sheet responses to the COVID-19 shock.\
This section applies techniques from **Lectures 4-6** (data summarisation, comparative visualisation, and faceting) and interprets results within a financial-resilience framework.

**3. Which Sweden industries showed unexpected financial resilience or vulnerability from 2019 to 2021, and how were these outcomes shaped by profitability, leverage, liquidity, and firm size?**

Between 2019 and 2021, Swedish industries showed mixed financial resilience. Manufacturing, IT, and professional services remained strong, maintaining profitability and liquidity despite the pandemic, helped by Sweden’s lighter restrictions and export strength. In contrast, hospitality, transport, and retail were more vulnerable, with sharp declines in profitability and higher leverage as firms relied on debt to survive. Larger firms recovered faster due to stronger cash reserves and financing access, while smaller firms faced liquidity pressures. Overall, industries with high profitability and low leverage before 2020 proved most resilient through 2021.

**4. How did the fundamental relationship between corporate liquidity and profitability evolve and fracture within industries during the pandemic?**

This study compares firms in **Germany** and **Sweden**, two advanced European economies with different financial systems and policy responses.
The analysis explores how the ROA–CR relationship evolved and fractured within industries during and after the pandemic using time-series features and scagnostic metrics.
**ROA** and **Current Ratio** reflect how efficiently firms generate profit and how safely they can cover short-term obligations.
The goal is to reveal which sectors were most resilient, which experienced unequal recovery, and how national environments influenced these patterns.
Understanding the changing relationship between ROA and CR can provide insight into the financial resilience of different industries.
It also helps identify whether recovery followed a stable path or diverged into unequal outcomes, such as K-shaped patterns, where some firms recovered faster while others declined.

## IDA

Here is the IDA process for this project:

First, we develop this code to prepare the dataset for analysis by converting numeric columns, standardising text fields, and reporting missing values overall and by year.
It also performs basic sanity checks for duplicate firm-years, impossible values, and consolidation codes.

```{r table-checking}

num_cols <- c(
  "total_assets_eur","total_liabilities_eur","total_equity_eur",
  "total_revenue_eur","net_sales_eur","gross_sales_eur","net_income_eur",
  "ebit_margin_pct","ebitda_margin_pct","roa_pct","roe_pct",
  "net_assets_turnover","stock_turnover",
  "solvency_pct","gearing_pct","current_ratio","liquidity_ratio",
  "shareholders_liq_pct","interest_cover"
)

#symbol formatting/coerce numerics
filtered_germany_sweden <- filtered_germany_sweden |>
  mutate(
    across(all_of(num_cols) & where(is.character), readr::parse_number),
    country = stringr::str_to_title(country),
    city    = stringr::str_to_title(city)
  )

#missingness report check (overall and by year)
missing_overall <- filtered_germany_sweden |>
  summarise(across(everything(), ~ mean(is.na(.)))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "missing_rate")

missing_by_year <- filtered_germany_sweden |>
  group_by(acct_year) |>
  summarise(across(everything(), ~ mean(is.na(.)))) |>
  pivot_longer(-acct_year, names_to = "variable", values_to = "missing_rate")

#sanity checks
dup_firm_year <- filtered_germany_sweden |>
  count(company_id, fy_year, acct_year, country) |>
  filter(n > 1)

impossible_assets <- filtered_germany_sweden |>
  filter(!is.na(total_assets_eur) & total_assets_eur < 0)

consol_levels <- filtered_germany_sweden |>
  count(consolidation_code, sort = TRUE)

```

Then, we create this code to extract the 2-digit NACE code, categorises companies into broad industry groups using NACE and ICB, and creates a unified 'industry_group' factor with a fixed order for plotting purposes.
It also provides a small sample of unmapped records for verification.
After checking, there are no duplicates in the data using company_id, fy_year, and acct_year as the primary key.

```{r NACE-column}
#clean 2-digit NACE as integer
filtered_germany_sweden <- filtered_germany_sweden |>
  mutate(
    nace_num2 = suppressWarnings(as.integer(str_sub(readr::parse_number(as.character(nace_code)), 1, 2)))
  )

#map NACE divisions to broad section names
industry_from_nace <- function(x){
  case_when(
    x %in% 1:3 ~ "Agriculture, Forestry & Fishing",
    x %in% 5:9 ~ "Mining & Quarrying",
    x %in% 10:33 | x %in% 15:37 ~ "Manufacturing",
    x %in% 35 ~ "Electricity, Gas, Steam",
    x %in% 36:39 ~ "Water Supply & Waste",
    x %in% 41:43 ~ "Construction",
    x %in% 45:47 ~ "Wholesale & Retail Trade",
    x %in% 49:53 ~ "Transport & Storage",
    x %in% 55:56 ~ "Accommodation & Food",
    x %in% 58:63 ~ "Information & Communication",
    x %in% 64:66 ~ "Financial & Insurance",
    x %in% 68 ~ "Real Estate",
    x %in% 69:75 ~ "Professional, Scientific & Technical",
    x %in% 77:82 ~ "Administrative & Support",
    x %in% 84 ~ "Public Administration",
    x %in% 85 ~ "Education",
    x %in% 86:88 ~ "Human Health & Social Work",
    x %in% 90:93 ~ "Arts, Entertainment & Recreation",
    x %in% 94:96 ~ "Other Service Activities",
    x %in% 97:98 ~ "Household Activities",
    x %in% 99 ~ "Extraterritorial Organizations",
    TRUE ~ NA_character_
  )
}

#ICB top-level sector labels
icb_map <- c(
  "0001" = "Oil & Gas",
  "1000" = "Basic Materials",
  "2000" = "Industrials",
  "3000" = "Consumer Goods",
  "4000" = "Health Care",
  "5000" = "Consumer Services",
  "6000" = "Telecommunications",
  "7000" = "Utilities",
  "8000" = "Financials",
  "9000" = "Technology"
)

#normalize ICB to a 4-digit “bucket” (example: 2573 to 2000)
normalize_icb <- function(x){
  x_chr <- str_extract(as.character(x), "\\d+")
  ifelse(is.na(x_chr), NA_character_,
         sprintf("%04d", as.integer(floor(as.numeric(x_chr) / 1000) * 1000)))
}

filtered_germany_sweden <- filtered_germany_sweden |>
  mutate(
    industry_nace = industry_from_nace(nace_num2),
    icb_bucket    = normalize_icb(icb_code),
    industry_icb  = icb_map[icb_bucket],
    industry_group = coalesce(industry_nace, industry_icb, "Other / Unmapped")
  )

#check
unmapped_sample <- filtered_germany_sweden |>
  filter(industry_group == "Other / Unmapped") |>
  select(company_id, fy_year, acct_year, nace_code, icb_code) |>
  head(15)

#factor order for plots
ordered_levels <- c(
  "Agriculture, Forestry & Fishing","Mining & Quarrying","Manufacturing",
  "Electricity, Gas, Steam","Water Supply & Waste","Construction",
  "Wholesale & Retail Trade","Transport & Storage","Accommodation & Food",
  "Information & Communication","Financial & Insurance","Real Estate",
  "Professional, Scientific & Technical","Administrative & Support",
  "Public Administration","Education","Human Health & Social Work",
  "Arts, Entertainment & Recreation","Other Service Activities",
  "Household Activities","Extraterritorial Organizations",
  "Oil & Gas","Basic Materials","Industrials","Consumer Goods","Health Care",
  "Consumer Services","Telecommunications","Utilities","Financials","Technology",
  "Other / Unmapped"
)
filtered_germany_sweden <- filtered_germany_sweden |>
  mutate(industry_group = factor(industry_group, levels = ordered_levels))
```

After that, we create this code to reshape numeric variables into a long format for Initial Data Analysis, separate scale variables from ratios, and assign clear labels for charts.
The result is a tidy table ready for profiling and visualisation.

```{r prepare-IDA}
#using long data
ida_germany_sweden <- filtered_germany_sweden |>
  dplyr::select(acct_year, country, dplyr::all_of(num_cols)) |>
  tidyr::pivot_longer(dplyr::all_of(num_cols),
                      names_to = "variable_orig", values_to = "value") |>
  tidyr::drop_na()

#split between scale and ratio
scale_vars <- c("total_assets_eur","total_liabilities_eur","total_equity_eur",
                "total_revenue_eur","net_sales_eur","gross_sales_eur","net_income_eur")
ratio_vars <- setdiff(unique(ida_germany_sweden$variable_orig), scale_vars)

#label mapping for the charts
nice <- c(
  total_assets_eur   = "Total assets (EUR)",
  total_liabilities_eur="Total liabilities (EUR)",
  total_equity_eur   = "Equity (EUR)",
  total_revenue_eur  = "Total revenue (EUR)",
  net_sales_eur      = "Net sales (EUR)",
  gross_sales_eur    = "Gross sales (EUR)",
  net_income_eur     = "Net income (EUR)",
  ebit_margin_pct    = "EBIT margin (%)",
  ebitda_margin_pct  = "EBITDA margin (%)",
  roa_pct            = "ROA (%)",
  roe_pct            = "ROE (%)",
  net_assets_turnover= "Net assets turnover (x)",
  stock_turnover     = "Stock turnover (x)",
  solvency_pct       = "Solvency (%)",
  gearing_pct        = "Gearing (%)",
  current_ratio      = "Current ratio (x)",
  liquidity_ratio    = "Liquidity ratio (x)",
  shareholders_liq_pct = "Shareholders’ liquidity (%)",
  interest_cover     = "Interest cover (x)"
)

pretty_var <- function(x) ifelse(x %in% names(nice), nice[x], x)

ida_germany_sweden <- ida_germany_sweden |>
  mutate(variable = pretty_var(variable_orig),
         variable = factor(variable, levels = unique(variable)))

```

Then, we create this code to visualise the distribution of scale variables by year using violin and jitter plots on a log scale, with medians highlighted.
It allows quick checks of spread, skewness, and outliers over time for monetary measures.

```{r IDA-scale}
#define variables sets
vars_balance <- c("total_assets_eur","total_liabilities_eur","total_equity_eur")
vars_flows   <- c("total_revenue_eur","net_sales_eur","gross_sales_eur","net_income_eur")
```

**Balanced Sheet Scale Variables**

```{r balanced-sheet-scale}
p_scale_balance <- ida_germany_sweden |>
  filter(variable_orig %in% vars_balance) |>
  mutate(
    y    = if_else(value > 0, value, NA_real_),
    year = factor(acct_year)
  ) |>
  ggplot(aes(x = year, y = y, colour = country,
             group = interaction(country, year))) +
  geom_violin(trim = FALSE, fill = NA, linewidth = 0.4, na.rm = TRUE,
              position = position_dodge(width = 0.75)) +
  geom_jitter(alpha = 0.12, size = 0.6, na.rm = TRUE,
              position = position_jitterdodge(jitter.width = 0.12,
                                              jitter.height = 0,
                                              dodge.width = 0.75)) +
  stat_summary(fun = median, geom = "point", shape = 95, size = 6, na.rm = TRUE,
               position = position_dodge(width = 0.75)) +
  scale_y_log10(labels = scales::label_number(scale_cut = scales::cut_short_scale())) +
  facet_wrap(~ variable, scales = "free_y", ncol = 3) +
  labs(x = "Year", y = NULL, title = "Violin and Jitter by Year (log scale) for Balance Sheet Variables") +
  theme_minimal(base_size = 11) +
  theme(
  plot.title.position = "plot",
  legend.position     = "top",
  legend.spacing.y    = unit(-0.6, "lines"),
  legend.justification = "center",
  legend.direction     = "horizontal",
  legend.box           = "horizontal",
  legend.margin        = margin(b = 4),
  axis.text.x          = element_text(size = 10),
  strip.text           = element_text(face = "bold"),
  panel.grid.minor     = element_blank()
)

p_scale_balance
```

According to the chart, most companies in both countries are at the lower end of the asset and liability spectrum, while a small number of very large firms form the long upper tails, indicating strong right skewness.
The medians for total assets, liabilities, and equity remain stable over time, suggesting no significant structural shift in balance sheets between 2018 and 2021.
Sweden consistently reports higher levels across the balance-sheet variables than Germany, reflecting generally larger firm sizes in the Swedish sample.
The range widens in 2020 – 2021 for both countries, showing increased variation during the pandemic.
Median equity remains positive and steady across both nations, suggesting that most firms maintained solvency despite the disruptions.

**Flow Scale Variables**

```{r}
p_scale_flows <- ida_germany_sweden |>
  dplyr::filter(variable_orig %in% vars_flows) |>
  dplyr::mutate(
    y    = dplyr::if_else(value > 0, value, NA_real_),
    year = factor(acct_year)
  ) |>
  ggplot(aes(x = year, y = y, colour = country,
             group = interaction(country, year))) +
  geom_violin(trim = FALSE, fill = NA, linewidth = 0.4, na.rm = TRUE,
              position = position_dodge(width = 0.75)) +
  geom_jitter(alpha = 0.12, size = 0.6, na.rm = TRUE,
              position = position_jitterdodge(jitter.width = 0.12,
                                              jitter.height = 0,
                                              dodge.width = 0.75)) +
  stat_summary(fun = median, geom = "point", shape = 95, size = 6, na.rm = TRUE,
               position = position_dodge(width = 0.75)) +
  scale_y_log10(labels = scales::label_number(scale_cut = scales::cut_short_scale())) +
  facet_wrap(~ variable, scales = "free_y", ncol = 2) +
  labs(x = "Year", y = NULL, title = "Violin and Jitter by Year (log scale) for Flow Variables") +
  theme_minimal(base_size = 11) +
  theme(
  plot.title.position = "plot",
  legend.position     = "top",
  legend.spacing.y    = unit(0.3, "lines"),
  legend.justification = "center",
  legend.direction     = "horizontal",
  legend.box           = "horizontal",
  legend.margin        = margin(t = 6, b = 2),
  axis.text.x          = element_text(size = 10),
  strip.text           = element_text(face = "bold"),
  panel.grid.minor     = element_blank()
)

p_scale_flows
```

According to the chart, he flow measures are strongly right-skewed in both countries: most firms are small to mid-sized, with a few very large firms creating long upper tails.
Sweden ranks above Germany across all years, with higher medians for total revenue, net/gross sales, and net income.
Distributions expand in 2020 – 2021 for both countries.
Additionally, the most significant change is in net income, which shows a thicker lower tail, indicating more firms with very small profits or losses during the pandemic.
Median revenues and sales remain broadly stable over 2018 – 2021, with only modest year-to-year shifts compared to the cross-country gap.
Due to skew and extreme values, the median and IQR (not the mean) are the appropriate summaries, and any modelling should consider robust or winsorised treatments of the tails.

After that, we visualise ratio and percentage metrics by year on a linear scale using violin and jitter plots with median markers, faceted by variable.
It is used to assess distribution shape, central tendency, spread, and potential outliers across time without log transformation.

```{r IDA-ratio}
#define variable sets
vars_profit_eff <- c(
  "ebit_margin_pct", "ebitda_margin_pct",
  "roa_pct", "roe_pct",
  "net_assets_turnover", "stock_turnover"
)

vars_liquidity_solvency <- c(
  "solvency_pct", "gearing_pct",
  "current_ratio", "liquidity_ratio",
  "shareholders_liq_pct", "interest_cover"
)
```

**Profit and Efficiency Ratio Variables**

```{r p-ratio-profit-eff}
p_ratio_profit_eff <- ida_germany_sweden |>
  dplyr::filter(variable_orig %in% vars_profit_eff) |>
  dplyr::mutate(year = factor(acct_year)) |>
  ggplot(aes(x = year, y = value, colour = country,
             group = interaction(country, year))) +
  geom_violin(trim = FALSE, fill = NA, linewidth = 0.4, na.rm = TRUE,
              position = position_dodge(width = 0.75)) +
  geom_jitter(alpha = 0.12, size = 0.6, na.rm = TRUE,
              position = position_jitterdodge(jitter.width = 0.12,
                                              jitter.height = 0,
                                              dodge.width = 0.75)) +
  stat_summary(fun = median, geom = "point",
               shape = 95, size = 6, na.rm = TRUE,
               position = position_dodge(width = 0.75)) +
  facet_wrap(~ variable, scales = "free_y", ncol = 3) +
  labs(x = "Year", y = NULL,
       title = "Violin and Jitter by Year for Profitability & Efficiency Ratios") +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_text(size = 10),
    strip.text  = element_text(face = "bold"),
    panel.grid.minor = element_blank(),
    legend.position = "top"
  )

p_ratio_profit_eff
```

According to the chart, the profitability ratios (EBIT margin, EBITDA margin, and ROA) are mainly close to zero for most firms in both countries, with many observations slightly positive and some negative.
The distributions become wider in 2020 – 2021, indicating increased volatility during the pandemic, especially in ROA, where more firms became negative.
ROE shows the greatest extent of variation, including very large negative values, suggesting that equity returns were more affected by shocks.
Net asset turnover and stock turnover remain relatively stable over time, suggesting operational efficiency did not change much during the pandemic.
Sweden consistently has a broader spread than Germany in these ratios, reflecting greater firm-level variation.
Due to the heavy tails and outliers, medians and distribution shapes are more useful than means.

**Liquidity and Solvency Ratio Variables**

```{r}
p_ratio_liq_solv <- ida_germany_sweden |>
  dplyr::filter(variable_orig %in% vars_liquidity_solvency) |>
  dplyr::mutate(year = factor(acct_year)) |>
  ggplot(aes(x = year, y = value, colour = country,
             group = interaction(country, year))) +
  geom_violin(trim = FALSE, fill = NA, linewidth = 0.4, na.rm = TRUE,
              position = position_dodge(width = 0.75)) +
  geom_jitter(alpha = 0.12, size = 0.6, na.rm = TRUE,
              position = position_jitterdodge(jitter.width = 0.12,
                                              jitter.height = 0,
                                              dodge.width = 0.75)) +
  stat_summary(fun = median, geom = "point",
               shape = 95, size = 6, na.rm = TRUE,
               position = position_dodge(width = 0.75)) +
  facet_wrap(~ variable, scales = "free_y", ncol = 3) +
  labs(x = "Year", y = NULL,
       title = "Violin and Jitter by Year for Liquidity & Solvency Variables") +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_text(size = 10),
    strip.text  = element_text(face = "bold"),
    panel.grid.minor = element_blank(),
    legend.position = "top"
  )

p_ratio_liq_solv
```

Solvency ratios for both countries generally range from 40% to 60%, indicating that most firms maintain moderate long-term stability.
Sweden's median solvency ratios are slightly higher than Germany's, suggesting stronger capital structures.
Gearing ratios are low for the majority of firms but exhibit long right tails, reflecting a few highly leveraged cases in both countries.
Liquidity and current ratios are close to 1, consistent with tight short-term asset coverage, although some firms hold extremely high liquidity buffers.
Shareholders’ liquidity and interest coverage show wide spreads, with a few large positive outliers indicating firms with strong financial reserves or low debt burdens.
Overall, the distributions are heavily skewed, suggesting that medians and IQRs provide more meaningful insights than means for assessing firm solvency and liquidity patterns.

Then, we measure missing data for key numeric variables by industry and by year, then visualise year-wise gaps with a heatmap.
It highlights variables or periods with substantial missingness, guiding imputation, exclusion, or robustness checks before analysis.

```{r numeric-missingness-check}
key_vars <- c("total_assets_eur","total_liabilities_eur","total_equity_eur",
              "total_revenue_eur","net_income_eur",
              "ebit_margin_pct","ebitda_margin_pct","roa_pct","roe_pct",
              "current_ratio","liquidity_ratio","gearing_pct","solvency_pct")

# missingness by industry and country
missing_by_industry <- filtered_germany_sweden |>
  group_by(country, industry_group) |>
  summarise(across(all_of(key_vars), ~ mean(is.na(.)), .names = "{.col}"),
            .groups = "drop") |>
  pivot_longer(-c(country, industry_group),
               names_to = "variable", values_to = "missing_rate")

#heatmap by year for numeric variables
missing_heat <- filtered_germany_sweden |>
  group_by(country, acct_year) |>
  summarise(across(all_of(key_vars), ~ mean(is.na(.)), .names = "{.col}"),
            .groups = "drop") |>
  pivot_longer(-c(country, acct_year),
               names_to = "variable", values_to = "missing_rate")

ggplot(missing_heat,
       aes(variable, factor(acct_year), fill = missing_rate)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "white", high = "red",
                      labels = scales::percent_format(accuracy = 1)) +
  facet_wrap(~ country, nrow = 1) +
  labs(x = NULL, y = "Year",
       title = "Missingness by Year & Country (Numeric Variables)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

The heatmap indicates that missingness patterns remain consistent across 2018 – 2021, with no significant changes during the pandemic.
In both Germany and Sweden, the highest levels of missing data are found in gearing_pct and liquidity_ratio, reaching about 20 – 35%, followed by ebit_margin_pct and ebitda_margin_pct, which have moderate gaps around 10 – 20%.
Sweden consistently exhibits slightly higher missingness rates than Germany, especially for gearing and liquidity indicators.
Core balance-sheet and flow variables such as assets, liabilities, equity, and revenues are nearly complete, with almost no missing values.
Overall, ratio-based metrics are the primary source of data gaps, meaning that profitability and liquidity analyses will be based on smaller samples compared to total-level variables.

After that, we used this code to detect outliers in all numeric variables with two rules - IQR (1.5×) and standardized scores (\|z\| \> 3) - and summarise counts and proportions per variable, ordered by the IQR share.
It flags variables with heavy tails or extreme values to guide winsorisation, transformation, or sensitivity checks before modelling.

```{r outliers-check}
num_cols <- c("total_assets_eur","total_liabilities_eur","total_equity_eur",
              "total_revenue_eur","net_sales_eur","gross_sales_eur","net_income_eur",
              "ebit_margin_pct","ebitda_margin_pct","roa_pct","roe_pct",
              "net_assets_turnover","stock_turnover",
              "solvency_pct","gearing_pct","current_ratio","liquidity_ratio",
              "shareholders_liq_pct","interest_cover")

qfun <- function(x,p) quantile(x, probs=p, na.rm=TRUE, names=FALSE)

outliers_by_country <- purrr::map_dfr(num_cols, function(v){
  x  <- filtered_germany_sweden[[v]]
  ct <- filtered_germany_sweden$country
  
  q1 <- qfun(x,.25); q3 <- qfun(x,.75); iqr <- q3-q1
  lo <- q1 - 1.5*iqr; hi <- q3 + 1.5*iqr
  z  <- as.numeric(scale(x))
  
  tibble::tibble(
    variable = v,
    country  = ct,
    n        = !is.na(x),
    iqr_flag = (x < lo | x > hi),
    z_flag   = abs(z) > 3
  )
}) |>
  dplyr::group_by(variable, country) |>
  dplyr::summarise(
    n = sum(n),
    iqr_outliers = sum(iqr_flag, na.rm = TRUE),
    z_outliers   = sum(z_flag, na.rm = TRUE),
    share_iqr    = iqr_outliers/n,
    share_z      = z_outliers/n,
    .groups = "drop"
  ) |>
  dplyr::arrange(country, desc(share_iqr))

var_levels <- intersect(num_cols, unique(outliers_by_country$variable))

outliers_by_country_ord <- outliers_by_country |>
  dplyr::mutate(variable = factor(variable, levels = var_levels)) |>
  dplyr::arrange(variable, country)

outliers_by_country_ord |>
  kable(align = "l", booktabs = TRUE) |>
  kable_styling(full_width = FALSE, font_size = 11)
```

According to the table, outliers are common in scale variables such as assets, liabilities, equity, and revenues, with 13 – 22% of observations falling outside the IQR range, reflecting the broad firm-size distribution.
Sweden consistently shows a higher proportion of outliers than Germany across most variables, aligning with its wider value dispersion seen in violin plots.
Profitability ratios (EBIT, EBITDA, ROA, ROE) also display many outliers.
Notably, ROE and EBITDA margins have extreme positive or negative values due to small equity or volatile income.
Liquidity and solvency ratios like gearing_pct, liquidity_ratio, and interest_cover have narrower but still significant tails, indicating some firms hold exceptionally high debt or cash reserves.
Overall, these findings confirm the presence of heavy-tailed distributions, justifying the later use of median-based summaries, log transformations, or winsorisation for more robust comparison and modelling.

Given these outlier patterns, the next step is to transform or winsorise selected variables (particularly profitability and liquidity ratios) to reduce the impact of extreme values and improve comparability across firms and periods before proceeding with further EDA and modelling.

Then, we standardise key character fields by trimming whitespace, normalising case, and converting blanks to NA, and report missingness for identifiers and classification codes.
This ensures consistent keys for joins and grouping, and highlights data quality issues to address before analysis.

```{r character-check}
#character variables to check
char_vars <- c("company_id","company_name","country","city",
               "consolidation_code","nace_code","icb_code","sic_code")
char_vars <- intersect(char_vars, names(filtered_germany_sweden))

#cleaning
chars_clean <- filtered_germany_sweden |>
  mutate(
    across(all_of(char_vars),
           ~ .x |> as.character() |> str_squish() |> na_if("")),
    #casing
    country = str_to_title(country),
    city    = str_to_title(city),
    company_name = str_squish(company_name)
  )

#missingness check
char_missing <- chars_clean |>
  summarise(across(all_of(char_vars), ~ mean(is.na(.)))) |>
  tidyr::pivot_longer(everything(), names_to="variable", values_to="missing_rate")

char_missing |>
  kable(align = "l", booktabs = TRUE) |>
  kable_styling(full_width = FALSE, font_size = 11)
```

According to the table, all key identifiers and location fields are complete, with no missing data in company_id, company_name, country, city, or consolidation_code.
This confirms that entity tracking and data joins can be performed reliably without risk of key mismatches.
Industry classifications are also mostly complete: nace_code shows only about 2% missing values, while icb_code has around 9% missingness, making it the least complete among the coding systems.
sic_code is fully populated, ensuring that every firm can still be categorised under at least one industrial classification.
Overall, this pattern indicates good metadata quality, and most firms can be consistently classified using either SIC or NACE as the primary identifiers, with ICB serving as a secondary, more detailed but less complete classification.

After completing the general IDA, we developed the specific IDA to address our sub-questions.
Here are the details:

### Question 1

We initially added a period label (Pre-pandemic 2018 – 2019 vs. Pandemic 2020 – 2021).
For visualisations, net_income_eur was transformed using asinh() to compare both gains and losses on a consistent scale.
When ROA/ROE are missing, we recalculated simple estimates from the accounting totals and noted the sources of the values (reported vs. recomputed).

```{r q1_prepare_profitability_inputs}
stopifnot("net_income_eur" %in% names(filtered_germany_sweden))

ida_question1 <- filtered_germany_sweden |>
  mutate(
    country = country,
    # period flag
    period = if_else(acct_year >= 2020, "Pandemic (2020–2021)", "Pre-pandemic (2018–2019)"),
    # transform for visuals (handles negatives)
    net_income_eur_asinh = asinh(as.numeric(net_income_eur)),
    # recompute ROA / ROE (fallbacks)
    roa_calc = 100 * (net_income_eur / total_assets_eur),
    roe_calc = if_else(total_equity_eur > 0,
                       100 * (net_income_eur / total_equity_eur),
                       NA_real_),
    roa_use = coalesce(roa_pct, roa_calc),
    roe_use = coalesce(roe_pct, roe_calc),
    roa_src = case_when(
      !is.na(roa_pct) ~ "reported",
      !is.na(roa_calc) ~ "recomputed",
      TRUE ~ NA_character_
    ),
    roe_src = case_when(
      !is.na(roe_pct) ~ "reported",
      !is.na(roe_calc) ~ "recomputed",
      TRUE ~ NA_character_
    )
  )
```

Then, we carried out a conservative within-firm, within-period “one-gap” fill, only patching a single missing year when adjacent values match.
It then restored reported values where available.

```{r q1_firm_one_gap_fill}
#Firm one-gap fill within the same period only

fill_one_gap <- function(x){
  f <- dplyr::lag(x); b <- dplyr::lead(x)
  ifelse(is.na(x) & !is.na(f) & !is.na(b) & (f == b), f, x)
}

ida_question1 <- ida_question1 |>
  arrange(company_id, acct_year) |>
  group_by(company_id, period) |>  #period boundary blocks carryover
  mutate(
    ebit_step1   = fill_one_gap(ebit_margin_pct),
    ebitda_step1 = fill_one_gap(ebitda_margin_pct),
    roa_step1    = fill_one_gap(roa_use),
    roe_step1    = fill_one_gap(roe_use)
  ) |>
  ungroup() |>
  mutate(
    ebit_step1   = coalesce(ebit_margin_pct, ebit_step1),
    ebitda_step1 = coalesce(ebitda_margin_pct, ebitda_step1),
    roa_step1    = coalesce(roa_use, roa_step1),
    roe_step1    = coalesce(roe_use, roe_step1)
  )
```

After that, we imputed remaining gaps following a strict hierarchy: first industry and year, then industry and period, and finally period, without crossing period boundaries.
This process produced the final Q1 variables (EBIT, EBITDA, ROA, ROE) and records the source of each value.

```{r q1_impute_hierarchy_finalize}
#median fills in strict order (no cross-period borrowing)
#country x industry × year medians
med_ixy <- ida_question1 |>
  group_by(country, industry_group, acct_year) |>
  summarise(
    med_ebit = median(ebit_step1,   na.rm=TRUE),
    med_ebitda = median(ebitda_step1, na.rm=TRUE),
    med_roa = median(roa_step1,    na.rm=TRUE),
    med_roe = median(roe_step1,    na.rm=TRUE),
    .groups="drop"
  )

#country x industry × period medians
med_ip <- ida_question1 |>
  group_by(country, industry_group, period) |>
  summarise(
    med_ebit_ip = median(ebit_step1,   na.rm=TRUE),
    med_ebitda_ip = median(ebitda_step1, na.rm=TRUE),
    med_roa_ip = median(roa_step1,    na.rm=TRUE),
    med_roe_ip = median(roe_step1,    na.rm=TRUE),
    .groups="drop"
  )

#country x period medians
med_p <- ida_question1 |>
  group_by(country, period) |>
  summarise(
    med_ebit_p = median(ebit_step1,   na.rm=TRUE),
    med_ebitda_p = median(ebitda_step1, na.rm=TRUE),
    med_roa_p = median(roa_step1,    na.rm=TRUE),
    med_roe_p = median(roe_step1,    na.rm=TRUE),
    .groups="drop"
  )

#join and impute with precedence, also explicit source flags
ida_question1 <- ida_question1 |>
  left_join(med_ixy, by = c("country","industry_group","acct_year")) |>
  left_join(med_ip,  by = c("country","industry_group","period")) |>
  left_join(med_p,   by = c("country", "period")) |>
  mutate(
    #final values, using your precedence
    ebit_margin_q1   = if_else(is.na(ebitda_step1), ebit_step1, ebit_step1),
    ebit_margin_q1   = if_else(is.na(ebit_step1),   coalesce(med_ebit,   med_ebit_ip,   med_ebit_p),   ebit_step1),
    ebitda_margin_q1 = if_else(is.na(ebitda_step1), coalesce(med_ebitda, med_ebitda_ip, med_ebitda_p), ebitda_step1),
    roa_q1           = if_else(is.na(roa_step1),    coalesce(med_roa,    med_roa_ip,    med_roa_p),    roa_step1),
    roe_q1           = if_else(is.na(roe_step1),    coalesce(med_roe,    med_roe_ip,    med_roe_p),    roe_step1),

    #provenance for EBIT/EBITDA margins
    ebit_src = dplyr::case_when(
      !is.na(ebit_margin_pct) ~ "reported",
      is.na(ebit_margin_pct) & !is.na(ebit_step1) ~ "firm-1gap(period)",
      is.na(ebit_step1) & !is.na(med_ebit) ~ "ind×year",
      is.na(ebit_step1) & is.na(med_ebit) & !is.na(med_ebit_ip) ~ "ind×period",
      TRUE ~ "period"
    ),
    ebitda_src = dplyr::case_when(
      !is.na(ebitda_margin_pct) ~ "reported",
      is.na(ebitda_margin_pct) & !is.na(ebitda_step1) ~ "firm-1gap(period)",
      is.na(ebitda_step1) & !is.na(med_ebitda) ~ "ind×year",
      is.na(ebitda_step1) & is.na(med_ebitda) & !is.na(med_ebitda_ip) ~ "ind×period",
      TRUE ~ "period"
    ),

    #ROA/ROE provenance
    roa_src = dplyr::case_when(
      !is.na(roa_src) ~ roa_src,
      is.na(roa_src) & !is.na(roa_step1) ~ "firm-1gap(period)",
      is.na(roa_step1) & !is.na(med_roa) ~ "ind×year",
      is.na(roa_step1) & is.na(med_roa) & !is.na(med_roa_ip) ~ "ind×period",
      TRUE ~ "period"
    ),
    roe_src = dplyr::case_when(
      !is.na(roe_src) ~ roe_src,
      is.na(roe_src) & !is.na(roe_step1) ~ "firm-1gap(period)",
      is.na(roe_step1) & !is.na(med_roe) ~ "ind×year",
      is.na(roe_step1) & is.na(med_roe) & !is.na(med_roe_ip) ~ "ind×period",
      TRUE ~ "period"
    )
  ) |>
  select(-ends_with("_step1"))
```

Then, we summarised imputation provenance by period, counting observations that are reported, firm-filled, or imputed from industry/year/period medians.
It provided a transparent audit of data construction for Question 1.

```{r q1_imputation_summary}
impute_summary_q1 <- ida_question1 |>
  summarise(
    .by = c(country, period),
    n_rows = n(),
    ebit_reported   = sum(ebit_src   == "reported",            na.rm=TRUE),
    ebit_firmgap    = sum(ebit_src   == "firm-1gap(period)",   na.rm=TRUE),
    ebit_ind_year   = sum(ebit_src   == "ind×year",            na.rm=TRUE),
    ebit_ind_period = sum(ebit_src   == "ind×period",          na.rm=TRUE),
    ebit_period     = sum(ebit_src   == "period",              na.rm=TRUE),

    ebitda_reported   = sum(ebitda_src   == "reported",          na.rm=TRUE),
    ebitda_firmgap    = sum(ebitda_src   == "firm-1gap(period)", na.rm=TRUE),
    ebitda_ind_year   = sum(ebitda_src   == "ind×year",          na.rm=TRUE),
    ebitda_ind_period = sum(ebitda_src   == "ind×period",        na.rm=TRUE),
    ebitda_period     = sum(ebitda_src   == "period",            na.rm=TRUE),

    roa_reported    = sum(roa_src == "reported",            na.rm=TRUE),
    roa_recomputed  = sum(roa_src == "recomputed",          na.rm=TRUE),
    roa_firmgap     = sum(roa_src == "firm-1gap(period)",   na.rm=TRUE),
    roa_ind_year    = sum(roa_src == "ind×year",            na.rm=TRUE),
    roa_ind_period  = sum(roa_src == "ind×period",          na.rm=TRUE),
    roa_period      = sum(roa_src == "period",              na.rm=TRUE),

    roe_reported    = sum(roe_src == "reported",            na.rm=TRUE),
    roe_recomputed  = sum(roe_src == "recomputed",          na.rm=TRUE),
    roe_firmgap     = sum(roe_src == "firm-1gap(period)",   na.rm=TRUE),
    roe_ind_year    = sum(roe_src == "ind×year",            na.rm=TRUE),
    roe_ind_period  = sum(roe_src == "ind×period",          na.rm=TRUE),
    roe_period      = sum(roe_src == "period",              na.rm=TRUE)
  )
```

1.  EBIT & EBITDA Imputation

```{r impute-margins}
impute_margins <- impute_summary_q1 |>
  select(country, period, n_rows,
         ebit_reported, ebit_ind_year, ebit_period, ebit_firmgap,
         ebitda_reported, ebitda_ind_year, ebitda_period, ebitda_firmgap)

impute_margins |>
  kable(align = "l", booktabs = TRUE) |>
  kable_styling(full_width = FALSE, font_size = 11)
```

According to the table, Germany shows strong coverage of EBIT and EBITDA margins in both periods, with about 82 – 84% of values reported directly.
The remaining 16 – 18% are filled using industry and year medians, and Germany does not need any firm-gap or period-level imputations.
In comparison, Sweden has lower direct reporting, around 69 – 71% across both periods.
Approximately 29 – 31% of Swedish records depend on industry and year medians, which is a higher dependence than in Germany.
Only four rows in each period use a period-level fallback, indicating that heavy imputations are very limited.
Overall, imputations are present in both countries, but Sweden requires more support due to lower direct reporting.

2.  ROA Imputation

```{r impute-roa}
impute_roa <- impute_summary_q1 |>
  select(country, period, n_rows,
         roa_reported, roa_recomputed, roa_ind_year, roa_ind_period, roa_period)

impute_roa |>
  kable(align = "l", booktabs = TRUE) |>
  kable_styling(full_width = FALSE, font_size = 11)
```

According to the table, ROA is nearly fully available in Germany, with 98 – 99% of rows reported directly in both periods.
Only about 1% of observations depend on recomputation from accounting totals, and industry and year replacement is below 0.3%.
Sweden also shows strong ROA coverage, around 94 – 95% reported or recomputed.
However, Sweden relies more on recomputation, approximately 5% in each period.
Industry-level fallback is very limited.
Overall, data quality remains stable in both countries, though Sweden requires slightly more reconstruction of ROA than Germany.

3.  ROE Imputation

```{r impute-roe}
impute_roe <- impute_summary_q1 |>
  select(country, period, n_rows,
         roe_reported, roe_recomputed, roe_ind_year, roe_ind_period, roe_period)

impute_roe |>
  kable(align = "l", booktabs = TRUE) |>
  kable_styling(full_width = FALSE, font_size = 11)
```

According to the table, Germany had strong ROE availability, with about 96% of records reported directly in both periods.
Around 0.4 – 0.5% were recomputed, and roughly 3 – 4% used industry and year fallbacks.
Sweden also had good coverage, above 94% reported in both periods.
However, Sweden relied more on recomputation and on industry and year fallbacks combined, accounting for around 5 – 6% of observations.
No period-level imputations occurred in either country.
Overall, both datasets provided reliable ROE values, with Sweden needing slightly more support from industry medians.

### Question 2

```{r q2-ida}
# Create leverage and liquidity indicators for both countries
lev_liq_data <- filtered_germany_sweden |>
  filter(acct_year %in% 2018:2021, status == "Active") |>
  mutate(
    debt_to_assets = total_liabilities_eur / total_assets_eur,
    equity_ratio   = total_equity_eur / total_assets_eur
  ) |>
  select(acct_year, country, industry_group,
         debt_to_assets, equity_ratio, gearing_pct,
         solvency_pct, current_ratio) |>
  drop_na(debt_to_assets, current_ratio)

# Summarise by year and country
lev_liq_summary <- lev_liq_data |>
  group_by(country, acct_year) |>
  summarise(
    median_debt_assets = median(debt_to_assets, na.rm = TRUE),
    iqr_debt_assets    = IQR(debt_to_assets, na.rm = TRUE),
    median_liquidity   = median(current_ratio, na.rm = TRUE),
    iqr_liquidity      = IQR(current_ratio, na.rm = TRUE),
    .groups = "drop"
  )

lev_liq_summary |>
  knitr::kable(
    caption = "Median and IQR of Leverage and Liquidity by Country (2018–2021)",
    digits = 2
  ) |>
  kableExtra::kable_styling(full_width = FALSE, font_size = 11)
```

Observation: Both countries maintained median liquidity near 1.8×, while median leverage hovered around 0.50–0.52.
Sweden shows slightly higher liquidity dispersion, suggesting a broader mix of firm sizes and capital structures.

### Question 3

For this question, Initial Data Analysis (IDA) is conducted using only approved ETC5521 methods (see lectures and tutorials, Weeks 3–4), focusing on group-wise statistics by industry and robust distributional checks.

1. Financial Indicators Summaries

```{r industry-financial}

industry_financial_summary <- filtered_germany_sweden |> 
  filter(country == "Germany", !is.na(industry_group), acct_year >= 2019, acct_year <= 2021) |> 
  mutate(period = ifelse(acct_year < 2020, "Pre-pandemic", "Pandemic")) |>
  group_by(industry_group, period) |> 
  summarise(
    n_firms = n(),
    median_profit = median(net_income_eur, na.rm = TRUE),
    median_roa = median(roa_pct, na.rm = TRUE),
    median_liq = median(current_ratio, na.rm = TRUE),
    median_lev = median(gearing_pct, na.rm = TRUE),
    .groups = "drop"     
  ) |> 
  arrange(industry_group, period)

# Display table
knitr::kable(
  industry_financial_summary,
  digits = 2,
  booktabs = TRUE,
  caption = "Industry-level summary as per IDA procedure"
)

```

**Interpretation**

The summary of key financial indicators for German industries between 2019 and 2021 shows clear differences in performance before and during the pandemic. Overall, **profitability** (measured by net income and ROA) declined across most sectors during the pandemic years, while **leverage** tended to rise as firms relied more on debt to manage financial pressures. **Liquidity** patterns were mixed—capital-intensive industries maintained relatively stable current ratios, whereas service-based sectors such as hospitality and transport faced tighter cash flow conditions. These trends suggest that manufacturing and technology industries demonstrated stronger financial resilience compared to consumer-facing sectors that were more directly affected by pandemic disruptions.


2. Outlier Detection — IQR Rule

```{r Outlier detection}
outlier_summary <- function(x) {
  q1 <- quantile(x, 0.25, na.rm=TRUE)
  q3 <- quantile(x, 0.75, na.rm=TRUE)
  iqr <- q3 - q1
  lower <- q1 - 1.5 * iqr
  upper <- q3 + 1.5 * iqr
  sum(x < lower | x > upper, na.rm=TRUE)
}
financial_vars <- c("ebit_margin_pct", "gearing_pct", "current_ratio", "total_assets_eur")
outlier_counts_sweden <- filtered_germany_sweden  |> 
  filter(country == "Sweden", acct_year >= 2019, acct_year <= 2021)  |> 
  summarise(across(all_of(financial_vars), outlier_summary))

knitr::kable(outlier_counts_sweden, caption="Outlier counts for Sweden: Key financial variables (IQR rule)")

```

**Interpretation**

The outlier analysis for Sweden from 2019 to 2021 reveals noticeable variability in firms’ financial performance across key indicators. A relatively high number of outliers in **EBIT margin** and **gearing ratio** suggests that some firms experienced unusually strong or weak profitability and varying levels of financial leverage during the pandemic. This highlights the uneven financial effects across industries, with certain firms taking on more debt or achieving exceptional margins. In contrast, **current ratio** and **total assets** show fewer outliers, indicating that liquidity positions and firm sizes remained more consistent across Swedish companies during this period.

3. Data Completeness (Missingness Bar Plot)

```{r missing-values-sweden}
# Prepare long-format data (already done)
missing_long_compare <- filtered_germany_sweden |>  
  filter(country %in% c("Sweden", "Germany"), acct_year >= 2019, acct_year <= 2021) |>  
  select(country, acct_year, all_of(financial_vars)) |>  
  pivot_longer(-c(country, acct_year), names_to = "variable", values_to = "value") |>  
  mutate(missing = is.na(value)) |>  
  group_by(country, acct_year, variable) |>  
  summarise(prop_missing = mean(missing), .groups = "drop")

# Ribbon plot
ggplot(missing_long_compare, aes(x = acct_year, y = prop_missing, fill = country, group = country)) +
  geom_ribbon(aes(ymin = 0, ymax = prop_missing), alpha = 0.3) +
  geom_line(aes(color = country), size = 1) +
  facet_wrap(~variable, scales = "free_y") +
  labs(
    title = "Missingness Trends by Variable for Sweden and Germany (2019–2021)",
    x = "Year",
    y = "Proportion Missing",
    fill = "Country",
    color = "Country"
  ) +
  theme_bw(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 13),
    axis.text.x = element_text(face = "bold"),
    legend.position = "right"
  )

```

**Interpretation**

The ribbon plot compares the proportion of missing values for key financial variables between Swedish and German firms from 2019 to 2021. Overall, most variables show relatively low levels of missing data in both countries, indicating generally reliable datasets. However, some variables, such as **EBIT margin** and **gearing ratio**, exhibit higher missingness, particularly during the pandemic period. The plot highlights country-specific differences: Sweden tends to have slightly higher missingness in certain measures, whereas Germany shows more consistent completeness across variables. The ribbon effectively visualizes both the magnitude and trend of missingness over time, making it easier to identify variables and years that may require careful interpretation or imputation.



4. Comparative Summary Table: Germany vs Sweden (Pandemic Years)

```{r summary table}
combined_financial_summary <- filtered_germany_sweden  |> 
  filter(country %in% c("Germany", "Sweden"), !is.na(industry_group), acct_year >= 2019, acct_year <= 2021)  |> 
  group_by(country, industry_group)  |> 
  summarise(
    median_roa = median(roa_pct, na.rm = TRUE),
    median_liq = median(current_ratio, na.rm = TRUE),
    median_lev = median(gearing_pct, na.rm = TRUE)
  )  |> 
  pivot_wider(names_from = country, values_from = c(median_roa, median_liq, median_lev))

knitr::kable(combined_financial_summary, 
             digits = 2, 
             booktabs = TRUE, 
             caption = "Comparison of Key Financial Medians by Industry: Germany vs Sweden (2019-2021)")

```

**Interpretation**

The comparative table presents key financial medians—return on assets (ROA), liquidity (current ratio), and leverage (gearing ratio)—for German and Swedish firms across industries during the pandemic years (2019–2021). Overall, the table highlights differences in financial resilience between the two countries. For example, some Swedish industries, particularly technology and manufacturing, exhibit higher median ROA and stronger liquidity compared to their German counterparts, suggesting relatively better profitability and cash management. Conversely, certain German industries, such as transport and hospitality, show higher leverage levels, indicating greater reliance on debt financing during the pandemic. These contrasts underscore how country-specific factors, including market structure, government support, and industry composition, shaped firms’ financial outcomes during the pandemic period.

### Question 4

1.  Data Cleaning on Consolidation Codes

The raw data from Orbis contained financial statements under different consolidation codes (C1, C2, C\*, U1, U2).
To ensure consistency and comparability, we prioritized consolidated accounts which represent the entire economic entity.
For each firm-year observation, we retained the highest-priority available consolidated statement following the order: C1 \> C2 \> C\*.
Unconsolidated statements (U1, U2) were excluded from the analysis.

```{r Q4 IDA 1. Data Cleaning on Consolidation Codes}
ger_swe_q4 <- filtered_germany_sweden |>
  mutate(
    consolidation_priority = case_when(
      consolidation_code == "C1" ~ 1,  # highest priority
      consolidation_code == "C2" ~ 2,
      consolidation_code == "C*" ~ 3,
      consolidation_code == "U1" ~ 99, # Very low priority, about to be filtered
      consolidation_code == "U2" ~ 99,
      TRUE ~ 99  # Handle any unexpected code
    )
  ) |>
  
  # Group by company and year, and keep only the highest priority records
  group_by(company_id, fy_year) |>
  arrange(consolidation_priority, .by_group = TRUE) |>
  slice(1) |> # Take the first row of each group, which has the highest priority
  ungroup() |>
  # Filter out the remaining non-consolidated records
  filter(consolidation_priority %in% c(1, 2, 3))
```

2.  Select the columns to be analyzed

```{r Q4 IDA 2. Select the columns to be analyzed}
ger_swe_q4 <- ger_swe_q4 |>
  select(company_id, fy_year, roa_pct, current_ratio, industry_group, country)
```

The analysis in this study is based on a dataset of company financial information.
It is important to note that in this question we use the fiscal year, not the calendar year.
A company's fiscal year is its own 12-month reporting period, which may not end in December.
Using the fiscal year is more accurate for this research because it matches the company's true business cycle.
This means that when we refer to "2020" in the data, we are talking about the fiscal year that was most affected by the pandemic, even if it does not exactly match the calendar year 2020.

3.  Convert data into a tsibble object using the tsibble package (Lecture 9, slide 5 & Reading week 9)

Each company was defined as a key and the financial year as the index.
Basic screening was performed to check for missing values, outliers, and duplicates.

```{r Q4 IDA 3. Convert data into a tsibble object using the tsibble package (Lecture 9, slide 5 & Reading week 9)}
ger_swe_ts <- ger_swe_q4 |>
  as_tsibble(key = company_id, index = fy_year) |>
  arrange(company_id, fy_year)
```

4.  Missing value - Check for missing value (Tutorial 8)

The initial examination of missing values revealed a relatively low proportion of incomplete data in the dataset.
As shown in @fig-missing_clu, the missing value analysis showed that only 0.9% of the total data was missing, while 99.1% of observations were complete.
ROA contained between 3% and 3.5% missing values across the dataset, while the current ratio had a slightly higher proportion of missing data, ranging from 1% to 2% (@fig-missing_plots).

```{r Q4 IDA 4. Missing value -Check for missing value}
#| label: fig-missing_plots
#| fig-cap: "The plots of Missing Value Distribution"

# Check for missing ROA and CR
miss_var <- 
  gg_miss_var(ger_swe_ts, show_pct = TRUE) + 
  labs(title = "Distribution of Missing Values", 
       subtitle = "For all variables")

# Check for missing values by industry group
miss_industry <- ger_swe_ts |> 
  group_by(industry_group) |> 
  gg_miss_var(show_pct = TRUE) + 
  labs(title = "Distribution of Missing Values", 
       subtitle = "By industry group")

  
miss_var / miss_industry
```

```{r Q4 EDA 4. Missing value -Check for missing value 2}
#| label: fig-missing_clu
#| fig-cap: "The plot of clustering to visualize missing patterns"

m3 <- vis_miss(ger_swe_ts, cluster = TRUE, sort_miss = TRUE) +
  theme(axis.text.x = element_text(size = 8, angle = 45, hjust = 1),
        axis.text.y = element_text(size = 8))

m3
```

**Address missingness issues**

To address missing data, observations with missing values in either ROA or current ratio were removed from the dataset.
Additionally, to ensure temporal consistency across the study period, only industry groups that contained data for all three years (2019-2021) were retained.
The 'Other / Unmapped' category was removed from the study as it comprises companies without clear industry classification.

```{r Q4 IDA 4. Address missingness issues}
ger_swe_ts <- ger_swe_ts |>
  as_tibble() |>   # Convert to a normal tibble
  # Remove observations with missing ROA or CR
  filter(!is.na(roa_pct) & !is.na(current_ratio)) |>
   # Remove industries that don't have data in 2019-2021
  group_by(industry_group) |>
  filter(all(2019:2021 %in% fy_year)) |>
  filter(industry_group != "Other / Unmapped") |>
  ungroup() |>
  # Convert back to tsibble format
  as_tsibble(key = company_id, index = fy_year) |>
  arrange(company_id, fy_year)
```

5.  Summary

The data preparation followed the same cleaning and preprocessing pipeline as described for Germany, ensuring comparability across the two countries.
Consolidated financial statements were prioritized, and only industries with complete data for 2019–2021 were retained.

The comparative analysis of ROA and Current Ratio trends (@tbl-summary) from 2017 to 2021 reveals stark differences between Germany and Sweden, with distinct pandemic-related patterns.
German companies demonstrated remarkable resilience in profitability, with ROA recovering strongly from a low of 0.37% in 2020 to peak at 5.00% in 2021.
It indicates a robust post-pandemic recovery.
In contrast, Swedish firms maintained consistently negative ROA throughout the period, ranging from -9.69% to -11.29%.
It suggests persistent profitability challenges unaffected by the pandemic.
Regarding liquidity, both countries maintained stable Current Ratios.
German companies showed more volatility, dipping to 2.62 in 2019 before recovering to 3.11 in 2020.
However, Swedish firms maintained consistently higher ratios between 2.95-3.48, indicating stronger short-term liquidity positions.

This difference suggests that while the pandemic significantly impacted German profitability with a subsequent strong recovery, Swedish companies faced deeper structural profitability issues.
But both maintained adequate short-term financial stability throughout the period.

```{r Q4 IDA 5. Summary}
#| label: tbl-summary
#| tbl-cap: "Trends in Mean Return on Assets (ROA) and Current Ratio from Pre-Pandemic to Post-Pandemic Years (2017-2021) in Two Countries"

kable(ger_swe_ts |>
  as_tibble() |>
  group_by(fy_year, country) |>
  summarise(mean_roa = mean(roa_pct, na.rm = TRUE),
            mean_cr = mean(current_ratio, na.rm = TRUE)) |>
  ungroup())
```

## EDA

Here are the EDA processes and results for the sub-questions addressing the primary question "How did the pandemic affect corporate financials in Germany?"

### Question 1

**1. How did profitability change across industries?**

#### Profitability Measures Analysis

This code recodes years into Pre (2018 – 2019) versus Pandemic (2020 – 2021), reshapes five profitability measures (EBIT, EBITDA, ROA, ROE, net income) to long format, and plots violin plots with jitter, median, and IQR overlays.
It allows us to compare central tendency and spread before versus during the pandemic across all firms, and net income uses the asinh function, so negatives remain visible.

```{r q1_profitability_distributions_overall}
#recode period to short labels
ida_question1 <- ida_question1 |>
  mutate(
    period = if_else(acct_year >= 2020, "Pandemic (2020–21)", "Pre (2018–19)"),
    period = factor(period, levels = c("Pre (2018–19)", "Pandemic (2020–21)"))
  )

#build long data
q1_vars <- c("ebit_margin_q1","ebitda_margin_q1","roa_q1","roe_q1","net_income_eur_asinh")

plot_long <- ida_question1 |>
  select(company_id, country, period, all_of(q1_vars)) |>
  pivot_longer(-c(company_id, country, period), names_to = "metric", values_to = "value") |>
  mutate(
    metric = factor(
      metric,
      levels = q1_vars,
      labels = c("EBIT margin (%)","EBITDA margin (%)","ROA (%)","ROE (%)","Net income (asinh)")
    )
  ) |>
  drop_na(value)

#helper for IQR bar
iqr_bar <- function(y) data.frame(
  y    = median(y, na.rm=TRUE),
  ymin = quantile(y, 0.25, na.rm=TRUE),
  ymax = quantile(y, 0.75, na.rm=TRUE)
)

#helper to draw a panel
draw_q1_panel_country <- function(data_subset, ncol = 2, title_text = "") {
  ggplot(
    data_subset,
    aes(x = period, y = value, colour = country, group = interaction(country, period))
  ) +
    geom_violin(
      trim = FALSE, fill = NA, linewidth = 0.45, na.rm = TRUE,
      position = position_dodge(width = 0.75)
    ) +
    geom_jitter(
      alpha = 0.10, size = 0.55, na.rm = TRUE,
      position = position_jitterdodge(jitter.width = 0.08, jitter.height = 0, dodge.width = 0.75)
    ) +
    stat_summary(
      fun.data = iqr_bar, geom = "errorbar", width = 0.12, linewidth = 0.6, na.rm = TRUE,
      position = position_dodge(width = 0.75), show.legend = FALSE
    ) +
    stat_summary(
      fun = median, geom = "point", shape = 95, size = 8, na.rm = TRUE,
      position = position_dodge(width = 0.75), show.legend = FALSE
    ) +
    facet_wrap(~ metric, scales = "free_y", ncol = ncol) +
    scale_x_discrete(labels = c(
      `Pre (2018–19)`      = "Pre\n(2018 – 2019)",
      `Pandemic (2020–21)` = "Pandemic\n(2020 – 2021)"
    )) +
    labs(x = NULL, y = NULL, colour = "country", title = title_text) +
    theme_minimal(base_size = 11) +
    theme(
      strip.text  = element_text(face = "bold"),
      axis.text.x = element_text(size = 10, margin = margin(t = 8), lineheight = 0.95),
      plot.margin = margin(10, 15, 24, 10),
      legend.position = "top"
    )
}
```

**EBIT & EBITDA Analysis using Median and IQR**

```{r q1-ebit-ebitda}
figA_metrics <- c("EBIT margin (%)", "EBITDA margin (%)")
p_q1_A <- draw_q1_panel_country(
  dplyr::filter(plot_long, metric %in% figA_metrics),
  ncol = 2,
  title_text = "Pre vs Pandemic: Germany vs Sweden\nEBIT & EBITDA Analysis using Median and IQR"
)

p_q1_A
```

According to the chart, EBIT and EBITDA margins in both Germany and Sweden remain centred near zero across both periods.
The medians stay nearly the same between the pre-pandemic and pandemic years, indicating that the overall level of operating profitability did not change much for most firms.
However, the shape of the violins broadens in 2020 – 2021, especially for Sweden, showing increased variation in margins.
This suggests that while many firms maintained similar performance, a larger number experienced either significant gains or greater losses during the pandemic.
Germany shows a slightly narrower distribution, suggesting more stable profitability than Sweden.
Overall, the chart indicates that average profitability remained steady, but performance across firms became more uneven during the pandemic.

**ROA & ROE Analysis Using Median and IQR**

```{r q1-roa-roe}
figB_metrics <- c("ROA (%)", "ROE (%)")
p_q1_B <- draw_q1_panel_country(
  dplyr::filter(plot_long, metric %in% figB_metrics),
  ncol = 2,
  title_text = "Pre vs Pandemic: Germany vs Sweden\nROA & ROE analysis using Median and IQR"
)

p_q1_B
```

The chart compares ROA and ROE for Germany and Sweden before the pandemic (2018 – 2019) and during the pandemic (2020 – 2021).
For both countries, the medians stay close to zero, indicating that overall profitability on assets and equity did not change significantly between the two periods.
However, the range of values widens during the pandemic, especially in Sweden, where several firms exhibit extremely negative ROE.
This pattern suggests that while most firms maintained stable returns, some experienced substantial losses or reductions in equity in 2020 – 2021.
Germany also shows a wider spread compared to the pre-pandemic period, but the dispersion is less pronounced than in Sweden.
Overall, the chart indicates that profitability stability at the centre conceals greater volatility across firms, with Sweden experiencing stronger downside movements in both ROA and ROE during the pandemic.

**Net income (asinh) Analysis Using Median and IQR**

```{r}
figC_metrics <- c("Net income (asinh)")
p_q1_C <- draw_q1_panel_country(
  dplyr::filter(plot_long, metric %in% figC_metrics),
  ncol = 1,
  title_text = "Pre vs Pandemic: Germany vs Sweden\nNet income (asinh) Analysis Using Median and IQR"
)

p_q1_C
```

The chart compares net income (asinh) between Germany and Sweden before (2018 – 2019) and during the pandemic (2020 – 2021).
In both countries, the median values remain close to zero across the two periods, indicating that overall profitability did not change much.
However, Sweden shows a wider spread of values in both directions, especially during the pandemic, leading more firms to record stronger profits or deeper losses.
Germany’s distribution is narrower and more concentrated around the median, reflecting relatively stable firm performance.
The heavier tails in Sweden during 2020 – 2021 suggest increased volatility, with a few firms benefiting while others faced larger setbacks.
Overall, the figure shows that while average profitability remained steady, Sweden experienced greater variation in firm outcomes during the pandemic than Germany.

#### Industry Median for EBITDA Margin Analysis

This code retains the mapped industries, calculates industry medians for EBITDA margin during the Pre and Pandemic periods, and creates a dumbbell chart showing the two medians and the change (Pandemic – Pre), coloured by increase or decrease.
Using medians ensures the comparison is robust against outliers and highlights which sectors improved or worsened the most.

```{r q1_industry_ebitda_median_delta}
#mapped industries and build period label
eb_src <- ida_question1 |>
  filter(
    !is.na(ebitda_margin_q1),
    !is.na(industry_group),
    industry_group != "Other / Unmapped"
  ) |>
  mutate(
    period = ifelse(acct_year >= 2020, "Pandemic", "Pre"),
    period = factor(period, levels = c("Pre", "Pandemic"))
  )

#median by country × industry × period, plus delta
eb_med_cs <- eb_src |>
  group_by(country, industry_group, period) |>
  summarise(med = median(ebitda_margin_q1, na.rm = TRUE), .groups = "drop") |>
  pivot_wider(names_from = period, values_from = med) |>
  drop_na(Pre, Pandemic) |>
  mutate(
    delta     = Pandemic - Pre,
    direction = factor(ifelse(delta >= 0, "Increase", "Decrease"),
                       levels = c("Increase", "Decrease"))
  )

#one common ordering for industries
order_levels <- eb_med_cs |>
  group_by(industry_group) |>
  summarise(delta_mean = mean(delta, na.rm = TRUE), .groups = "drop") |>
  arrange(delta_mean) |>
  pull(industry_group)

eb_med_cs <- eb_med_cs |>
  mutate(industry_group = factor(industry_group, levels = order_levels))

#facet by country
ggplot(eb_med_cs,
       aes(y = industry_group)) +
  geom_segment(aes(x = Pre, xend = Pandemic,
                   yend = industry_group, colour = direction),
               linewidth = 1) +
  geom_point(aes(x = Pre),      colour = "#4E79A7", size = 2) +  #pre (blue)
  geom_point(aes(x = Pandemic), colour = "#E15759", size = 2) +  #pandemic (red)
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey70") +
  scale_color_manual(values = c("Increase" = "#2CA02C", "Decrease" = "#D62728")) +
  guides(colour = guide_legend(title = NULL, override.aes = list(linewidth = 3))) +
  labs(
    title = "Change in Median EBITDA Margin: Pandemic vs Pre\nGermany and Sweden by Industry",
    x = "Median EBITDA Margin (%)",
    y = NULL
  ) +
  facet_grid(. ~ country) +
  theme_minimal(base_size = 11) +
  theme(
    legend.position      = "top", #legend below title
    legend.box           = "vertical",#stack under the title
    plot.title.position  = "plot",
    strip.text           = element_text(face = "bold"),
    panel.grid.minor     = element_blank(),
    plot.margin          = margin(10, 20, 10, 10)
  )
```

This chart compares the median EBITDA margin for each industry between the pre-pandemic period (2018 – 2019) and the pandemic period (2020 – 2021) in Germany and Sweden.
Each line shows the change in industry profitability, with the blue dot indicating the pre-pandemic median and the red dot indicating the pandemic median.
The line colour reveals whether the median increased or decreased.
In both countries, most industries experienced moderate or small shifts.
In Germany, Accommodation & Food showed the strongest improvement, moving from a negative to a positive margin, while Health Care recorded the largest decline from its previously high level.
Technology and Utilities also performed relatively well, with mild increases.
In Sweden, the direction of change is similar across many sectors, but the shifts are generally smaller.
Accommodation & Food improved slightly, while Health Care, Consumer Services, and Construction saw reductions in profitability.
Overall, both economies demonstrate that the pandemic affected industries unevenly: service-based and cyclical sectors were more volatile, while Technology, Utilities, and Education remained relatively stable.
Using medians offers a clearer view of typical firm performance without distortion from extreme values.

### Question 2

#### EDA – Year-wise Comparison

```{r}
lev_liq_long <- lev_liq_summary |>
  pivot_longer(cols = c(median_debt_assets, median_liquidity),
               names_to = "indicator", values_to = "value")

ggplot(lev_liq_long,
       aes(x = acct_year, y = value, colour = country, group = country)) +
  geom_line(linewidth = 1.1) +
  geom_point(size = 2) +
  facet_wrap(~ indicator, scales = "free_y",
             labeller = as_labeller(c(
               median_debt_assets = "Median Debt-to-Assets",
               median_liquidity   = "Median Current Ratio"
             ))) +
  scale_y_continuous(labels = scales::number_format(accuracy = 0.01)) +
  scale_colour_brewer(palette = "Set1") +
  labs(
    title = "Leverage and Liquidity Trends (2018–2021)",
    subtitle = "Germany vs Sweden: Stability in liquidity, mild re-leveraging during pandemic",
    x = "Accounting Year", y = "Median Value", colour = "Country"
  ) +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(size = 11),
        panel.grid.minor = element_blank())
```

Interpretation: • Both economies experienced a slight leverage increase in 2020 (Germany +0.01, Sweden +0.02).
• Liquidity remained broadly constant across the pandemic, stabilising near 1.8–1.9×.
• The trends suggest that German firms relied more on credit support, while Swedish firms drew on retained liquidity, consistent with each country’s policy emphasis (Germany’s Kurzarbeit subsidies vs Sweden’s fiscal buffers).

#### EDA – Leverage–Solvency Relationship

```{r}
#| label: fig-leverage-solvency-facets
#| fig-cap: "Leverage–Solvency Relationship by Country and Year"
#| fig-width: 14
#| fig-height: 9

# build plotting data for the leverage–solvency faceted chart
lev_liq_plot <- lev_liq_data |>
  # keep sensible ranges and finite values
  filter(
    is.finite(debt_to_assets), is.finite(solvency_pct),
    between(debt_to_assets, 0, 1.5),   # 0–150%
    between(solvency_pct, -50, 100)    # -50% to 100%
  ) |>
  drop_na(country, acct_year)

ggplot(lev_liq_plot, aes(x = debt_to_assets, y = solvency_pct, colour = country)) +
  geom_point(alpha = 0.25, size = 1.2) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 1.1) +
  facet_grid(acct_year ~ country) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, 1.5)) +
  scale_y_continuous(limits = c(-50, 100)) +
  scale_colour_manual(values = c("Germany" = "#D55", "Sweden" = "#0072B2")) +
  labs(
    title = "Leverage–Solvency Relationship by Country and Year",
    subtitle = "Stronger negative slope in Germany, especially during 2020",
    x = "Debt-to-Assets (%)",
    y = "Solvency (%)",
    colour = "Country"
  ) +
  theme_minimal(base_size = 15) +
  theme(
    strip.text = element_text(face = "bold", size = 14),
    plot.title = element_text(face = "bold", size = 18, hjust = 0.5),
    plot.subtitle = element_text(size = 13, hjust = 0.5),
    legend.position = "none",  # each facet is already labelled
    axis.text = element_text(size = 12),
    panel.spacing.x = unit(1.2, "lines"),
    panel.spacing.y = unit(1.0, "lines"),
    panel.grid.minor = element_blank()
  )
```

Interpretation:

Across both Germany and Sweden, leverage shows a **strong negative relationship with solvency** from 2018 to 2021 — firms with higher debt-to-asset ratios consistently report lower solvency levels.
The **pattern remains stable** over time, confirming that higher leverage structurally reduces long-term financial stability in both economies.

However, the **slope is notably steeper in Germany**, especially in **2020**, indicating that increases in debt had a stronger adverse effect on solvency during the pandemic.
This likely reflects the **manufacturing-heavy structure** of German industries, which faced sharper revenue disruptions and slower balance sheet recovery.
In contrast, **Swedish firms**, with **more diversified sectors and stronger equity positions**, experienced a milder solvency decline, demonstrating greater resilience to leverage-induced stress.

EDA – Industry-level Leverage Distributions

```{r}
#| label: fig-leverage-industry-comparison
#| fig-cap: "Leverage Distribution by Industry and Year for Germany and Sweden (2018–2021)"
#| fig-width: 16
#| fig-height: 20

lev_liq_industry <- filtered_germany_sweden |>
  filter(acct_year %in% 2018:2021,
         total_assets_eur > 0,
         total_liabilities_eur >= 0,
         total_liabilities_eur / total_assets_eur <= 1.5) |>
  mutate(
    debt_to_assets = total_liabilities_eur / total_assets_eur
  ) |>
  drop_na(industry_group, country)

ggplot(lev_liq_industry,
       aes(x = debt_to_assets,
           y = reorder(industry_group, debt_to_assets),
           fill = country)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7, width = 0.6) +
  facet_wrap(~ acct_year, ncol = 2, scales = "free_x") +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1),
                     limits = c(0, 1.5)) +
  scale_fill_manual(
    values = c("Germany" = "#D55", "Sweden" = "#0072B2"),
    name = "Country"
  ) +
  labs(
    title = "Leverage Distribution by Industry and Year (2018–2021)",
    subtitle = "Comparison between Germany and Sweden across industry groups and years",
    x = "Debt-to-Assets Ratio",
    y = NULL
  ) +
  theme_minimal(base_size = 16) +  # increased from 13 → 16
  theme(
    strip.text = element_text(face = "bold", size = 15),
    plot.title = element_text(face = "bold", size = 20, hjust = 0.5),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    legend.position = "top",
    legend.text = element_text(size = 13),
    legend.title = element_text(size = 14),
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    panel.grid.minor = element_blank(),
    panel.spacing.y = unit(1.2, "lines")
  )
```

Interpretation:

The boxplots reveal that leverage remained broadly stable across both countries from 2018 to 2021, with **capital-intensive sectors**-notably *Manufacturing*, *Construction*, and *Transport & Storage*-maintaining higher debt-to-asset ratios than service-oriented industries such as *Information & Communication* or *Professional & Technical Services*.

The median levels changed little across years, though a modest upward shift appears in 2020, consistent with increased borrowing during the early pandemic period.

Overall, **Germany and Sweden display similar leverage structures**, suggesting comparable financial resilience and corporate funding patterns.

When leverage patterns are compared with liquidity (current ratio) trends, a complementary story emerges.

Firms in both countries **maintained adequate short-term liquidity**, with median current ratios remaining around 1.8–2.0 throughout 2018–2021.

This indicates that, despite slightly greater reliance on debt during 2020, most firms preserved the capacity to meet near-term obligations.

The combination of stable leverage and steady liquidity implies that corporate balance sheets were **buffered by prudent debt management and strong internal financing**, helping both economies avoid a widespread solvency crisis during the COVID-19 shock.

```{r}
#| warning: false
# Sensitivity check: winsorise extreme ratios
lev_liq_summary_w <- lev_liq_data |>
  mutate(across(c(debt_to_assets, current_ratio),
                ~ scales::squish(., c(quantile(., 0.01, na.rm=TRUE),
                                      quantile(., 0.99, na.rm=TRUE))))) |>
  group_by(country, acct_year) |>
  summarise(across(c(debt_to_assets, current_ratio), median, na.rm=TRUE), .groups="drop")
knitr::kable(head(lev_liq_summary_w), caption="Winsorised medians confirm robustness.")
```

Summary interpretation: • Leverage: Both countries show small increases during 2020, reverting by 2021.
• Liquidity: Stable across years and nations, implying effective cash-flow management.
• Cross-country pattern: Sweden’s firms appear slightly more liquid and less leveraged, signalling greater resilience to credit-market stress.
• Conclusion: Despite pandemic disruptions, neither economy faced systemic liquidity deterioration; Germany’s higher industrial exposure drove modest leverage stress, while Sweden’s diversified base cushioned impact.

### Question 3

The purpose of this section is to examine the structure and main financial indicators of German and Swedish firms before and during the COVID-19 pandemic (2018–2021).
We explore firm size, profitability, leverage, and liquidity across industries and over time.


1. Industry-level Financial Summary for Sweden

```{r Industry summary}
industry_financial_summary_sweden <- filtered_germany_sweden |> 
  filter(country == "Sweden", !is.na(industry_group), acct_year >= 2019, acct_year <= 2021) |> 
  group_by(industry_group, acct_year) |> 
  summarise(
    n_firms = n(),
    median_profit = median(net_income_eur, na.rm = TRUE),
    median_roa = median(roa_pct, na.rm = TRUE),
    median_liq = median(current_ratio, na.rm = TRUE),
    median_lev = median(gearing_pct, na.rm = TRUE),
    .groups = "drop"
  ) |> 
  mutate(period = ifelse(acct_year < 2020, "Pre-pandemic", "Pandemic"))

```


```{r Industry level summary}

combined_summary <- bind_rows(
industry_financial_summary |> mutate(country="Germany"),
industry_financial_summary_sweden |> mutate(country="Sweden")
)


ggplot(combined_summary, aes(
    x = industry_group,
    y = median_profit,
    fill = country
  )) +
  geom_col(position = "dodge", width = 0.7) +
  facet_wrap(~period) +
  coord_flip() +
  labs(
    title = "Median Profit Comparison by Industry",
    x = "Industry",
    y = "Median Net Income (EUR)",
    fill = "Country"
  ) +
  scale_fill_manual(values = c("Germany" = "#E69F00", "Sweden" = "#56B4E9")) +
  scale_y_continuous(labels = scales::label_comma()) +
  theme_minimal(base_size = 10) +
  theme(
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10),
    axis.title.x = element_text(size = 10, face = "bold"),
    axis.title.y = element_text(size = 10, face = "bold"),
    legend.position = "top",
    strip.text = element_text(face = "bold", size = 10)
  )

```


2. Trend of Median Financial Ratios Over Time

```{r median financial ratio}
median_trends <- filtered_germany_sweden |> 
  filter(acct_year >= 2019, acct_year <= 2021) |> 
  group_by(country, acct_year) |> 
  summarise(across(all_of(financial_vars), median, na.rm=TRUE), .groups="drop") |> 
  pivot_longer(cols = all_of(financial_vars), names_to="variable", values_to="median_value")

ggplot(median_trends, aes(x=acct_year, y=median_value, color=country, group=country)) +
  geom_line(size=1) +
  geom_point(size=2) +
  facet_wrap(~ variable, scales="free_y") +
  labs(title="Median Financial Indicators Over Time", x="Year", y="Median Value", color="Country") +
  theme_minimal()

```

3. Boxplots of Key Financial Ratios

```{r key financial ratios}
# Reshape data for plotting
eda_data <- filtered_germany_sweden |> 
  filter(acct_year >= 2019, acct_year <= 2021) |> 
  select(country, acct_year, all_of(financial_vars)) |> 
  pivot_longer(cols = all_of(financial_vars), names_to = "variable", values_to = "value")

ggplot(eda_data, aes(x=country, y=value, fill=country)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 16) +
  facet_wrap(~ variable, scales = "free_y") +
  labs(title="Distribution of Key Financial Ratios (2019-2021)", 
       y="Value",
       x="Country") +
  theme_minimal()
```



### Question 4

**4. How did the fundamental relationship between corporate liquidity and profitability evolve and fracture within industries during the pandemic?**

#### 1. Macro-level evidence: Industry recovery patterns (ROA vs CR) - feasts (Lecture 9, slide 24-26, feasts and tsibbletalk)

We will analyze the fundamental relationship between corporate liquidity and profitability across industries during the pandemic period in two countries, with particular focus on identifying anomalous recovery patterns.

```{r Q4 EDA 1. Macro-level evidence: Calculate}
# Calculate the average ROA and current ratio for each industry each year
industry_ts <- ger_swe_ts |>
  group_by(industry_group, country) |>
  index_by(fy_year) |>
  summarise(
    industry_roa = mean(roa_pct, na.rm = TRUE),
    industry_cr = mean(current_ratio, na.rm = TRUE),
    n_companies = n()  
  )

# Calculating time series features
industry_features <- industry_ts |>
  group_by(industry_group, country) |>
  features(industry_roa, feat_stl) |>
  rename_with(~ paste0("roa_", .), -c(industry_group, country))  |>
  left_join(
    industry_ts |>
      features(industry_cr, features = feat_stl) |>
      rename_with(~ paste0("cr_", .), -c(industry_group, country)),
    by = c("industry_group", "country")
  )

# Focus on trend_strength and spikiness
focus_features <- industry_features |>
  select(industry_group, country,
         roa_trend_strength, roa_spikiness,
         cr_trend_strength, cr_spikiness)
```

**Trend strength analysis**

The trend strength visualization in @fig-combined_analysis reveals significant disruptions to industry financial fundamentals.
Five industries in the bottom-left quadrant, indicating severe deterioration in both ROA and Current Ratio trends.
Accommodation & Food has the largest bubble, reflecting extreme volatility and weak performance during pandemic restrictions.
This pattern reflects the extreme operational challenges faced by hospitality-related sectors during pandemic restrictions.

**Volatility analysis**

After trend strength analysis, the volatility examination in @fig-combined_analysis demonstrates divergent patterns in financial metric instability.
Five industries show high ROA volatility, but most have moderate Current Ratio spikiness.
However, Consumer Services is exceptional with the highest Current Ratio spikiness, suggesting unique liquidity challenges.

```{r Q4 EDA 1. Macro-level evidence -Combined analysis}
#| label: fig-combined_analysis
#| fig-cap: "Combined analysis of trend strength disruption and volatility across industries and countries during the pandemic period. Left: Trend strength analysis with 0.3 threshold lines. Right: Volatility analysis with median lines."

# Trend strength analysis to identify the hardest-hit sectors
trend_analysis <- focus_features |>
  group_by(country) |>
  mutate(
    # Identify sectors where trends are broken (low trend strength)
    roa_trend_disrupted = roa_trend_strength < 0.3,
    cr_trend_disrupted = cr_trend_strength < 0.3,
    overall_trend_impact = (roa_trend_strength + cr_trend_strength) / 2
  ) |>
  ungroup()

# Volatility analysis to identify the sectors with the greatest volatility during the pandemic
volatility_analysis <- focus_features |>
  group_by(country) |>
  mutate(
    high_roa_volatility = roa_spikiness > median(roa_spikiness),
    high_cr_volatility = cr_spikiness > median(cr_spikiness),
    overall_volatility = (scale(roa_spikiness) + scale(cr_spikiness)) / 2)  |>
  ungroup()

# Create trend strength plot
trend_plot <- focus_features |>
  ggplot(aes(x = roa_trend_strength, y = cr_trend_strength)) +
  geom_point(aes(size = roa_spikiness, 
                 color = industry_group, 
                 shape = country,  
                 text = paste("<b>Industry:</b>", industry_group, 
                             "<br><b>Country:</b>", country,  
                             "<br><b>ROA Trend:</b>", round(roa_trend_strength, 3),
                             "<br><b>CR Trend:</b>", round(cr_trend_strength, 3))), 
             alpha = 0.7) +
  geom_hline(yintercept = 0.3, linetype = "dashed", color = "red", alpha = 0.7) +
  geom_vline(xintercept = 0.3, linetype = "dashed", color = "red", alpha = 0.7) +
  labs(x = "ROA Trend Strength", y = "Current Ratio Trend Strength",
       size = "ROA Spikiness", shape = "Country") +
  theme_minimal() + 
  theme(legend.position = "none",
         axis.text = element_text(size = 9)) +
  scale_shape_manual(values = c(16, 17)) 

# Convert to interactive plotly
trend_in <- ggplotly(trend_plot, tooltip = "text") 

# Create volatility plot
volatility_plot <- focus_features |>
  ggplot(aes(x = roa_spikiness, y = cr_spikiness)) +
  geom_point(aes(color = industry_group, 
                 size = roa_trend_strength, 
                 shape = country,  
                 text = paste("<b>Industry:</b>", industry_group,
                             "<br><b>Country:</b>", country,  
                             "<br><b>ROA Spikiness:</b>", round(roa_spikiness, 3),
                             "<br><b>CR Spikiness:</b>", round(cr_spikiness, 3))), 
             alpha = 0.7) +
  geom_hline(yintercept = median(focus_features$cr_spikiness), 
             linetype = "dashed", color = "blue", alpha = 0.5) +
  geom_vline(xintercept = median(focus_features$roa_spikiness), 
             linetype = "dashed", color = "blue", alpha = 0.5) +
  coord_cartesian(xlim = c(0, quantile(focus_features$roa_spikiness, 0.90))) +
  labs(x = "ROA Spikiness", y = "Current Ratio Spikiness",
       size = "ROA Trend Strength", shape = "Country") +
  theme_minimal() +
  theme(legend.position = "none",
         axis.text = element_text(size = 9)) +
  scale_shape_manual(values = c(16, 17))  

volatility_in <- ggplotly(volatility_plot, tooltip = "text")

# Combined analysis
combined_analysis <- subplot(
  trend_in, volatility_in,
  nrows = 1,
  shareY = FALSE,
  titleX = TRUE,
  titleY = TRUE,
  margin = 0.05
) |>
  plotly::layout(
    title = list(
      text = "<b>Industry Financial Impact Analysis During Pandemic</b>",
      x = 0.5,
      xanchor = "center",
      y = 1.05,
      font = list(size = 16, color = "black"),
      automargin = TRUE,
      margin = list(l = 60, r = 60, b = 100, t = 150)
    ),
    hoverlabel = list(
      bgcolor = "white", 
      font = list(size = 11, color = "black"),
      bordercolor = "lightgray"
    ),
    margin = list(l = 60, r = 60, b = 80),
    hovermode = "closest",
    annotations = list(
      list(
      x = 0.5, y = 1.05,
      xref = "paper", yref = "paper",
      text = "Left: Trend strength disruption | Right: Volatility patterns | Shapes indicate countries",
      showarrow = FALSE,
      font = list(size = 10, color = "gray40")
    ),
      list(
        x = 0, y = -0.15,
        xref = "paper", yref = "paper",
        text = "Red lines: 0.3 threshold",
        showarrow = FALSE,
        font = list(size = 10, color = "red")
      ),
      list(
        x = 0.85, y = -0.15,
        xref = "paper", yref = "paper",
        text = "Blue lines: Median volatility values",
        showarrow = FALSE,
        font = list(size = 10, color = "blue")
      )
    )
  )

combined_analysis
```

**Threshold-based industry classification**

Given the comparative focus of our study between Germany and Sweden, we used country-specific threshold-based segmentation to classify industries into recovery archetypes.
This approach categorizes industries based on their ROA trend strength and volatility using the following criteria:

-   High trend strength (≥0.7) + Low spikiness (≤country-specific median) → "Resilient industries"
-   Low trend strength (≤0.3) + High spikiness (≥country-specific 75th percentile) → "K-shaped recovery industries"
-   High trend strength (≥0.7) + High spikiness (≥country-specific 75th percentile) → "Hidden crisis industries"
-   All other combinations → "Moderate impact industries"

The comparative analysis reveals significant differences in sectoral recovery patterns between Germany and Sweden during the post-pandemic period.
While some industries demonstrate consistent performance across both countries, others exhibit notable variations that highlight distinct economic environments.

As shown in @tbl-recovery-classification, a key finding is the divergent performance in several important sectors.
Germany shows stronger resilience in Construction and Financial Services, where these sectors maintained stable performance throughout the recovery period.
Conversely, Sweden demonstrates more robust recovery in Manufacturing and Administrative Support sectors.
This contrast suggests differing national economic priorities and support mechanisms.

The analysis also identifies interesting similarities.
The Information & Communication sector displays a consistent K-shaped recovery pattern in both countries, indicating that technological industries experienced similar uneven recovery trajectories regardless of national context.
This pattern suggests that sector-specific characteristics may sometimes outweigh national economic policies in determining recovery outcomes.

These comparative insights provide valuable understanding of how similar industries can experience different recovery paths depending on their national economic environment, while also highlighting sectors where recovery patterns transcend national boundaries.

```{r Q4 EDA 1. Macro-level evidence -industry classes}
#| label: tbl-recovery-classification
#| tbl-cap: Industry Recovery Classification for Sweden and Germany based on Country-Specific Thresholds

industry_classes <- focus_features |>
  group_by(country) |>
  mutate(
    # Calculate threshold
    trend_threshold_high = 0.7,
    trend_threshold_low = 0.3,
    spikiness_threshold_high = quantile(roa_spikiness, 0.75),
    spikiness_median = median(roa_spikiness), 
    
    # Classification
    recovery_pattern = case_when(
      roa_trend_strength >= trend_threshold_high & 
        roa_spikiness <= median(roa_spikiness) ~ "Resilient",
      roa_trend_strength <= trend_threshold_low & 
        roa_spikiness >= spikiness_threshold_high ~ "K-shaped",
      roa_trend_strength >= trend_threshold_high & 
        roa_spikiness >= spikiness_threshold_high ~ "Hidden Crisis",
      TRUE ~ "Moderate Impact"
    )
  )

# Comparing the classification differences between the two countries
classification_comparison <- industry_classes |>
  select(industry_group, country, recovery_pattern) |>
  pivot_wider(names_from = country, values_from = recovery_pattern) |>
  mutate(
    pattern_match = Germany == Sweden,
    cross_country_insight = case_when(
      Germany == "Resilient" & Sweden != "Resilient" ~ "Resilient only in Germany",
      Sweden == "Resilient" & Germany != "Resilient" ~ "Resilient only in Sweden",
      Germany == "K-shaped" & Sweden != "K-shaped" ~ "K-shaped only in Germany", 
      Sweden == "K-shaped" & Germany != "K-shaped" ~ "K-shaped only in Sweden",
      TRUE ~ "Similar pattern"
    )
  ) |>
  filter(!(Germany == "Moderate Impact" & Sweden == "Moderate Impact")) |>
  select(industry_group, Germany, Sweden, cross_country_insight)

kable(classification_comparison) |>
  kable_styling(font_size = 11)
```

#### 2. Micro-level evidence of anomalous recovery patterns across industries (ROA and liquidity) - scagnostics (Tignostics) (Lecture 8, 22-27 & Lecture 9, slide 24)

In this section, we compare firm-level relationships between liquidity (Current Ratio) and profitability (ROA) across industries in Germany and Sweden.

We focus on the industries that exhibited distinct recovery patterns in the macro analysis (see @tbl-recovery-classification).
Specifically, we select:

-   Information & Communication (K-shaped in both countries)
-   Wholesale & Retail Trade (K-shaped only in Germany)
-   Consumer Services (K-shaped only in Sweden)
-   Financial & Insurance (Resilient only in Germany)
-   Manufacturing (Resilient only in Sweden)

**Data and methods**

For each country × industry × fiscal year, we:

-   keep firm-level pairs (current_ratio, roa_pct);
-   compute scagnostic metrics (clumpy, outlying, monotonic, stringy);
-   fit a simple linear model ROA \~ CR to obtain slope and residual SD;
-   compare scagnostic between countries to represent variability among firms.
-   compute year-to-year slope changes and count “jumps” (a jump = \|slope change\| \> 0.5);
-   compute pre/post differences in mean residual SD and mean slope (post − pre).
-   When an industry × year has too few firms (\< 3) the scagnostics or regression are not reported (NA).

**How to read the scagnostic metrics**

-   Clumpy: higher values indicate the points form separate clusters → suggests within-industry divergence (possible K-shaped pattern).
-   Outlying: higher values indicate presence of extreme firms (very strong survivors or very weak firms).
-   Stringy / Monotonic: higher values indicate a clearer linear or monotonic relationship between CR and ROA.
-   Residual SD: higher residual SD from the linear fit signals more unexplained scatter (higher micro volatility).

```{r Q4 EDA 2. Micro-level evidence -data}
selected_industries <- c("Information & Communication", "Wholesale & Retail Trade", "Consumer Services", "Financial & Insurance", "Manufacturing")

# Extract the companies time series data of these industries
selected_companies_ts <- ger_swe_ts |>
  filter(industry_group %in% selected_industries)
```

```{r Q4 EDA 2. Micro-level evidence -Calculate scagnostic}
#| label: tbl-company
#| tbl-cap: "Summary statistics for industries with significant slope changes (jumps > 0). The table shows average slope, maximum absolute slope change, mean residual standard deviation, number of jumps, and differences in volatility and slope before and after the COVID-19 pandemic for three industries: Wholesale & Retail Trade, Information & Communication, and Financial & Insurance."

# Aggregate the original point set by industry × year
by_ind_year <- selected_companies_ts |>
  group_by(industry_group) |>
  index_by(fy_year) |>
  nest(data = c(company_id, roa_pct, current_ratio)) |>
  mutate(n_obs = map_int(data, ~ nrow(.x)))

# Calculate scagnostics
scag_names <- c("outlying", "stringy", "striated", "clumpy", "sparse", "monotonic", "dcor")

by_ind_year <- by_ind_year |>
  mutate(
    scags = map(data, ~ {
      df <- .x
      # When there are too few samples, return a row of tibble with the same column names and types.
      if (nrow(df) < 3) {
        return(as_tibble(set_names(as.list(rep(NA_real_, length(scag_names))), scag_names)))
      }
      res <- calc_scags(df$current_ratio, df$roa_pct, scags = scag_names)
      # Make sure to convert to atomic vectors and name them, then convert to tibble
      res_vec <- unlist(res)
      res_named <- set_names(as.numeric(res_vec[scag_names]), scag_names)
      as_tibble(as.list(res_named))
    })
  ) |>
  unnest(cols = scags)  # Expand into multiple columns

# Fit a linear regression for each industry × year and extract slope, intercept, and resid_sd
by_ind_year <- by_ind_year |>
  mutate(
    fit = map(data, ~ {
      df <- .x
      if (nrow(df) < 3 || all(is.na(df$current_ratio)) || all(is.na(df$roa_pct))) return(NULL)
      lm(roa_pct ~ current_ratio, data = df)
    }),
    slope = map_dbl(fit, ~ if (is.null(.x)) NA_real_ else coef(.x)[["current_ratio"]]),
    intercept = map_dbl(fit, ~ if (is.null(.x)) NA_real_ else coef(.x)[["(Intercept)"]]),
    resid_sd = map_dbl(fit, ~ if (is.null(.x)) NA_real_ else sd(residuals(.x), na.rm = TRUE))
  ) |>
  select(-fit)

# Industry time-series features, grouped by industry_group AND country
industry_tignostics <- by_ind_year |>
  arrange(industry_group, country, fy_year) |>
  group_by(industry_group, country) |>
  mutate(
    slope_lag = dplyr::lag(slope),
    slope_change = slope - slope_lag,   # year-to-year change in slope
    slope_pct_change = if_else(!is.na(slope_lag) & slope_lag != 0,
                               (slope - slope_lag) / abs(slope_lag) * 100,
                               NA_real_),
    resid_sd_lag = dplyr::lag(resid_sd),
    resid_sd_change = resid_sd - resid_sd_lag,
    big_slope_jump = if_else(!is.na(slope_change) & abs(slope_change) > 0.5, 1L, 0L),
    jumps_cumulative = cumsum(replace_na(big_slope_jump, 0L))
  ) |>
  ungroup()

# Calculate the overall volatility_shift (post - pre) before and after the epidemic (taking 2020 as the epidemic year)
vol_shift <- by_ind_year |>
  mutate(period = case_when(
    fy_year < 2020 ~ "pre",
    fy_year == 2020 ~ "covid_year",
    fy_year > 2020 ~ "post"
  )) |>
  group_by(industry_group, country, period) |>
  summarise(mean_resid_sd = mean(resid_sd, na.rm = TRUE),
            mean_slope = mean(slope, na.rm = TRUE),
            .groups = "drop") |>
  pivot_wider(names_from = period, values_from = c(mean_resid_sd, mean_slope)) |>
  mutate(
    vol_pre_post_diff = mean_resid_sd_post - mean_resid_sd_pre,
    slope_pre_post_diff = mean_slope_post - mean_slope_pre
  )

# Merge the results to generate the final summary
industry_summary <- industry_tignostics |>
  group_by(industry_group, country) |>
  summarise(
    n_years = n_distinct(fy_year),
    avg_slope = mean(slope, na.rm = TRUE),
    max_abs_slope_change = if (all(is.na(slope_change))) NA_real_ else max(abs(slope_change), na.rm = TRUE),
    mean_resid_sd = mean(resid_sd, na.rm = TRUE),
    jumps = if (all(is.na(jumps_cumulative))) 0 else max(jumps_cumulative, na.rm = TRUE),
    .groups = "drop"
  ) |>
  left_join(vol_shift |> select(industry_group, country, vol_pre_post_diff, slope_pre_post_diff), 
            by = c("industry_group", "country"))

kable(industry_summary |> 
        filter( !is.na(max_abs_slope_change)) |>
        select(-n_years)) |>
  kable_styling(font_size = 12)
```

```{r Q4 EDA 2. Micro-level evidence -scagnostic figure}
#| label: fig-scagnostic
#| fig-cap: "Scagnostic metrics across industries and countries. Dashed line marks 2020 (COVID onset). We can see variations in outlying (extreme values), clumpy (clustering patterns), stringy (linear relationships), and monotonic (consistent trends) metrics before and after the pandemic."
#| warning: false

scag_long <- by_ind_year |>
  select(industry_group, country, fy_year, clumpy, outlying, stringy, monotonic) |>
  pivot_longer(cols = c(clumpy, outlying, stringy, monotonic),
               names_to = "scag_metric", values_to = "value")

ggplot(scag_long, aes(x = fy_year, y = value, color = industry_group)) +
  geom_line(linewidth = 0.7) +
  facet_grid(country~ scag_metric, scales = "free_y") +
  geom_vline(xintercept = 2020, linetype = "dashed", color = "red") +
  labs(
    title = "Evolution of Scagnostic Metrics over Time",
    subtitle = "Dashed line = COVID year (2020)",
    x = "Fiscal Year", y = "Scagnostic Value",
    color = "Industry group"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.title = element_text(size = 9),
    legend.text = element_text(size = 6),
    legend.key.width = unit(0.8, "cm"),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 9),
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 11)) +
  guides(color = guide_legend(
    nrow = 2, byrow = TRUE, override.aes = list(size = 3))
    )
```

**Main results**

@tbl-company summarises the scagnostic-based firm-level patterns for three major industries across Germany and Sweden.
Except for Consumer Services, which has too few firms to generate valid scagnostic values, all other industries experienced at least three major slope jump.
This indicates significant shifts in the relationship between ROA and CR during the pandemic years.

@fig-scagnostic visualises the evolution of four scagnostic metrics (outlying, clumpy, stringy, and monotonic) over time for both countries.
The red dashed line marks 2020, when COVID-19 began to affect firm performance and liquidity.

Overall, Germany shows stronger stability with lower variability after COVID-19, while Sweden displays larger fluctuations and volatility increases, especially in the Information & Communication and Financial & Insurance sectors.

**Key patterns observed by industry (Germany vs. Sweden):**

**Wholesale & Retail Trade**

-   Germany: avg_slope ≈ 2.08, max_abs_slope_change ≈ 2.97, mean_resid_sd ≈ 7.79.
    Scagnostics: stable *stringy* and ↓ *clumpy*

    Interpretation: Germany shows moderate slope changes but fairly stable structure.
    Many firms kept similar ROA–CR patterns, and the sector did not become strongly outlying.

-   Sweden: avg_slope ≈ -1.23, max_abs_slope_change ≈ 4.00, mean_resid_sd ≈ 15.61.
    Scagnostics: ↓ *clumpy* and ↓ *stringy* after 2020; ↑ outlying in 2019–2021 up to about 0.8.

    Interpretation: Sweden experienced larger volatility and dispersion.
    This is consistent with a K-shaped response where some firms did much better while others worsened.

**Information & Communication**

-   Germany: avg_slope ≈ 2.50, mean_resid_sd ≈ 8.25.
    Scagnostics: sharply ↓ clumpy after 2020; ↑ monotonic.

    Interpretation: German firms became more predictable in their ROA–CR relation after 2020 (higher monotonic, lower clumpy), suggesting more uniform responses across firms.

-   Sweden: avg_slope ≈ 3.64, mean_resid_sd ≈ 24.62 (much larger).
    Scagnostics: ↑ monotonic but volatility and residual dispersion are much higher; clumpy and outlying patterns differ from Germany.

    Interpretation: Although average slope rose in Sweden, the much higher residual SD indicates wide firm-level dispersion.
    It is a “hidden crisis” where averages hide deep heterogeneity.

**Financial & Insurance**

-   Germany: avg_slope ≈ 0.83, mean_resid_sd ≈ 11.87.
    Scagnostics: ↑ clumpy after 2020, slightly ↓ stringy, small ↓ outlying.

    Interpretation: German financial firms show moderate changes.
    There are some clustering but not huge instability.

-   Sweden: avg_slope ≈ -0.67, mean_resid_sd ≈ 21.47 (higher).
    Scagnostics: rising trends in scagnostics for some metrics, higher volatility overall.

    Interpretation: Swedish financial firms experienced larger instability and dispersion, consistent with stronger liquidity or profitability shocks.

**Manufacturing**

-   Germany: avg_slope ≈ -0.002, mean_resid_sd ≈ 14.32, max_abs_slope_change ≈ 0.54.
    Scagnostics: High but slightly ↓ clumpy, ↓ monotonic after 2020, ↑ outlying in 2020–2021, stable stringy

    Interpretation: Although slope changes were moderate, the rise in outlying and decline in monotonic suggest that the relationship between ROA and CR became less predictable and more fragmented among firms after the pandemic began.

-   Sweden: avg_slope ≈ -0.60, mean_resid_sd ≈ 28.14, max_abs_slope_change ≈ 0.43.
    Scagnostics: High but ↓ clumpy after 2020, ↑ monotonic, mild fluctuations in outlying, stable stringy

    Interpretation: Despite a negative average slope and high residual standard deviation, the increase in monotonic and relatively stable scagnostic values point to a moderately resilient structure.
    Firms became more aligned in their ROA–CR behaviour over time, though some outliers remained.

**Summary**

In short, Germany displays more stable post-2020 behaviour across these industries (smaller volatility shifts and steadier scagnostic patterns).
Sweden shows larger volatility and stronger divergence across firms after 2020, especially in Information & Communication and Financial & Insurance.
This suggests national differences in how firms adjusted liquidity and profitability during the pandemic.
For example, institutional settings, firm composition, or policy responses may have led Swedish firms to react more heterogeneously.

**Limitations**

Results are descriptive: scagnostics indicate shape changes but do not prove causality.
We will triangulate these micro results with macro indicators (feasts features) in the final section.

#### 3. Comparative Analysis: Macro-Level Dynamics vs. Micro-Level Firm Relationships

This section combines both macro-level and micro-level results to explain how firm recovery patterns differed between Germany and Sweden during and after the COVID-19 crisis.
Overall, the two countries showed distinct adjustment behaviours in profitability (ROA) and liquidity (Current Ratio, CR), reflecting their financial and institutional environments.

At the macro level, Germany displayed stronger resilience.
Several industries, such as Financial & Insurance and Manufacturing, showed steady or quickly recovering ROA–CR relationships.
It suggests that firms maintained more stable capital and liquidity structures.
In contrast, Sweden exhibited larger variability and stronger K-shaped dynamics, especially in Information & Communication and Consumer Services, where some firms recovered rapidly while others experienced severe declines.
These results indicate that Swedish industries were more sensitive to shocks and policy uncertainty.

At the micro level, scagnostic patterns further support this interpretation.
German industries showed more stable or even declining variability in metrics such as clumpy and stringy, suggesting relatively coherent behaviour among firms.
Meanwhile, Swedish industries showed rising outlying and higher residual dispersion, revealing more fragmented recovery paths.
In other words, Swedish firms reacted more heterogeneously within the same sector, consistent with the idea of uneven or K-shaped recovery.

Bringing these results together, the comparison highlights that Germany’s corporate environment may have provided stronger structural support for liquidity management and profitability recovery, while Sweden’s more market-driven system allowed greater flexibility but also led to higher volatility.
This difference suggests that national financial systems and policy responses played a key role in shaping how firms adjusted after the pandemic shock.

## Conclusions

### Question 1

Overall, the analysis shows that profitability remained mostly stable in both Germany and Sweden during the pandemic years compared with the pre-pandemic period.
Median values for EBIT, EBITDA, ROA, ROE, and net income stayed close to zero, indicating that most firms kept similar levels of profitability.
However, the variation across firms increased during 2020 – 2021, especially in Sweden, where more companies experienced bigger profits or deeper losses.
At the industry level, the median EBITDA margin changed only slightly for most sectors, but Accommodation & Food saw the strongest recovery while Health Care declined the most.
Technology, Utilities, and Education remained relatively stable across both countries.
These results suggest that although the overall trend in profitability didn’t shift much, the pandemic led to more uneven outcomes across firms and industries, with service-based sectors showing higher volatility.

### Question 2

AAA

### Question 3

AAA

### Question 4

The analysis shows that the pandemic affected the link between corporate liquidity and profitability in complex and uneven ways.
Germany demonstrated overall resilience, with relatively stable adjustments in most industries.
Firms maintained consistent ROA–CR structures and showed limited volatility, suggesting strong financial discipline and institutional support.
In contrast, Sweden experienced more volatile but also more adaptive responses.
The stronger fluctuations in scagnostic metrics and residual dispersion indicate that Swedish firms reacted more dynamically, allowing some to recover quickly while others struggled.
It is a clear sign of K-shaped recovery in several sectors.

Overall, the comparison suggests that national environments and policy frameworks played an important role in shaping how firms balanced profitability and liquidity during the crisis.
Germany’s stability-oriented system promoted consistent recovery, while Sweden’s more flexible market structure encouraged diversity in outcomes.
These findings highlight that firm-level resilience depends not only on financial management but also on broader institutional conditions.

## References

The materials used for this report are:

-   Bureau van Dijk.
    (2025).
    OSIRIS \[Data set: Germany subset\].
    Bureau van Dijk – A Moody’s Analytics Company.
    https://www.bvdinfo.com/en-gb/our-products/data/international/osiris

-   **Tidyverse:** Wickham H, Averick M, Bryan J, Chang W, McGowan LD, François R, Grolemund G, Hayes A, Henry L, Hester J, Kuhn M, Pedersen TL, Miller E, Bache SM, Müller K, Ooms J, Robinson D, Seidel DP, Spinu V, Takahashi K, Vaughan D, Wilke C, Woo K, Yutani H (2019).
    “Welcome to the tidyverse.” *Journal of Open Source Software*, *4*(43), 1686.
    doi:10.21105/joss.01686 <https://doi.org/10.21105/joss.01686>.

-   **Conflicted:** Wickham H (2023).
    *conflicted: An Alternative Conflict Resolution Strategy*.
    doi:10.32614/CRAN.package.conflicted <https://doi.org/10.32614/CRAN.package.conflicted>, R package version 1.2.0, <https://CRAN.R-project.org/package=conflicted>.

-   **Dplyr:** Wickham H, François R, Henry L, Müller K, Vaughan D (2023).
    *dplyr: A Grammar of Data Manipulation*.
    doi:10.32614/CRAN.package.dplyr <https://doi.org/10.32614/CRAN.package.dplyr>, R package version 1.1.4, <https://CRAN.R-project.org/package=dplyr>.

-   **Stringr:** Wickham H (2025).
    *stringr: Simple, Consistent Wrappers for Common String Operations*.
    doi:10.32614/CRAN.package.stringr <https://doi.org/10.32614/CRAN.package.stringr>, R package version 1.5.2, <https://CRAN.R-project.org/package=stringr>.

-   **Janitor:** Firke S (2024).
    *janitor: Simple Tools for Examining and Cleaning Dirty Data*.
    doi:10.32614/CRAN.package.janitor <https://doi.org/10.32614/CRAN.package.janitor>, R package version 2.2.1, <https://CRAN.R-project.org/package=janitor>.

-   **Skimr:** Waring E, Quinn M, McNamara A, Arino de la Rubia E, Zhu H, Ellis S (2025).
    *skimr: Compact and Flexible Summaries of Data*.
    doi:10.32614/CRAN.package.skimr <https://doi.org/10.32614/CRAN.package.skimr>, R package version 2.2.1, <https://CRAN.R-project.org/package=skimr>.

-   **Tsibble:** Wang, E, D Cook, and RJ Hyndman (2020).
    A new tidy data structure to support exploration and modeling of temporal data, Journal of Computational and Graphical Statistics, 29:3, 466-478, doi:10.1080/10618600.2019.1695624.

-   **Knitr:** Xie Y (2025).
    *knitr: A General-Purpose Package for Dynamic Report Generation in R*.
    R package version 1.50, <https://yihui.org/knitr/>.

    Yihui Xie (2015) Dynamic Documents with R and knitr.
    2nd edition.
    Chapman and Hall/CRC.
    ISBN 978-1498716963

    Yihui Xie (2014) knitr: A Comprehensive Tool for Reproducible Research in R.
    In Victoria Stodden, Friedrich Leisch and Roger D. Peng, editors, Implementing Reproducible Computational Research.
    Chapman and Hall/CRC.
    ISBN 978-1466561595

-   **Naniar:** Tierney N, Cook D (2023).
    “Expanding Tidy Data Principles to Facilitate Missing Data Exploration, Visualization and Assessment of Imputations.” *Journal of Statistical Software*, *105*(7), 1-31.
    doi:10.18637/jss.v105.i07 <https://doi.org/10.18637/jss.v105.i07>.

-   **Patchwork:** Pedersen T (2025).
    *patchwork: The Composer of Plots*.
    doi:10.32614/CRAN.package.patchwork <https://doi.org/10.32614/CRAN.package.patchwork>, R package version 1.3.2, <https://CRAN.R-project.org/package=patchwork>.

-   **Plotly:** C. Sievert.
    Interactive Web-Based Data Visualization with R, plotly, and shiny.
    Chapman and Hall/CRC Florida, 2020.

-   **tsibbletalk:** Wang E, Cook D (2020).
    *tsibbletalk: Interactive Graphics for Tsibble Objects*.
    doi:10.32614/CRAN.package.tsibbletalk <https://doi.org/10.32614/CRAN.package.tsibbletalk>, R package version 0.1.0, <https://CRAN.R-project.org/package=tsibbletalk>.

-   **Feasts:** O'Hara-Wild M, Hyndman R, Wang E (2025).
    *feasts: Feature Extraction and Statistics for Time Series*.
    doi:10.32614/CRAN.package.feasts <https://doi.org/10.32614/CRAN.package.feasts>, R package version 0.4.2, <https://CRAN.R-project.org/package=feasts>.

-   **Broom:** Robinson D, Hayes A, Couch S (2025).
    *broom: Convert Statistical Objects into Tidy Tibbles*.
    doi:10.32614/CRAN.package.broom <https://doi.org/10.32614/CRAN.package.broom>, R package version 1.0.9, <https://CRAN.R-project.org/package=broom>.

-   **Purrr:** Wickham H, Henry L (2025).
    *purrr: Functional Programming Tools*.
    doi:10.32614/CRAN.package.purrr <https://doi.org/10.32614/CRAN.package.purrr>, R package version 1.1.0, <https://CRAN.R-project.org/package=purrr>.

-   **Cassowaryr:** L. Wilkinson, A. Anand and R.
    Grossman, Graph-theoretic scagnostics, IEEE Symposium on Information Visualization, 2005.
    INFOVIS 2005., 2005, pp. 157-164, doi: 10.1109/INFVIS.2005.1532142.

    L. Wilkinson and G.
    Wills (2008) Scagnostics Distributions, Journal of Computational and Graphical Statistics, 17(2), pp 473-491, doi:10.1198/106186008X320465

    K. Grimm, Kennzahlenbasierte Grafikauswahl, doctoral thesis, Universitat Augsburg, 2016.

    H. Mason, Lee, S., Laa, U.
    and Cook, D.

    (2022) cassowaryr: Compute Scagnostics on Pairs of Numeric Variables in a Data Set. R package version 2.0.0. https://CRAN.R-project.org/package=cassowary

-   **Gganimate:** Pedersen T, Robinson D (2025).
    *gganimate: A Grammar of Animated Graphics*.
    doi:10.32614/CRAN.package.gganimate <https://doi.org/10.32614/CRAN.package.gganimate>, R package version 1.0.11, <https://CRAN.R-project.org/package=gganimate>.

-   **Gifski:** Ooms J, Kornel Lesiński, Authors of the dependency Rust crates (2025).
    *gifski: Highest Quality GIF Encoder*.
    doi:10.32614/CRAN.package.gifski <https://doi.org/10.32614/CRAN.package.gifski>, R package version 1.32.0-2, <https://CRAN.R-project.org/package=gifski>.

-   **KableExtra:** Zhu H (2024).
    *kableExtra: Construct Complex Table with 'kable' and Pipe Syntax*.
    doi:10.32614/CRAN.package.kableExtra <https://doi.org/10.32614/CRAN.package.kableExtra>, R package version 1.4.0, <https://CRAN.R-project.org/package=kableExtra>.

-   **Ggplot2:** H. Wickham.
    ggplot2: Elegant Graphics for Data Analysis.
    Springer-Verlag New York, 2016.

-   **Scales:** Wickham H, Pedersen T, Seidel D (2025).
    *scales: Scale Functions for Visualization*.
    doi:10.32614/CRAN.package.scales <https://doi.org/10.32614/CRAN.package.scales>, R package version 1.4.0, <https://CRAN.R-project.org/package=scales>.

-   **Tidyr:** Wickham H, Vaughan D, Girlich M (2024).
    *tidyr: Tidy Messy Data*.
    doi:10.32614/CRAN.package.tidyr <https://doi.org/10.32614/CRAN.package.tidyr>, R package version 1.3.1, <https://CRAN.R-project.org/package=tidyr>.

-   **Ggrepel:** Slowikowski K (2024).
    *ggrepel: Automatically Position Non-Overlapping Text Labels with 'ggplot2'*.
    doi:10.32614/CRAN.package.ggrepel <https://doi.org/10.32614/CRAN.package.ggrepel>, R package version 0.9.6, <https://CRAN.R-project.org/package=ggrepel>.

-   **ViridisLite:** Simon Garnier, Noam Ross, Robert Rudis, Antônio P. Camargo, Marco Sciaini, and Cédric Scherer (2023).
    viridis(Lite) - Colorblind-Friendly Color Maps for R.
    viridisLite package version 0.4.2.

-   The Associated Press.
    (2020).
    Europe braces for next wave of coronavirus pandemic in Berlin.
    AP News.
    https://apnews.com/article/coronavirus-pandemic-health-europe-epidemics-berlin-b61de99739774c1f52b4ba6860054d6d

## AI Acknowledgement

The links to my use of ChatGPT and other AI for help on this assignment are:

-   <https://chatgpt.com/share/68f563ee-9040-8010-ab1c-4969cfb31ee0> Used to help know when the pandemic happened in Germany, understand the accounting terms related to profitability, how to create a violin chart and jitter, how to make a dumbbell chart for comparison, and perform imputation on the data.

-   <https://chatgpt.com/share/68f4378d-bb24-8003-8471-ed2147a8fab3> Used for formatting assistance, debugging R code, and improving interpretation flow.

-   <https://chatgpt.com/share/68f3046d-1ad8-800d-bc11-1b005a9e600f> helped to implement animated graphics and optimize content.

-   <https://chatgpt.com/share/68f304c8-a088-800d-be54-939dfcd7e278> helped to evaluate whether the questions are answered adequately and where to make revisions.

-   <https://chatgpt.com/share/68f305ad-f15c-800d-b724-fc917c1a93ef> helped to deal with the problem that package ‘tignostics’ is not available for this version of R and how to use scagnostics.

-   <https://chatgpt.com/share/68f571d7-0810-8011-960e-b969c4d9a752> helped us format the explanation in a better and concise manner.
