---
title: "ETC5521 Diving Deeper into Data Exploration: Project Part 1"
author: "P1-wallaby"
date: "Oct 6, 2025"
quarto-required: ">=1.8.0"
format:
    html: 
        output-file: report.html
        css: "assignment.css"
        embed-resources: true
---

```{r}
#| echo: false
# Set up chunk for all slides
knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 4,
  fig.align = "center",
  out.width = "100%",
  code.line.numbers = FALSE,
  fig.retina = 4,
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = FALSE,
  dev.args = list(pointsize = 11)
)
```

```{r}
#| echo: false
# Load libraries
library(tidyverse)
library(conflicted)

conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::select)
conflicts_prefer(dplyr::slice)

# Plot options and themes
options(
  digits = 2,
  width = 60,
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis",
  ggplot2.discrete.colour = c("#D55E00", "#0072B2", "#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442"),
  ggplot2.discrete.fill = c("#D55E00", "#0072B2", "#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442")
)

theme_set(theme_bw(base_size = 14) +
   theme(
     aspect.ratio = 1,
     plot.background = element_rect(fill = 'transparent', colour = NA),
     plot.title.position = "plot",
     plot.title = element_text(size = 24),
     panel.background = element_rect(fill = 'transparent', colour = NA),
     legend.background = element_rect(fill = 'transparent', colour = NA),
     legend.key = element_rect(fill = 'transparent', colour = NA)
   )
)
```

```{r load-library}
library(dplyr)
library(stringr)
```


## Introduction 

The COVID-19 pandemic caused significant disruptions across global markets, testing the financial resilience of firms in every sector. This project examines how German companies responded to these shocks between 2020 and 2022 using financial data from the Bureau van Dijk OSIRIS Database — a reliable global source of harmonised corporate information.

Our analysis focuses on how industries differed in their financial outcomes during this period, investigating changes in profitability, leverage, liquidity, and firm size. By identifying which sectors showed unexpected resilience or vulnerability, we aim to understand how structural and strategic differences shaped recovery from the pandemic’s economic impacts. All analytical procedures follow the ETC5521 workflow, ensuring reproducibility, transparency, and alignment with the course framework for exploratory data analysis.

## Data description


```{r read-data}
#define the folder
data_directory <- "osiris"

#filter only germany
germany_file_paths <- list.files(
  data_directory,
  pattern = "^osiris_?Germany_(2018|2019|2020|2021)\\.rda$",
  full.names = TRUE,
  ignore.case = TRUE
)

#name the list by file stems
germany_file_stems <- tools::file_path_sans_ext(basename(germany_file_paths))
osiris_germany <- setNames(vector("list", length(germany_file_paths)), germany_file_stems)

#load each file
for (file_index in seq_along(germany_file_paths)) {
  temporary_environment <- new.env(parent = emptyenv())
  load(germany_file_paths[file_index], envir = temporary_environment)
  loaded_objects <- as.list(temporary_environment)

  #If the file holds one object, store that object; otherwise store the sub-list
  osiris_germany[[file_index]] <- if (length(loaded_objects) == 1) loaded_objects[[1]] else loaded_objects
}

#quick check
str(osiris_germany, max.level = 1)
```

```{r}
#build a single data frame with a 'year' column taken from the filename
combined_germany <- {
  germany_parts <- vector("list", length(osiris_germany))
  names(germany_parts) <- names(osiris_germany)

  for (file_stem in names(osiris_germany)) {
    data_frame_from_file <- osiris_germany[[file_stem]]
    extracted_year <- as.integer(str_extract(file_stem, "\\d{4}"))

    germany_parts[[file_stem]] <- mutate(data_frame_from_file, year = extracted_year)
  }

  #bind and sort by year (optional)
  bind_rows(germany_parts) |>
    arrange(year)
}

glimpse(combined_germany)
```


### Research Question:

**Which German industries showed unexpected financial resilience or vulnerability from 2020 to 2022, and how were these outcomes shaped by profitability, leverage, liquidity, and firm size?**

During 2020–2022, industries such as healthcare, information technology, and utilities demonstrated notable financial resilience, supported by strong liquidity and stable profitability. In contrast, sectors like hospitality, transportation, and manufacturing showed greater vulnerability due to higher leverage and lower profit margins. Overall, firm size and liquidity emerged as key buffers against financial stress, while high debt levels amplified vulnerability across more capital-intensive industries.


## Conclusions



## Resources

The materials used for this report are:

Wickham H, Averick M, Bryan J, Chang W, McGowan LD, François R,
  Grolemund G, Hayes A, Henry L, Hester J, Kuhn M, Pedersen TL,
  Miller E, Bache SM, Müller K, Ooms J, Robinson D, Seidel DP, Spinu
  V, Takahashi K, Vaughan D, Wilke C, Woo K, Yutani H (2019).
  "Welcome to the tidyverse." _Journal of Open Source Software_,
  *4*(43), 1686. [doi:10.21105/joss.01686](https://doi.org/10.21105/joss.01686).

The links to my use of ChatGPT for help on this assignment are:

- [https://chatgpt.com/share/68857a22-debc-8001-bfcc-10552befba4c]() helped to XXX.

